\providecommand{\Abulk}{A_{\mathrm{bulk}}}
\providecommand{\Bbound}{B_{\partial}}
\providecommand{\Zder}{\operatorname{Z}_{\mathrm{der}}}
\providecommand{\mc}{\operatorname{MC}}
\providecommand{\gSC}{\mathfrak{g}^{\mathrm{SC}}}
\providecommand{\Res}{\operatorname{Res}}
\providecommand{\HH}{\operatorname{HH}}
\providecommand{\RHom}{\operatorname{RHom}}

\section{Three-dimensional gravity from $E_1$-chiral Koszul duality}
\label{sec:3d-gravity}
\index{3d gravity|textbf}
\index{Chern--Simons!gravitational}
\index{Brown--Henneaux central charge}
\index{BTZ black hole}

This section derives the complete holographic dictionary of
three-dimensional quantum gravity from a single algebraic input:
the Virasoro $\lambda$-bracket
\begin{equation}\label{eq:gravity-input}
\{T{}_\lambda T\}
\;=\;
\partial T + 2T\lambda + \frac{c}{12}\,\lambda^3.
\end{equation}
Three terms, one generator, one parameter $c$.  The quartic pole
($\lambda^3$ in the $\lambda$-bracket, equivalently
$(z-w)^{-4}$ in the OPE) is the single datum that generates the
entire structure: the non-associativity of $m_2$, the infinite
$\Ainf$ tower $\{m_k\}_{k \ge 3}$, the central charge shift
$c \mapsto 26-c$ under Koszul duality, the gravitational
Laplace kernel $r^L(z) = (c/2)/z^4 + 2T/z^2 + \partial T/z$
(equivalently, collision residue $r^{\mathrm{coll}}(z)
= (c/2)/z^3 + 2T/z$), and the
genus tower controlled by the curvature $\kappa = c/2$.

The development unfolds in five movements, each forced by the
previous:
\begin{enumerate}[label=\textup{(\Roman*)}]
\item The $\lambda$-bracket determines $m_2$, whose
  non-associativity (from the quartic pole) forces $m_3$, which
  forces $m_4$, and so on: the $\Ainf$ structure is genuinely
  infinite.
\item The bar complex $\barB(\mathrm{Vir}_c)$ is a dg coalgebra
  encoding this $\Ainf$ structure; Verdier duality converts this
  coalgebra to an algebra, and on the Koszul locus the resulting
  algebra is $\mathrm{Vir}_{26-c}$ (the Koszul dual).  Line operators
  are $\mathrm{Vir}_{26-c}$-$\mathbf{mod}$.
\item The Laplace transform of the $\lambda$-bracket produces
  the classical $r$-matrix, whose Yang--Baxter equation is Arnold
  cancellation on $\FM_3(\C)$.
\item The genus tower arises from the curvature
  $d_{\mathrm{fib}}^2 = \kappa_{\mathrm{eff}}\cdot\omega_1$,
  itself a consequence of trace and clutching on the
  gravitational open-sector category
  $\cC_{\mathrm{grav}}^{\mathrm{op}}$;
  the scalar all-genus generating function is the
  $\hat A$-genus weighted by $\kappa_{\mathrm{eff}} = (c-26)/2$
  (the $\hat{A}$-genus, not the Todd class, because the Hodge
  bundle carries a real structure via Serre duality, forcing the
  KO-theory orientation; $\hat{A}(x)^2 = \mathrm{Td}(x)\cdot
  \mathrm{Td}(-x)$).
\item The Koszul triangle closes with boundary
  $\mathrm{Vir}_c$, bulk $\C[\![c]\!]$, and an abstract
  line-side package conjecturally identified with modules for
  $\mathrm{Vir}_{26-c}$
  (Conjecture~\ref{conj:gravity-line-identification}); the
  boundary face recovers
Brown--Henneaux.  The two-sided Cardy asymptotic entropy formula would be expected
from the Virasoro bar-complex argument via modular $S$-transform
  and saddle-point extraction, and the BTZ Bekenstein--Hawking entropy
  formula $S_{\mathrm{BTZ}} = 2\pi\sqrt{ch/6} + 2\pi\sqrt{c\bar h/6}$
  would then follow from that two-sided Cardy asymptotic entropy formula;
  Verdier duality and the
  modular $S$-transform commute on genus-$1$ MC elements.
\end{enumerate}
At $c = 0$ (no quartic pole), the scalar curvature disappears:
the dual point is the Witt algebra on the uncurved scalar locus,
and the scalar genus package vanishes.  This does not prove that the
full $\Ainf$ structure is trivial or that the full higher-arity tower
vanishes.  The quartic pole is the engine of the nonlinear corrections.
This is an instance of the pole-order principle
(\S\ref{sec:pole-order-principle}): the maximal OPE pole order
controls the $\Ainf$ complexity class.  Affine Kac--Moody, with
poles of order at most~$2$, lives in class~$\mathbf{L}$: the
bar differential is finite, and the shadow tower terminates.
Virasoro, with its quartic pole, lives in class~$\mathbf{M}$:
the bar differential is infinite, and three-dimensional gravity
acquires its full perturbative hierarchy.  Drinfeld--Sokolov
reduction manufactures the quartic pole from the double pole ---
the Sugawara double Wick contraction --- and thereby transports
$\mathbf{L} \to \mathbf{M}$.

\begin{proposition}[Pole-order classification; \ClaimStatusProvedHere]%
\label{prop:pole-order-classification}%
\index{pole-order classification|textbf}%
\index{class L@class $\mathbf{L}$}%
\index{class M@class $\mathbf{M}$}%
Let $\cA$ be a one-generator chiral algebra with OPE of maximal
pole order~$p$.
\begin{enumerate}
\item If $p = 2$ \textup{(}double pole\textup{)}, then $\cA$ is
  of class~$\mathbf{L}$: the $\lambda$-bracket is associative
  \textup{(}$A_3 = 0$\textup{)}, the $\Ainf$ tower terminates at
  arity~$2$, and the dg-shifted Yangian
  $\Ydg(\cA)$ is a strict filtered deformation
  of $U(\fg[t])$ with simple-pole $r$-matrix
  $r(z) = \Omega/((k+h^\vee)z)$.
  \textup{Example}: Heisenberg $\cH_k$.  \textup{(}The multi-generator
  affine Kac--Moody algebra $V_k(\fg)$ is also class~$\mathbf{L}$:
  inter-generator OPEs have at most simple poles, so the
  multi-generator associator vanishes by the same
  mechanism.\textup{)}
\item If $p = 3$ \textup{(}cubic pole\textup{)}, then $\cA$ is
  of class~$\mathbf{C}$: the quartic contact invariant $m_4 \ne 0$
  but the tower terminates at shadow depth $r_{\max} = 4$ by
  rank-one abelian rigidity.
  \textup{Example}: $\beta\gamma$ system.
\item If $p = 4$ \textup{(}quartic pole\textup{)}, then $\cA$ is
  of class~$\mathbf{M}$: the $\lambda$-bracket is non-associative
  \textup{(}$A_3 \neq 0$ with scalar part proportional to
  $c/12$\textup{)}, the $\Ainf$ tower is infinite, and
  $\Ydg(\cA)$ is a curved $\Ainf$-deformation with
  cubic-pole collision residue
  $r(z) = (c'/2)/z^3 + 2T/z$.
  \textup{Example}: $\mathrm{Vir}_c$.
\item Drinfeld--Sokolov reduction transports
  $\mathbf{L} \to \mathbf{M}$: the Sugawara construction
  manufactures the quartic pole from the double pole via the
  double Wick contraction, and the DS contracting homotopy
  generates the infinite $\Ainf$ tower from the formal input
  through the resolvent tree formula.
\end{enumerate}
\end{proposition}

\begin{proof}
Part~(1): $p = 2$ means $\{a{}_\lambda a\} = c_1\lambda + c_0$
with $c_1$ central and $c_0 = a_{(0)}a$.  By antisymmetry,
$c_0 = [a,a] = 0$ for a single generator.  The iterated
bracket $\{a{}_\mu\{a{}_\lambda a\}\} = \{a{}_\mu\, c_1\lambda\} = 0$
(since $c_1\lambda$ is proportional to the identity,
$\{a{}_\mu\,\mathbf{1}\} = 0$ by the PVA vacuum axiom),
so $A_3 = 0$.  The shadow tower
terminates by the depth formula
$d' = \lfloor(p-1)/2\rfloor = 0$.  For multi-generator algebras at
$p = 2$, the discriminant is the \emph{centrality} of the
double-pole coefficient: if the residue $c_0 = \{a_{(0)} b\}$ is
central (Heisenberg, class~$\mathbf{G}$, depth~$2$) the
$\lambda$-bracket is abelian; if $c_0$ is non-central (affine KM,
class~$\mathbf{L}$, depth~$3$) the Lie bracket generates $m_3
\ne 0$ through the Jacobi identity.
Part~(2): the quartic pole $a_{(3)}a = c/2$ forces
$A_3 \neq 0$ (cf.~\eqref{eq:gravity-associator} below), which
forces $m_3 = -A_3 \neq 0$; the Stasheff identity then forces
$m_4, m_5, \ldots$ in perpetuity.
Part~(3) is the DS transfer theorem
(Theorem~\ref{thm:ds-hpl-transfer}).
\end{proof}

\begin{remark}[$\mathrm{Vir}_0$: uncurved but non-formal]
\label{rem:vir-zero-type-V}
\index{Witt algebra!uncurved non-formal}
At $c = 0$ the quartic pole vanishes and the $\lambda$-bracket
reduces to $\{T{}_\lambda T\} = \partial T + 2T\lambda$.  The
curvature $\kappa(\mathrm{Vir}_0) = 0$, so the bar complex is
flat at every genus: $d^2 = 0$.  However, the $\Ainf$ tower
does \emph{not} truncate: the non-central double-pole coefficient
$2T$ (a field, not a scalar) produces a non-vanishing associator
$A_3 \ne 0$, forcing $m_3 \ne 0$ and hence
$m_k \ne 0$ for all $k \ge 3$.  The scalar shadow free energies
vanish ($F_g = 0$ for all $g \ge 1$), but the $T$-dependent
terms in every $m_k$ survive unchanged (they are $c$-independent).
$\mathrm{Vir}_0$ is therefore uncurved ($d^2 = 0$ at every genus,
since $\kappa = c/2 = 0$ gives $d_{\mathrm{fib}}^2 = \kappa \cdot
\omega_g = 0$ by Vol~I Theorem~D) but non-formal (the $\Ainf$ tower
$\{m_k\}_{k \ge 3}$ does not truncate).  The Koszul dual is
$\mathrm{Vir}_0^! = \mathrm{Vir}_{26-0} = \mathrm{Vir}_{26}$
(Proposition~\ref{prop:wn-complementarity} at $N = 2$),
which sits at the critical string, and the passage
$c = 0 \leftrightarrow c = 26$ under the Koszul involution maps
the uncurved algebra to the topological point.
\end{remark}

\begin{corollary}[Gauge-gravity complexity dichotomy; \ClaimStatusProvedHere]%
\label{cor:gauge-gravity-dichotomy}%
\index{gauge-gravity dichotomy|textbf}%
The two standard HT families occupy opposite corners of the
$(\Ainf,\,\Delta)$-complexity plane:
\[
\begin{array}{c|cc}
& \Ainf\textup{ products } m_k\;(k \ge 3) &
  \textup{coproduct } \Delta_z \\
\hline
\textup{Gauge \textup{(}CS\textup{)}}
  & m_k = 0\;\textup{(}formal\textup{)}
  & \textup{nontrivial Yangian } \Delta_z \\
\textup{Gravity \textup{(}Vir\textup{)}}
  & m_k \neq 0\;\forall\, k\;\textup{(}non-formal\textup{)}
  & \textup{strictly primitive } \Delta_z
\end{array}
\]
Drinfeld--Sokolov reduction transports between corners:
it preserves chiral Koszulness but destroys Swiss-cheese
formality.  The Sugawara double Wick contraction manufactures
the quartic pole; the DS contracting homotopy~$h$ generates
the infinite $\Ainf$ tower via the resolvent tree formula;
and the ghost-number obstruction
($|h| = -1$ in ghost degree) kills all non-primitive
coproduct terms.
\end{corollary}

\begin{proof}
The $\Ainf$ column follows from
Proposition~\ref{prop:pole-order-classification} (class~$\mathbf{L}$
formal, class~$\mathbf{M}$ non-formal).  The coproduct column:
for gauge theory $\Delta_z$ is the standard Yangian coproduct
(non-primitive, encoding the spectral braiding); for gravity
$\Delta_z(x) = \tau_z(x) \otimes 1 + 1 \otimes x$ is primitive
(Theorem~\ref{thm:gravitational-primitivity}).  The DS transport
is Theorem~\ref{thm:ds-hpl-transfer}.
\end{proof}

%====================================================================
% MOVEMENT I: THE A-INFINITY STRUCTURE
%====================================================================

\subsection{Movement I: the $\Ainf$ structure from the quartic pole}
\label{subsec:gravity-ainf}

\subsubsection*{The binary operation $m_2$}

The $\lambda$-bracket~\eqref{eq:gravity-input} IS the binary
operation:
\begin{equation}\label{eq:gravity-m2}
m_2(T, T;\, \lambda)
\;=\;
\{T{}_\lambda T\}
\;=\;
\frac{c}{12}\,\lambda^3
\;+\; 2T\,\lambda
\;+\; \partial T.
\end{equation}
The three terms correspond to the three OPE poles:
quartic $\mapsto$ $\frac{c}{12}\lambda^3$,
double $\mapsto$ $2T\lambda$,
simple $\mapsto$ $\partial T$.
In the bar complex: the differential on degree-$2$ elements is
$d_{\barB}[s^{-1}T \,|\, s^{-1}T] = \partial T$ (the FM boundary
integral over $E_{12}$ extracts the simple-pole mode $T_{(0)}T$).

\subsubsection*{The ternary operation $m_3$: forced by non-associativity}

Compute the associator.  By the PVA Jacobi identity
$A_3(\ell,\mu) = -\{T_\mu\{T_\ell T\}\}$, or
equivalently by direct sesquilinearity computation
(see~\eqref{eq:vir-associator}):
\begin{equation}\label{eq:gravity-associator}
\boxed{
A_3(T,T,T;\,\lambda_{12},\lambda_{23})
\;=\;
-\partial^2 T
\;-\; (2\lambda_{12} + 3\lambda_{23})\,\partial T
\;-\; 2\lambda_{23}(2\lambda_{12} + \lambda_{23})\,T
\;-\; \frac{c}{12}\lambda_{23}^3(2\lambda_{12} + \lambda_{23}).}
\end{equation}
The scalar term $\tfrac{c}{12}\lambda_{23}^3(2\lambda_{12}+\lambda_{23})$
is proportional to the central charge: for affine Kac--Moody
(where the quartic pole vanishes) $A_3 = 0$ and the $\lambda$-bracket
is strictly associative.

The full vertex algebra OPE (singular $+$ regular) is associative
(Borcherds identity).  The non-associativity $A_3 \neq 0$ of
the singular part ($\lambda$-bracket) is compensated by the regular
part (the normally ordered product $:TT:$), giving $m_3 = -A_3$:
\begin{equation}\label{eq:gravity-m3}
m_3(T,T,T;\,\lambda_{12},\lambda_{23})
\;=\;
\partial^2 T
\;+\; (2\lambda_{12} + 3\lambda_{23})\,\partial T
\;+\; 2\lambda_{23}(2\lambda_{12} + \lambda_{23})\,T
\;+\; \frac{c}{12}\lambda_{23}^3(2\lambda_{12} + \lambda_{23}).
\end{equation}
The four terms have direct gravitational content:
$\partial^2 T$ is the second-order geometric correction;
$(2\lambda_{12}+3\lambda_{23})\partial T$ is the first-order response;
$2\lambda_{23}(2\lambda_{12}+\lambda_{23})T$ is the stress-tensor backreaction;
$\tfrac{c}{12}\lambda_{23}^3(2\lambda_{12}+\lambda_{23})$
is the three-graviton coupling ($\propto 1/G$).

\subsubsection*{The quaternary operation $m_4$: forced by the Stasheff identity}

The arity-$4$ Stasheff identity on $\FM_4(\C) \cong K_4$ involves five
compositions: two from $(r,s) = (2,3)$ and three from $(r,s) = (3,2)$,
with Koszul signs $(-1)^j$ from the desuspended degree $|sT| = -1$.
The obstruction $\mathrm{Obs}_4$ is assembled from all compositions of
$m_2$ and $m_3$, and $m_4 = -\mathrm{Obs}_4$ is determined uniquely on the
generator~$T$ by sesquilinearity.

\begin{proposition}[Quaternary Virasoro $\Ainf$ operation; \ClaimStatusProvedHere]
\label{prop:gravity-m4}
The quaternary $\Ainf$ operation for the Virasoro algebra is
\begin{equation}\label{eq:gravity-m4}
\begin{aligned}
m_4(T,T,T,T;\,&\lambda_{12},\lambda_{23},\lambda_{34})
\\[3pt]
&=\;
(\lambda_{23} + 2\lambda_{34})\,\partial^2 T
\\[3pt]
&\quad +\;
(5\lambda_{12}^2 - 4\lambda_{12}\lambda_{23}
  + 3\lambda_{12}\lambda_{34}
  - 5\lambda_{23}^2 + 10\lambda_{23}\lambda_{34}
  + 4\lambda_{34}^2)\,\partial T
\\[3pt]
&\quad +\;
2(\lambda_{12}^3 + 2\lambda_{12}^2\lambda_{23}
  + 3\lambda_{12}^2\lambda_{34}
  - 2\lambda_{12}\lambda_{23}^2
  - 4\lambda_{12}\lambda_{23}\lambda_{34}
  + 4\lambda_{12}\lambda_{34}^2
\\
&\qquad\qquad
  - \lambda_{23}^3 - 2\lambda_{23}^2\lambda_{34}
  + 6\lambda_{23}\lambda_{34}^2)\,T
\\[3pt]
&\quad +\;
\frac{c}{12}
(\lambda_{12}^5 + 2\lambda_{12}^4\lambda_{23}
  + 3\lambda_{12}^4\lambda_{34}
  + 4\lambda_{12}^3\lambda_{23}\lambda_{34}
  + 2\lambda_{12}^3\lambda_{34}^2
\\
&\qquad\qquad
  - 2\lambda_{12}\lambda_{23}^4
  - 4\lambda_{12}\lambda_{23}^3\lambda_{34}
  - 4\lambda_{12}\lambda_{23}\lambda_{34}^3
  + 2\lambda_{12}\lambda_{34}^4
\\
&\qquad\qquad
  - \lambda_{23}^5 - 2\lambda_{23}^4\lambda_{34}
  + 6\lambda_{23}\lambda_{34}^4).
\end{aligned}
\end{equation}
At the symmetric point $\lambda_{12} = \lambda_{23} = \lambda_{34} = \lambda$:
\begin{equation}\label{eq:gravity-m4-sym}
m_4(T^4;\,\lambda,\lambda,\lambda)
\;=\;
3\lambda\,\partial^2 T
+ 13\lambda^2\,\partial T
+ 14\lambda^3\, T
+ \frac{7c}{12}\,\lambda^5.
\end{equation}
\end{proposition}

\begin{proof}
Solve the arity-$4$ Stasheff relation~\eqref{eq:gravity-associator}
with five terms (two from $(r,s)=(2,3)$, three from $(r,s)=(3,2)$),
using the sesquilinearity of $m_2$~\eqref{eq:gravity-m2} and
$m_3$~\eqref{eq:gravity-m3}.  The first-slot rule
$m_3(\partial^j T, \ldots) = (-\lambda_{12})^j m_3(T, \ldots)$
and the last-slot rule
$m_3(\ldots, \partial^j T) = (\lambda_{23}+\partial)^j m_3(\ldots, T)$
reduce every composition to a polynomial in
$\lambda_{12}, \lambda_{23}, \lambda_{34}$, verified by direct symbolic
computation.
\end{proof}

\noindent
\textit{Structural features.}
The leading derivative in $m_4$ is $\partial^2 T$ with a
\emph{linear} coefficient $(\lambda_{23}+2\lambda_{34})$ --- not $\partial^3 T$.
The absence of a $\partial^3 T$ term reflects the associahedron $K_4$
cell structure.
The central charge~$c$ enters exclusively in the scalar
contact term $\frac{c}{12}Q_4(\lambda)$, where $Q_4$ has degree~$5$;
all $T$-dependent coefficients are $c$-independent.
At zero spectral parameters $m_4 = 0$: no constant-mode quartic
self-coupling.

\subsubsection*{The quinary operation $m_5$: sign alternation and
  the infinite tower}

At arity~$5$, the Stasheff identity involves compositions from
$(r,s) = (2,4)$, $(3,3)$, and $(4,2)$, giving nine terms.
The $(4,2)$ compositions require the internal sesquilinearity
of~$m_4$, obtained by recomputing the obstruction with modified inputs.

\begin{proposition}[Quinary Virasoro $\Ainf$ operation; \ClaimStatusProvedHere]
\label{prop:gravity-m5}
The quinary $\Ainf$ operation for the Virasoro algebra is
\begin{equation}\label{eq:gravity-m5}
\begin{aligned}
m_5(T^5;\,
  &\lambda_{12},\lambda_{23},\lambda_{34},\lambda_{45})
\\[3pt]
&=\;
-\partial^4 T
\;-\;
(2\lambda_{12} + 4\lambda_{23}
  + \lambda_{34} + 3\lambda_{45})\,\partial^3 T
\\[3pt]
&\quad +\;
(3\lambda_{12}^2 - 16\lambda_{12}\lambda_{23}
  - 6\lambda_{12}\lambda_{45}
  - 13\lambda_{23}^2 + 7\lambda_{23}\lambda_{34}
\\
&\qquad\qquad
  - 9\lambda_{23}\lambda_{45}
  - 7\lambda_{34}^2
  + 3\lambda_{34}\lambda_{45}
  - 3\lambda_{45}^2)\,\partial^2 T
\\[3pt]
&\quad +\; P_1(\lambda)\,\partial T
\;+\; P_0(\lambda)\, T
\;+\; \frac{c}{12}\, Q_5(\lambda),
\end{aligned}
\end{equation}
where $P_1$, $P_0$ are explicit polynomials of degree~$3$, $4$
respectively in $\lambda_{12}, \ldots, \lambda_{45}$, and
$Q_5$ is a degree-$6$ polynomial whose symmetric-point value is
$Q_5(\lambda,\ldots,\lambda) = -34\lambda^6$.
At the symmetric point:
\begin{equation}\label{eq:gravity-m5-sym}
\begin{aligned}
m_5(T^5;\,\lambda,\lambda,\lambda,\lambda)
&\;=\;
-\partial^4 T
- 10\lambda\,\partial^3 T
- 41\lambda^2\,\partial^2 T
\\
&\quad
- 84\lambda^3\,\partial T
- 68\lambda^4\, T
- \frac{17c}{6}\,\lambda^6.
\end{aligned}
\end{equation}
All symmetric-point coefficients are negative.
\end{proposition}

\begin{proof}
Same method as Proposition~\ref{prop:gravity-m4}, now solving the
arity-$5$ Stasheff relation with nine compositions.  The new ingredient
is the internal sesquilinearity of $m_4$ (second and third slots),
extracted from the obstruction computation with modified inputs.
Verified by symbolic computation.
\end{proof}

\noindent
\textit{Structural features.}
The leading term $-\partial^4 T$ carries a minus sign --- the first
sign alternation in the tower.
Operations $m_2$, $m_3$ have positive leading terms ($+\partial T$,
$+\partial^2 T$); $m_4$ has no $\partial^3 T$ term; $m_5$ begins
with $-\partial^4 T$.
The sign alternation is the Koszul sign from the bar desuspension
$s^{-1}$.
Unlike $m_4$, the operation $m_5$ populates the full derivative range
$\partial^4 T$ through $T$ plus a scalar.
As for all lower operations, $c$ enters only through the scalar
contact term; all $T$-dependent coefficients are $c$-independent.

\subsubsection*{The degree pattern and gravitational Feynman rules}

The explicit computations $m_2$--$m_5$ reveal a uniform pattern:

\begin{center}
\small
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}cccc@{}}
\textbf{Operation}
  & \textbf{Leading term}
  & \textbf{$\deg_\lambda$ (max)}
  & \textbf{Scalar} \\
\hline
$m_2$
  & $\partial T$
  & $3$
  & $\frac{c}{12}\lambda^3$ \\[3pt]
$m_3$
  & $\partial^2 T$
  & $4$
  & $\frac{c}{12}\lambda_{23}^3(2\lambda_{12}+\lambda_{23})$ \\[3pt]
$m_4$
  & $(\lambda_{23}+2\lambda_{34})\,\partial^2 T$
  & $5$
  & $\frac{c}{12}Q_4(\lambda)$ \\[3pt]
$m_5$
  & $-\partial^4 T$
  & $6$
  & $\frac{c}{12}Q_5(\lambda)$
\end{tabular}
\end{center}

\noindent
Three structural features persist across all computed arities:
\begin{enumerate}[label=\textup{(\alph*)},nosep]
\item \textit{Central charge as contact term.}
  The central charge enters every $m_k$ through a single scalar term
  $\frac{c}{12}Q_k(\lambda)$, where $Q_k$ has degree $k+1$.
  All $T$-dependent coefficients are $c$-independent.
  This reflects the Virasoro OPE structure: the only
  $c$-dependent singular term is $(c/2)/(z-w)^4$, and the bar kernel
  $d\log(z-w)$ absorbs one power (AP19) to produce
  $(c/2)/(z-w)^3$ in the collision residue.
  Gravitationally: the stress-tensor backreaction
  ($T$, $\partial T$, $\partial^2 T$ terms) is universal ---
  independent of the cosmological constant ---
  while the pure-graviton scattering amplitude is proportional
  to $c/12 = \kappa/6$.

\item \textit{Degree pattern.}
  The coefficient of $\partial^j T$ in $m_k$ has spectral degree
  $k - 1 - j$; the scalar has degree $k+1$.
  The total weight $j + \deg_\lambda = k-1$ is constant across
  $T$-dependent terms.

\item \textit{Vanishing at $\lambda = 0$.}
  Every $m_k(T^k)$ vanishes when all spectral parameters are zero:
  the $\Ainf$ structure is purely spectral.
  At $c = 0$ (the Witt algebra), the scalar terms vanish but the
  $T$-dependent terms persist unchanged: the tower remains infinite
  even without the central charge.  The quartic OPE pole is the
  engine of the infinite tower.
\end{enumerate}

\subsubsection*{The quartic contact invariant and infinite depth}

At $n = 4$, $\FM_4(\C)$ is the associahedron $K_4$.  The Stasheff
relation produces the quartic contact invariant from the Kac
determinant at level~$4$:
\begin{equation}\label{eq:gravity-quartic}
\mathfrak{Q}^{\mathrm{contact}}_{\mathrm{Vir}}
\;=\;
\frac{10}{c(5c+22)}.
\end{equation}
The numerator $10$ arises from the Shapovalov inner product at level~$4$;
the denominator is the Kac determinant.  Poles at $c = 0$ ($G = \infty$) and $c = -22/5$
(Lee--Yang) mark reorganisation points of the shadow tower.
The contact invariant is the \emph{normalised} version of
$m_4$~\eqref{eq:gravity-m4}: it divides $m_4$ by the Shapovalov inner
product at level~$4$.

The $\Ainf$ structure does not truncate:
$m_k \ne 0$ for all $k \ge 3$ and generic~$c$.  Each
$m_k$ is computed by a wheel diagram (a Hamiltonian cycle on the
complete graph $K_k$ with $k-1$ cubic vertices), and the quartic
OPE pole ensures a nonzero residue at every arity
(Proposition~\ref{prop:vir-truncation}).

\begin{remark}[The $\Ainf$ tower is manufactured by DS reduction;
\ClaimStatusProvedHere]%
\label{rem:gravity-ds-manufactured}%
\index{Drinfeld--Sokolov reduction!manufactures $\Ainf$ tower}%
The infinite $\Ainf$ tower $\{m_k\}_{k \ge 3}$ of the Virasoro
algebra is not an irreducible datum: it is \emph{manufactured} by
Drinfeld--Sokolov reduction from the finite Kac--Moody tower.
The principal DS reduction of $V_k(\mathfrak{sl}_2)$ produces
$\mathrm{Vir}_{c_{\mathrm{DS}}}$ with
$c_{\mathrm{DS}} = 1 - 6(k+1)^2/(k+2)$, transporting the
complexity class from $\mathbf{L}$ (affine: $m_k = 0$ for $k \ge 3$,
finitely many nonzero shadows) to $\mathbf{M}$ (Virasoro: $m_k \ne 0$
for all $k \ge 3$, infinitely many nonzero shadows).
The mechanism is precise: the DS stress tensor
$T^{\mathrm{DS}} = J^f + (J^h)^2/(4(k+2)) + \tfrac{1}{2}\partial J^h$
contains a quadratic Sugawara term $(J^h)^2/(4(k+2))$ whose
double-Wick self-contraction generates the quartic OPE pole
$c_{\mathrm{DS}}/12$ --- a pole order absent in the original affine
OPE (which has at most double poles).  This manufactured quartic pole
makes $m_2$ non-associative ($A_3 \ne 0$), which forces $m_3 = -A_3
\ne 0$ by the Stasheff identity, which seeds the entire infinite tower.
The DS--bar intertwining theorem (Vol~I,
Theorem~ds-koszul-intertwine) ensures that this chain-level
verification lifts to a coalgebra isomorphism:
$\mathrm{DS}_\chi(\barB(V_k(\mathfrak{sl}_2))) \simeq
\barB(\mathrm{Vir}_{c_{\mathrm{DS}}})$.
The quartic pole of the Virasoro OPE --- the single datum that
generates the entire gravitational structure --- is a
Wick-contraction consequence of Sugawara, not an independent
input.
\end{remark}

\subsubsection*{The $c$-linearity theorem}

The explicit computations $m_2$--$m_5$ reveal that the central
charge enters each $m_k$ through a single scalar term
$\frac{c}{12}Q_k(\lambda)$, with all $T$-dependent coefficients
$c$-independent.  This is not an accident of low arity, nor
special to the Virasoro algebra: it is a structural theorem
governing the $\Ainf$ tower of any logarithmic $\SCchtop$-algebra
with a central charge (Proposition~\ref{prop:c-linearity-general}).
The Virasoro specialisation pins down the polynomial degree.

\begin{theorem}[$c$-linearity of the Virasoro $\Ainf$ tower; \ClaimStatusProvedHere]
\label{thm:gravity-c-linearity}
\index{c-linearity@$c$-linearity!Virasoro A-infinity@Virasoro $\Ainf$ tower|textbf}
Specialising Proposition~\ref{prop:c-linearity-general} to
$\cA = \mathrm{Vir}_c$:
the scalar part of the chain-level $m_k(T^{\otimes k};\,
\lambda_1, \ldots, \lambda_{k-1})$ is \textbf{linear} in~$c$
for all $k \ge 2$:
\begin{equation}\label{eq:gravity-c-linearity}
m_k\big|_{\textup{scalar}}
\;=\;
\frac{c}{12}\,P_k(\lambda_1, \ldots, \lambda_{k-1}),
\end{equation}
where $P_k$ is a homogeneous polynomial of degree $k+1$
in $(\lambda_1, \ldots, \lambda_{k-1})$.  All
$T$-dependent coefficients (the polynomials multiplying
$\partial^j T$ for $0 \le j \le k-2$) are independent of~$c$.
\end{theorem}

\begin{proof}
The general mechanism is
Proposition~\ref{prop:c-linearity-general}: the central
charge enters the $\lambda$-bracket as a scalar contact term,
scalars are annihilated by the bracket in both slots, and
induction on the Stasheff relations preserves $c$-linearity.
The Virasoro-specific content is the degree count:
$m_2|_{\text{scalar}} = (c/12)\lambda^3$ has degree $3 = 2+1$,
and the Stasheff composition raises the degree by~$2$ at each
arity step, giving $\deg P_k = k+1$.  The BRST homotopy $h$ from
the DS reduction of $\widehat{\mathfrak{sl}}_2$ is a
$c$-independent operator (it depends on the ghost system, not on
the central charge), so it preserves the $c$-linearity.
\end{proof}

\begin{remark}[Gravitational content of $c$-linearity]
\label{rem:gravity-c-linearity-physics}
The $c$-linearity theorem has a direct gravitational reading.
The central charge $c = 3\ell/(2G)$ is proportional to the
inverse Newton constant.  The theorem states that the
stress-tensor backreaction ($T$, $\partial T$, $\ldots$,
$\partial^{k-2}T$ terms) is \emph{universal} --- independent
of $G$ --- while the pure-graviton scattering amplitude (the
scalar contact term) is proportional to $c/12 = \kappa/6$.
Graviton self-interaction at any arity is controlled by a single
coupling constant.

On BRST cohomology, the Shapovalov inverse introduces Kac
determinant denominators, so the transferred operation
$m_k^H$ is rational in~$c$:
\[
m_k^H\big|_{\text{scalar}}
\;=\;
\frac{c \cdot P_k(\lambda_1, \ldots, \lambda_{k-1})}
     {D_k(c)},
\]
where $D_k(c) = \prod_{n=2}^{k}\prod_{rs \le n}
(h_{r,s}(c))^{p(n-rs)}\big|_{h=0}$ is the product of Kac
determinant factors at levels $\le k$.  The poles in~$c$ are
the loci where null vectors appear in the intermediate Virasoro
Verma module: they are \emph{resonances} of the graviton
self-energy.  The level-$4$ Kac determinant gives the quartic
contact invariant denominator $c(5c+22)$; higher levels
contribute the denominators of $m_k^H$ for $k \ge 5$.
\end{remark}

\subsubsection*{Finite presentation of the infinite tower}

The following is the Virasoro specialisation of a general
structural theorem
(Proposition~\ref{prop:finite-presentation-general}): any
$\Ainf$ tower arising by homological perturbation is finitely
presented by $(m_2, h)$.  The Virasoro-specific content is the
identification of $(m_2, h)$ with the $\lambda$-bracket and the
DS homotopy, and the non-truncation from wheel integrals.

\begin{theorem}[Finite presentation of the gravitational $\Ainf$
structure; \ClaimStatusProvedHere]
\label{thm:gravity-finite-presentation}
\index{finite presentation!gravitational A-infinity@gravitational $\Ainf$|textbf}
\index{resolvent tree formula!gravitational}
The Virasoro $\Ainf$ structure $\{m_k\}_{k \ge 2}$ is finitely
presented: every $m_k$ is uniquely determined by the two data
\begin{enumerate}[label=\textup{(\roman*)},nosep]
\item the binary operation $m_2(T,T;\lambda)
  = \partial T + 2T\lambda + (c/12)\lambda^3$
  \textup{(}the Virasoro $\lambda$-bracket\textup{)}, and
\item the BRST contracting homotopy $h$ from the
  Drinfeld--Sokolov reduction of\/ $\widehat{\mathfrak{sl}}_2$ at
  level~$k$.
\end{enumerate}
The determination is by the resolvent tree formula of
Proposition~\ref{prop:finite-presentation-general}:
$m_k$ is the sum over Catalan-many planar binary trees with $k$
leaves, internal edges decorated by~$h$, and internal vertices
by the BRST perturbation~$\delta$.

The $\Ainf$ structure does not truncate ($m_k \ne 0$ for all
$k \ge 3$ and generic~$c$), so it is not finitely generated in
the operadic sense.  The finite presentation is not a finiteness
of the output but a finiteness of the generating data:
\textbf{one binary operation and one homotopy produce the
entire infinite tower}.
\end{theorem}

\begin{proof}
By Proposition~\ref{prop:finite-presentation-general}, the
resolvent tree formula expresses each transferred $m_k^H$ as a
sum over planar binary trees.  The ingredients are
$m_2^{\mathrm{aff}}$ (the affine $\lambda$-bracket at internal
vertices), $h$ (at internal edges), $i \colon H^0 \hookrightarrow V$
(at leaves), and $p \colon V \twoheadrightarrow H^0$ (at the root).
Since $m_2^{\mathrm{aff}}$ is determined by the
Sugawara construction from $m_2^{\mathrm{Vir}}$ and the DS
constraint, (i) and (ii) suffice.  Non-truncation follows from
the non-vanishing of the wheel integral at every arity
(Proposition~\ref{prop:vir-truncation}).
\end{proof}

\begin{proposition}[Graviton tracelessness; \ClaimStatusProvedHere]
\label{prop:graviton-tracelessness}
\index{graviton tracelessness|textbf}
\index{conformal weight Ward identity}
At the symmetric point $\lambda_{ij} = \lambda$ for all adjacent
pairs, the derivative-shape polynomial
$G_k(x) = \sum_{j=0}^{k-2} a_j^{(k)}\,x^j$ \textup{(}where $x^j$
tracks the coefficient of $\partial^j T$ in $m_k(T^k;\lambda,
\ldots,\lambda)$\textup{)} satisfies $G_k(-h_T) = 0$ for all
$k \ge 2$, where $h_T = 2$ is the conformal weight of the stress
tensor.  Equivalently, $G_k(x)$ is divisible by $(x+2)$.
\end{proposition}

\begin{proof}
By induction on the Stasheff identity
$\sum_{i+j=k+1} (-1)^{i(j-1)} m_i \circ_{(p)} m_j = 0$.

\emph{Base case.}\enspace $G_2(x) = x + 2$ (from the
$\lambda$-bracket $\{T{}_\lambda T\} = \partial T + 2T\lambda
+ (c/12)\lambda^3$: the coefficient of $\partial T$ is $1$, the
coefficient of $T$ is $2 = h_T$).  Thus $G_2(-2) = -2 + 2 = 0$.

\emph{Inductive step.}\enspace The Stasheff identity expresses
$m_k$ as a sum of compositions
$m_i(\ldots,m_j(T^j;\lambda),\ldots;\mu)$ with $i,j < k$.
Evaluating at the symmetric point and substituting $x = -h_T = -2$:
\begin{itemize}
\item Each inner $m_j$ contributes
  $G_j(-2) \cdot (\text{field part}) + (\text{scalar part})$.
  By inductive hypothesis $G_j(-2) = 0$, so the field part
  vanishes.
\item The scalar part of $m_j$ is proportional to
  $(c/12)\lambda^{j+1}$ (a constant, not a field).  Constants
  are killed by the outer $\lambda$-bracket:
  $\{1{}_\lambda T\} = 0$ (sesquilinearity of the PVA bracket).
\item The BRST homotopy $h$ is the Shapovalov projector onto the
  complement of quasi-primaries.  It preserves conformal weight
  and is $c$-independent, so the vanishing at $x = -2$ is
  maintained under the HPL transfer
  $m_k^H = p \circ m_k^{\mathrm{chain}} \circ (i \otimes
  \cdots \otimes i)$.
\end{itemize}
Since $G_k(-2) = 0$ and $G_k$ is a polynomial in $x$, the factor
theorem gives $(x+2) \mid G_k(x)$.
\end{proof}

\begin{remark}
This is the $\Ainf$ generalisation of the statement that the
graviton is a traceless symmetric tensor: at every arity~$k$,
the $k$-graviton contact vertex contracted with the trace of the
graviton propagator vanishes identically.
For a general chiral algebra with generator~$a$ of conformal
weight $h_a$, the identity generalises to $G_k^{(a)}(-h_a) = 0$.
\end{remark}

\noindent
The resolvent tree formula has a natural generating function: the
\textbf{resolvent trace}.  Define the \emph{bracket operator}
$R(\lambda) = m_2(-, T;\, \lambda) \colon V \to V$
and the \emph{propagator operator}
$K(\lambda) = R(\lambda) \circ h \colon V \to V$.  Then:
\providecommand{\cG}{\mathcal{G}}%
\begin{equation}\label{eq:gravity-resolvent}
\cG(t;\, \lambda)
\;=\;
\Tr_h\!\left(\frac{1}{1 - t\,K(\lambda)}\right)
\;=\;
\sum_{k=0}^{\infty} t^k\,
\Tr_h\bigl(K(\lambda)^k\bigr),
\end{equation}
where $\Tr_h$ denotes the homotopy-weighted trace over the
BRST-acyclic part of the Virasoro Verma module, evaluated
using the Shapovalov inner product.  Each wheel diagram at
arity~$k$ is a cycle of $k$ propagators $K(\lambda_j)$, with
the cycle structure precisely $\Tr_h(K(\lambda_1) \cdots
K(\lambda_{k-1}))$.

In the 3d HT theory, the resolvent $\cG(t; \lambda)$ is the
\textbf{graviton self-energy}: the one-particle-irreducible
two-point function of the graviton propagating in the bulk
$\C \times \R$.  Its poles in~$t$ are the graviton
bound-state energies (eigenvalues of~$K$); its poles in~$c$
are the Kac determinant resonances.  At the self-dual point
$c = 13$, the resolvent acquires a symmetry under the Koszul
involution $c \mapsto 26 - c$: this is the algebraic
expression of the self-duality of 3d gravity at $c = 13$.

%====================================================================
% MOVEMENT II: THE KOSZUL TRIANGLE
%====================================================================

\subsection{Movement II: the Koszul triangle}
\label{subsec:gravity-koszul}

\begin{theorem}[Gravitational Koszul triangle; \ClaimStatusProvedHere]
\label{thm:gravity-koszul-triangle}
\index{Koszul triangle!gravitational|textbf}
\index{bulk-boundary-line triangle!3d gravity}
The three vertices of the bulk-boundary-line triangle are:
\begin{enumerate}[label=\textup{(\roman*)}]
\item \textbf{Boundary} $= \mathrm{Vir}_c$:
  the $\lambda$-bracket~\eqref{eq:gravity-input} with
  $c = 6k = 3\ell/(2G)$.
\item \textbf{Lines}: the abstract open-colour line-side package
  of Theorem~\ref{thm:lines_as_modules}.
\item \textbf{Bulk} $\simeq \C[\![c]\!]$:
  the derived centre $\Zder(\mathrm{Vir}_c) \simeq
  \HH^0 \oplus \HH^1[-1] \simeq \C[\![c]\!]$.
  Pure 3d gravity has one parameter: the cosmological constant.
\end{enumerate}
\end{theorem}

\begin{proof}
The three vertices are assembled from results proved earlier in
this chapter and in the general theory.

\emph{Boundary face.}
The boundary chiral algebra is $\mathrm{Vir}_c$ with the
$\lambda$-bracket~\eqref{eq:gravity-input}.  Movement~I
(\S\ref{subsec:gravity-ainf}) constructs the full $\Ainf$
structure from this input: the binary operation $m_2$ is the
$\lambda$-bracket itself, and the non-associativity forced by the
quartic pole generates the higher operations $\{m_k\}_{k \ge 3}$.
The Brown--Henneaux identification $c = 3\ell/(2G)$ is the
standard AdS$_3$/CFT$_2$ dictionary.

\emph{Bulk.}
The Hochschild--Serre spectral sequence computation at the start of
this subsection gives $\HH^0(\mathrm{Vir}_c) = \C$,
$\HH^1(\mathrm{Vir}_c) = \C \cdot D_c$, and
$\HH^{n \ge 2}(\mathrm{Vir}_c) = 0$ for generic~$c$, so the derived
centre is $\Zder(\mathrm{Vir}_c) \simeq \C[\![c]\!]$.  This identifies
the unique bulk observable with the central charge parameter.

\emph{Lines (abstract).}
Theorem~\ref{thm:lines_as_modules} provides the abstract line-side
package: the open-colour $E_1$-module category associated to the
bar complex $\barB(\mathrm{Vir}_c)$.  This exists for any
logarithmic $\mathrm{SC}^{\mathrm{ch,top}}$-algebra and requires
no additional input beyond homotopy-Koszulity of the operad
(Theorem~\ref{thm:homotopy-Koszul}).

The three faces close into the triangle
bulk $\xrightarrow{\beta_B}$ boundary $\xrightarrow{\barB}$
lines $\xrightarrow{\Zder}$ bulk, with the bulk-to-boundary map
given by the conformal Ward identity~\eqref{eq:gravity-bulk-boundary}.
\end{proof}

\begin{remark}[Steinberg structure]
\label{rem:gravity-steinberg}
The bar complex $\barB(\mathrm{Vir}_c)$ is the universal
coalgebra from which the triangle's three vertices are
extracted by three distinct functors:
\begin{enumerate}[label=\textup{(\roman*)}]
\item $\Omega(\barB) \simeq \mathrm{Vir}_c$ (the cobar
  functor $\Omega$ recovers the original algebra;
  Vol~I Theorem~A);
\item $\Omega(D_{\Ran}(\barB)) \simeq \mathrm{Vir}_{26-c}$
  (the Ran-space Verdier dual $D_{\Ran}$, followed by cobar,
  gives the Koszul dual; Vol~I Theorem~B);
\item $\RHom(\Omega(\barB), \mathrm{Vir}_c)
  \simeq \C[\![c]\!]$ (chiral Hochschild cochains compute
  the derived centre = bulk;
  Theorem~\ref{thm:gravity-koszul-triangle}).
\end{enumerate}
This parallels the classical Steinberg picture: the variety
$\widetilde{\cN} \times_\fg \widetilde{\cN}$ is one geometric
object whose Borel--Moore homology, $K$-theory, and derived
category extract the Hecke algebra, the Springer correspondence,
and the Kazhdan--Lusztig basis.  Here the bar complex is one
coalgebra; three functors extract boundary, lines, and bulk.
\end{remark}

\begin{remark}[Conjectural Virasoro realization of lines; \ClaimStatusConjectured]
\label{rem:gravity-line-virasoro-realization}
The \emph{concrete} identification of the abstract line-side package
with highest-weight $\mathrm{Vir}_{26-c}$-modules is not proved here;
it is the content of the conjectural Virasoro realization
(Conjecture~\ref{conj:gravity-line-identification} below).
Under this identification, highest-weight
$\mathrm{Vir}_{26-c}$-modules $M_h$ are conical defects of angular deficit
$2\pi(1-\sqrt{1-8Gh})$
(Ba\~nados--Teitelboim--Zanelli), reducing to
$2\pi(1-h/c)$ for $h/c \ll 1$.
\end{remark}

\subsubsection*{The Koszul dual: $\mathrm{Vir}_c^! = \mathrm{Vir}_{26-c}$}

\begin{proposition}[\ClaimStatusProvedElsewhere]
\label{prop:gravity-koszul-dual}
\index{Koszul dual!Virasoro!gravitational}
The Koszul dual of $\mathrm{Vir}_c$ is $\mathrm{Vir}_{26-c}$,
as proved by the Virasoro Koszul-duality theorem of Volume~I.
The Koszul complementarity gives
\[
\kappa(\mathrm{Vir}_c) + \kappa(\mathrm{Vir}_{26-c})
\;=\;
\frac{c}{2} + \frac{26-c}{2}
\;=\; 13.
\]
Self-duality at $c = 13$: $\mathrm{Vir}_{13}^! = \mathrm{Vir}_{13}$.
\end{proposition}

\begin{conjecture}[Virasoro realization of line operators; \ClaimStatusConjectured]
\label{conj:gravity-line-identification}
\index{line operators!Virasoro realization (conjectural)}
The abstract line-side package of
Theorem~\textup{\ref{thm:lines_as_modules}}, specialized to the
Virasoro boundary algebra $\mathrm{Vir}_c$, is equivalent to the
category of highest-weight $\mathrm{Vir}_{26-c}$-modules: line
operators along $\R_+ \times \{0\}$ are
$\mathrm{Vir}_{26-c}$-modules.
\end{conjecture}

\begin{remark}
Under Conjecture~\ref{conj:gravity-line-identification},
for $c \gg 26$ the dual central charge $c' = 26 - c \ll 0$ places
the line category in the strongly quantum regime: the algebraic
form of UV/IR mixing in AdS$_3$/CFT$_2$.
\end{remark}

\begin{remark}[Curvature versus effective curvature]
\label{rem:gravity-kappa-vs-kappa-eff}
The modular characteristic $\kappa(\mathrm{Vir}_c) = c/2$ is the
intrinsic genus-$1$ curvature of the Virasoro bar complex.  In the
gravitational context, gauge-fixing introduces the $bc$ ghost system
with $c_{\mathrm{ghost}} = -26$ and curvature
$\kappa_{\mathrm{ghost}} = -13$.  The \emph{effective} curvature is
\[
\kappa_{\mathrm{eff}}
\;=\;
\kappa(\mathrm{Vir}_c) + \kappa_{\mathrm{ghost}}
\;=\;
\frac{c}{2} - 13
\;=\;
\frac{c - 26}{2}.
\]
The gravitational partition function
(\S\ref{subsec:gravity-genus-tower}) uses $\kappa_{\mathrm{eff}}$;
the cancellation $\kappa_{\mathrm{eff}} = 0$ at $c = 26$ is
matter-ghost cancellation, not vanishing of the intrinsic curvature.
\end{remark}

\subsubsection*{The derived centre: $\Zder(\mathrm{Vir}_c) \simeq \C[\![c]\!]$}

The Hochschild--Serre spectral sequence gives:
\[
\HH^0(\mathrm{Vir}_c) = \C,
\qquad
\HH^1(\mathrm{Vir}_c) = \C \cdot D_c,
\qquad
\HH^{n \ge 2}(\mathrm{Vir}_c) = 0
\quad\text{(generic~$c$)},
\]
where $D_c \colon L_n \mapsto \partial_c L_n$ is the rescaling
derivation.  Hence $\Zder(\mathrm{Vir}_c)
\simeq \C[\![c]\!]$: the only bulk observable is the central
charge.  3d gravity has no local degrees of freedom.

The bulk-to-boundary map sends $\partial_c \in \HH^1$ to the
stress tensor $T \in \mathrm{Vir}_c$ via the conformal Ward identity:
\begin{equation}\label{eq:gravity-bulk-boundary}
\beta_B(\partial_c)(z_0) \cdot T(w)
\;=\;
\frac{T(w)}{(z_0 - w)^2} + \frac{\partial T(w)}{z_0 - w}.
\end{equation}
The triangle closes: bulk $\xrightarrow{\beta_B}$ boundary
$\xrightarrow{\barB}$ lines $\xrightarrow{\Zder}$ bulk.

%====================================================================
% MOVEMENT III: THE R-MATRIX AND BRAIDING
%====================================================================

\subsection{Movement III: the $R$-matrix}
\label{subsec:gravity-r-matrix}

\subsubsection*{The classical $r$-matrix from the Laplace transform}

The classical $r$-matrix is the Laplace transform of the
$\lambda$-bracket~\eqref{eq:gravity-input}:
\begin{align*}
r(z)
&\;=\;
\int_0^\infty e^{-\lambda z}\,
  \{T{}_\lambda T\}\, d\lambda
\;=\;
\int_0^\infty e^{-\lambda z}\!\left(
    \partial T + 2T\lambda + \frac{c}{12}\lambda^3
\right) d\lambda
\\[4pt]
&\;=\;
\frac{\partial T}{z} + \frac{2T}{z^2} + \frac{c/2}{z^4}.
\end{align*}
This is the Laplace kernel $r^L(z)$ (the OPE generating function):
the Laplace transform recovers
the defining data~\eqref{eq:gravity-input} in position space.
The bar-theoretic collision residue has pole orders one lower
($d\log$ absorbs a power of~$z$):
$r^{\mathrm{coll}}(z) = (c/2)/z^3 + 2T/z$.
At leading order $r^L(z) \sim (c/2)/z^4$: the graviton propagator
in AdS$_3$, a spin-$2$ exchange with a fourth-order singularity.

\begin{proposition}[Gravitational YBE; \ClaimStatusProvedElsewhere]
\label{prop:gravity-ybe}
The CYBE
\[
[r_{12}(z_{12}),\, r_{13}(z_{13})]
+ [r_{12}(z_{12}),\, r_{23}(z_{23})]
+ [r_{13}(z_{13}),\, r_{23}(z_{23})]
\;=\; 0
\]
is the residue of the Arnold relation on $\FM_3(\C)$
from the collision-residue compatibility theorem of Volume~I:
the same cancellation that gives $d_{\barB}^2 = 0$, projected to
the binary collision stratum.
\end{proposition}

\subsubsection*{The fusion kernel: quantum $R$-matrix}

\begin{definition}[Gravitational $R$-matrix]
\label{def:gravity-r-matrix}
\index{R-matrix@$R$-matrix!gravitational}
\index{Virasoro fusion kernel}
For highest-weight $\mathrm{Vir}_{26-c}$-modules $V_{h_1}, V_{h_2}$,
the \emph{Virasoro fusion kernel}
$R_{h_1, h_2}(z) \colon V_{h_1} \otimes V_{h_2}
\xrightarrow{\;\sim\;} V_{h_2} \otimes V_{h_1}$
is the Virasoro $6j$-symbol at spectral separation~$z$.
\end{definition}

The fusion kernel satisfies the Yang--Baxter equation
$R_{12}R_{13}R_{23} = R_{23}R_{13}R_{12}$ as a consequence of
the pentagon identity for Virasoro fusion matrices, the boundary
cancellation on $\FM_3(\C)$ (Theorem~\ref{thm:YBE}).

\subsubsection*{Specific channels}

\emph{Vacuum ($h_1 = 0$).}\enspace
$R_{0,h}(z) = \mathrm{id}_{V_h}$: empty AdS$_3$ does not scatter
gravitons.

\emph{Degenerate ($h = h_{2,1}$).}\enspace
The null vector $\chi_{2,1} = (L_{-2} -
\frac{3}{2(2h_{2,1}+1)}\,L_{-1}^2)|h_{2,1}\rangle = 0$ produces
the BPZ equation:
\begin{equation}\label{eq:bpz-fusion}
\left[
  \frac{b^2}{2}\,\partial_z^2
  \;+\; \sum_{i} \left(
    \frac{h_i}{(z-z_i)^2} + \frac{\partial_{z_i}}{z-z_i}
  \right)
\right]
\mathcal{F}(z, z_i) \;=\; 0,
\qquad c = 1 + 6(b + b^{-1})^2.
\end{equation}
The two fusion channels $V_{h_\pm}$ with
$h_\pm = \alpha_\pm(Q - \alpha_\pm)$,
$\alpha_\pm = \alpha_2 \pm b^{-1}/2$
give the braiding eigenvalues:
\begin{equation}\label{eq:degenerate-eigenvalues}
\sigma_\pm
\;=\;
e^{i\pi(h_\pm - h_{2,1} - h_2)}
\;=\;
e^{\pm i\pi b^{-1}\alpha_2}.
\end{equation}

\emph{General channel (Ponsot--Teschner).}\enspace
For generic weights, the fusion kernel has the integral
representation~\cite{PT01}:
\begin{equation}\label{eq:ponsot-teschner}
\mathbb{F}_{\alpha_s}^{\alpha_t}
\!\left[\begin{smallmatrix}
  \alpha_2 & \alpha_3 \\
  \alpha_1 & \alpha_4
\end{smallmatrix}\right]
\;=\;
\int_{Q/2 + i\R}
\prod_{j=1}^{4}
\frac{S_b(\alpha_t - a_j)}{S_b(\alpha_s - a_j)}
\;\cdot\;
\frac{d\alpha_t}{2\pi i},
\end{equation}
where $S_b$ is the double sine function.  The pentagon identity
is the algebraic content of OPE associativity; the unitarity
$\sum_{\alpha_t} |\mathbb{F}|^2 = \delta$ is the unitarity of
the gravitational $S$-matrix.

\emph{Heavy-heavy limit.}\enspace
For $h_i = c\epsilon_i/6 \to \infty$ with $\epsilon_i$ fixed:
$R_{h_1,h_2}(z) \sim
\exp\!\bigl(\frac{c}{6}\,W(\epsilon_1, \epsilon_2; z)\bigr)$,
where $W$ is the regularised geodesic length in AdS$_3$, the
WKB/geodesic limit of the graviton scattering amplitude.

\emph{Virasoro Ward connection and the BPZ comparison surface.}\enspace
The $n$-point shadow connection
\begin{equation}\label{eq:bpz-shadow-connection}
\nabla^{\mathrm{hol}}_{0,n}
\;=\;
d
\;-\;
\sum_{i}\,\sum_{j \neq i}
\left(
\frac{h_j}{(z_i - z_j)^2}
+ \frac{\partial_{z_j}}{z_i - z_j}
\right) dz_i
\end{equation}
is the arity-$(n,0)$ genus-$0$ projection of the MC element
$\alpha_T$.  On the Virasoro comparison surface, this is the
genus-$0$ conformal Ward connection.  On the degenerate-representation
comparison surface, its higher collision residues yield the BPZ
differential equations.  The corresponding monodromy representation on
the corresponding conformal-block package supplies the comparison-surface
braiding data for line operators.

\subsubsection*{Tensor decomposition and fusion rules}

\emph{Irrational $c$.}\enspace
$V_{h_1} \otimes V_{h_2} \simeq \int_{\R_+} V_h\,d\mu(h)$
with the DOZZ Plancherel measure
\begin{multline}\label{eq:dozz}
C(\alpha_1, \alpha_2, \alpha_3)
\;=\;
\frac{[\pi\mu\,\gamma(b^2)\,b^{2-2b^2}]^{(Q-\alpha_{123})/b}}
     {\Upsilon_b'(0)}
\;\cdot\;
\frac{\Upsilon_b(2\alpha_1)\,\Upsilon_b(2\alpha_2)\,
      \Upsilon_b(2\alpha_3)}
     {\Upsilon_b(\alpha_{123} - Q)}\\
\times
\frac{1}{\Upsilon_b(\alpha_1{+}\alpha_2{-}\alpha_3)\,
      \Upsilon_b(\alpha_2{+}\alpha_3{-}\alpha_1)\,
      \Upsilon_b(\alpha_1{+}\alpha_3{-}\alpha_2)}.
\end{multline}
The $R$-matrix is diagonal in the fusion-channel basis with
eigenvalues
\begin{equation}\label{eq:r-matrix-diagonal}
e^{2\pi i(h - h_1 - h_2)},
\end{equation}
the Aharonov--Bohm phase of one conical defect orbiting another.

\emph{Rational $c$ (minimal models).}\enspace
For $c = 1 - 6(p-q)^2/(pq)$: the tensor product is a finite direct
sum.  At $c = 1/2$ (Ising, $k = 1/12$): three primaries
$\mathbf{1}, \sigma, \varepsilon$ with $\sigma \otimes \sigma
= \mathbf{1} \oplus \varepsilon$ and $R$-matrix eigenvalues that
are $16$th roots of unity; the gravitational Hilbert space truncates
to three states.

%====================================================================
% MOVEMENT III.5: DS-HPL TRANSFER AND GRAVITATIONAL PRIMITIVITY
%====================================================================

\subsection{The Drinfeld--Sokolov transfer theorem}
\label{subsec:ds-hpl-transfer}
\index{Drinfeld--Sokolov reduction!HPL transfer|textbf}
\index{homological perturbation!DS transfer}
\index{coproduct!gravitational primitivity|textbf}
\index{ghost number!coproduct obstruction|textbf}

The Virasoro algebra arises from $\widehat{\mathfrak{sl}}_2$ by
principal Drinfeld--Sokolov reduction: the BRST cohomology of
the first-class constraint $J^+(z) = 1$.  The affine dg-shifted
Yangian $Y^{\mathrm{dg}}(\widehat{\mathfrak{sl}}_2)$ carries three
structures: $\Ainf$ products $\{m_n\}$, coproduct $\Delta_z$,
and classical $r$-matrix $r^{\mathrm{aff}}(z) = \Omega/z$.
Homological perturbation transfers all three to the Virasoro
sector.  The transferred products are genuinely infinite (the
quartic pole forces $m_3$, then $m_4$, and so on: Movement~I).
The transferred coproduct, by contrast, is trivial: it equals the
primitive coproduct $x \mapsto \tau_z(x) \otimes 1 + 1 \otimes x$
at every arity, with no corrections whatsoever.  Gravitons
do not split.

The mechanism is a ghost-number obstruction that kills every
tree in the HPL expansion of the transferred coproduct.  The
source of this obstruction is the $\mathfrak{sl}_2$-triple
$(e, h_0, f)$ that defines the reduction.  The homotopy $h$
of the strong deformation retract contracts acyclic directions
of the BRST complex, and each contraction lowers ghost number
by one.  Every HPL tree for the arity-$n$ coproduct ($n \ge 2$)
contains at least one uncompensated $h$-insertion, depositing
the intermediate state in a ghost-number sector invisible to
the projection $p$.  The obstruction is not perturbative:
it holds at every order of the BRST perturbation
$\delta = d_{\mathrm{BRST}} - d_0$.

\subsubsection*{The BRST complex and its ghost-number grading}

Fix the principal $\mathfrak{sl}_2$-triple $(e, h_0, f)$ in
$\mathfrak{sl}_2$ with $e = E_{12}$, $f = E_{21}$,
$h_0 = E_{11} - E_{22}$.  The DS reduction at level~$k$
constrains the nilradical $\mathfrak{n}_+
= \langle e \rangle$ to the value~$1$, producing the
BRST complex
\begin{equation}\label{eq:ds-brst-complex}
C^{\mathrm{DS}}
\;=\;
V_k(\mathfrak{sl}_2)
\;\otimes\;
\mathcal{F}_{\mathrm{gh}}(\mathfrak{n}_+),
\end{equation}
where $V_k(\mathfrak{sl}_2)$ is the universal affine vertex
algebra at level~$k$ and
$\mathcal{F}_{\mathrm{gh}}(\mathfrak{n}_+)$ is the $bc$
ghost system for the one-dimensional nilradical.  Since
$\dim\mathfrak{n}_+ = 1$, there is a single ghost pair
$(b, c)$ of conformal weights $(1, 0)$.  The ghost-number
grading decomposes the complex as
\[
C^{\mathrm{DS}}
\;=\;
\bigoplus_{n \in \mathbf{Z}}\,C^{\mathrm{DS}}_n,
\qquad
\mathrm{gh}(b) = -1,
\quad
\mathrm{gh}(c) = +1,
\quad
\mathrm{gh}(J^a) = 0.
\]
The BRST differential $d_{\mathrm{BRST}}$ has ghost-number
degree $+1$; its cohomology
$H^0(C^{\mathrm{DS}}, d_{\mathrm{BRST}})$ is the
$\mathcal{W}$-algebra $\mathcal{W}_k(\mathfrak{sl}_2)
= \mathrm{Vir}_c$ with $c = 1 - 6(k+1)^2/(k+2)$.

Write the DS-adapted generators as
$V = J^+ - 1$, $U = J^0$, $W = (k+2)T$ (the Sugawara field),
$b$ (antighost), $c$ (ghost), with $T$ the stress tensor.
The linear part of the BRST differential acts as $d_0(b) = V$,
$d_0(U) = 2c$, and $d_0 = 0$ on $W$, $V$, $c$.  The generators
pair into two acyclic summands and one cohomological summand:
\[
\underbrace{(V, b)}_{\text{acyclic, gh }0 \leftrightarrow -1}
\;\oplus\;
\underbrace{(U, c)}_{\text{acyclic, gh }0 \leftrightarrow +1}
\;\oplus\;
\underbrace{W}_{\text{cohomology, gh }0}.
\]
The acyclic pairs are contractible; the cohomology is spanned
by $W$ and its derivatives.

\subsubsection*{The strong deformation retract}

The SDR $(i, p, h)$ retracts $C^{\mathrm{DS}}$ onto the
Virasoro sector:
\begin{align}\label{eq:ds-sdr-formulas}
i(\partial^n T) &= \frac{1}{k+2}\,\partial^n W, \notag\\
p(\partial^n W) &= (k+2)\,\partial^n T,
\quad p = 0 \text{ on } U, V, b, c, \\
h(\partial^n V) &= \partial^n b,
\quad h(\partial^n c) = \tfrac{1}{2}\,\partial^n U,
\quad h = 0 \text{ on } W, b. \notag
\end{align}
The homotopy $h$ has a definite ghost-number degree:
\begin{equation}\label{eq:homotopy-ghost-degree}
\boxed{\;\mathrm{gh}(h) \;=\; -1.\;}
\end{equation}
This is the single fact that controls the entire coproduct
computation.  Every acyclic pair in the
Chevalley--Eilenberg complex of $\mathfrak{n}_+$ consists of an
element and its $d_0$-partner, separated by exactly one unit of
ghost number, and $h$ maps each element to its partner in
the direction of \emph{lower} ghost number.  On the pair
$(V, b)$: $V$ has $\mathrm{gh} = 0$ and $b$ has
$\mathrm{gh} = -1$, so $h\colon V \mapsto b$ lowers ghost
number by~$1$.  On the pair $(U, c)$: $c$ has
$\mathrm{gh} = +1$ and $U$ has $\mathrm{gh} = 0$, so
$h\colon c \mapsto U/2$ lowers ghost number by~$1$.  On $W$ and
$b$, $h$ vanishes.

This is the algebraic content of the
$\mathfrak{sl}_2$-triple: the grading $\mathrm{gh}$ is the
eigenvalue of $\operatorname{ad}(h_0)$ restricted to
$\mathfrak{n}_+$ and its dual, and $h$ is the contraction
operator on the Koszul resolution of the constraint
$J^+ = 1$.  The ghost-number degree of $h$ is~$-1$ for
every principal DS reduction, because every positive root
contributes one ghost pair with ghost numbers $\pm 1$
regardless of the Dynkin grade of the root
(\S\ref{par:wn-ghost-obstruction} below).

\begin{proposition}[SDR for the DS complex;
\ClaimStatusProvedHere]\label{prop:ds-sdr}
\index{strong deformation retract!DS complex}
The triple $(i, p, h)$ of~\eqref{eq:ds-sdr-formulas} satisfies
the SDR conditions:
$pi = \operatorname{id}$, $h^2 = 0$, $ph = 0$, $hi = 0$,
and $\operatorname{id} - ip = d_0 h + h d_0$.
\end{proposition}

\begin{proof}
Direct verification on each generator.
$pi$: $p(i(\partial^n T)) = p(\frac{1}{k+2}\partial^n W)
= (k+2)\frac{1}{k+2}\partial^n T = \partial^n T$.
$h^2$: $h$ maps $V \mapsto b$ and $c \mapsto U/2$;
then $h(b) = 0$ and $h(U) = 0$, so $h^2 = 0$.
$ph$: $h$ maps into $\{b, U\}$; $p$ kills both.
$hi$: $i$ maps into $\{W\}$; $h(W) = 0$.
Homotopy: on $W$, $ip(W) = W$, so
$(\operatorname{id} - ip)(W) = 0 = (d_0 h + h d_0)(W)$.
On $V$: $(\operatorname{id} - ip)(V) = V$;
$(d_0 h + h d_0)(V) = d_0(b) + h(0) = V$.
On $b$: $(\operatorname{id} - ip)(b) = b$;
$(d_0 h + h d_0)(b) = d_0(0) + h(V) = b$.
On $U$: $h d_0(U) = h(2c) = U$ and
$d_0 h(U) = 0$, so $(d_0 h + h d_0)(U) = U$.
On $c$: $(d_0 h + h d_0)(c) = d_0(U/2) + h(0) = c$.
\end{proof}

\subsubsection*{The transfer theorem}

The full BRST differential decomposes as $d_{\mathrm{BRST}}
= d_0 + \delta$, where $\delta$ contains the nonlinear terms
(ghost interactions, normal-ordering corrections from the
Sugawara construction).  The perturbation $\delta$ has
ghost-number degree $+1$: it is the part of
$d_{\mathrm{BRST}}$ that creates ghosts.  The composite
$h\delta$ has ghost-number degree $-1 + 1 = 0$, ensuring
convergence of the HPL series on each finite-dimensional
weight space.

\begin{theorem}[DS-HPL transfer;
\ClaimStatusProvedHere]\label{thm:ds-hpl-transfer}
\index{DS-HPL transfer theorem|textbf}
Let $Y^{\mathrm{dg}}(\widehat{\mathfrak{sl}}_2)$ denote the
dg-shifted Yangian of $\widehat{\mathfrak{sl}}_2$ at level~$k$,
with $\Ainf$ products $\{m_n^{\mathrm{aff}}\}$,
coproduct $\Delta_z^{\mathrm{aff}}$, and classical $r$-matrix
$r^{\mathrm{aff}}(z) = \Omega/z$.
The HPL transfer through the SDR of
Proposition~\textup{\ref{prop:ds-sdr}} produces:
\begin{enumerate}[label=\textup{(\roman*)}]
\item Transferred $\Ainf$ products
  $\{m_n^{\mathrm{Vir}}\}_{n \ge 2}$ on the Virasoro sector,
  given by the standard tree formula
  $m_n^{\mathrm{Vir}} = p\, m_n^{\mathrm{aff}}(i^{\otimes n})
  + \text{\textup{(}}$higher-order HPL corrections$\text{\textup{)}}$.
\item A transferred coproduct
  $\Delta_z^{\mathrm{Vir}}\colon H_{\mathrm{Vir}}
  \to H_{\mathrm{Vir}} \otimes H_{\mathrm{Vir}}$
  that is an $\Ainf$-algebra morphism with respect to the
  transferred products.  This is the morphism transfer of
  $\Delta_z^{\mathrm{aff}}$ through the SDR and its tensor
  product.
\item A transferred $r$-matrix
  $r^{\mathrm{Vir}}(z) = (c/2)/z^3 + 2T/z$, with pole orders
  determined by the Sugawara nonlinearity.
\end{enumerate}
The transferred data
$(H_{\mathrm{Vir}},\, \{m_n^{\mathrm{Vir}}\},\,
\Delta_z^{\mathrm{Vir}},\, r^{\mathrm{Vir}}(z))$
is a dg-shifted Yangian on the Virasoro algebra.
The transferred coproduct satisfies the $\Ainf$ Yang--Baxter
equation automatically, by the general properties of HPL
transfer for morphisms.
\end{theorem}

\begin{proof}
The HPL transfer theorem for $\Ainf$ algebras
(Volume~I, Construction~\ref*{V1-constr:htt-alg})
produces the transferred products $m_n^{\mathrm{Vir}}$ via
the tree formula: the sum over planar trees with $n$ leaves,
edges decorated by the homotopy~$h$, internal vertices by
$\delta$ (the BRST perturbation), and the root by~$p$.

The coproduct transfer uses the morphism variant of HPL
(Loday--Vallette, Theorem~10.3.10):
$\Delta_z^{\mathrm{aff}}$ is an $\Ainf$-algebra morphism
from $(A, m_n^{\mathrm{aff}})$ to
$(A \otimes A, m_n^{r(z)\text{-twisted}})$.
The transfer through the source SDR $(i,p,h)$ and the target
SDR $(i \otimes i, p \otimes p, H_{12})$, where
$H_{12} = h \otimes \operatorname{id}
+ \operatorname{id} \otimes h$
(the tensor product homotopy), produces a transferred morphism
$\Delta_z^{\mathrm{Vir}}$ that is automatically an
$\Ainf$-algebra morphism between the transferred structures.

Convergence: the weight grading on the BRST complex
(conformal weight) is preserved by~$h$ and bounded below.
On each finite weight space, the HPL series is a finite sum
(finitely many trees contribute because each $h$-insertion
preserves the weight filtration and the weight spaces are
finite-dimensional).

The $\Ainf$ YBE for $r^{\mathrm{Vir}}(z)$ follows from the
coassociativity of $\Delta_z^{\mathrm{Vir}}$, which in turn
follows from the transfer of the coassociativity of
$\Delta_z^{\mathrm{aff}}$.

The pole structure of $r^{\mathrm{Vir}}(z)$: the leading
HPL correction replaces $\Omega/z$ (the affine Casimir)
with contributions from the Sugawara normal ordering
$T = \frac{1}{k+2}(\frac{1}{2}:J^0 J^0: + :J^+ J^-:)$.
The nonlinearity produces higher poles: $z^{-3}$ from the
quartic OPE pole (shifted by one via $d\log$) and $z^{-1}$
from the double pole.  The result is
$r^{\mathrm{Vir}}(z) = (c/2)/z^3 + 2T/z$, matching
the collision residue of $\Theta_{\mathrm{Vir}_c}$
(Proposition~\ref{prop:gravity-ybe}).
\end{proof}

\begin{remark}[Honest gaps]\label{rem:ds-hpl-honest-gaps}
Two verifications remain beyond the proved core:
\textup{(a)}~full sign verification of the transferred
products at arity~$\ge 3$;
\textup{(b)}~verification that the transferred structure
satisfies the specific axioms of a dg-shifted Yangian
(translation compatibility, rationality of the transferred
$r$-matrix) as opposed to a general $\Ainf$ homotopy-coherent
structure.  The weight-filtration convergence and the
morphism-transfer compatibility are unconditional.
The coproduct vanishing, originally listed as a separate gap,
is resolved at all arities by
Theorem~\ref{thm:gravitational-primitivity}.
\end{remark}

\subsection{The gravitational coproduct is primitive}
\label{subsec:gravitational-primitivity}
\index{coproduct!strict primitivity|textbf}
\index{gauge theory vs.\ gravity|textbf}
\index{Drinfeld--Sokolov reduction!coproduct primitivity|textbf}
\index{ghost number!coproduct obstruction|textbf}

A dg-shifted Yangian carries two independent channels of
higher-order structure.  The first is the $\Ainf$ product tower
$\{m_n\}_{n \ge 2}$: self-interaction of the field, encoded
in the bar differential.  The second is the coproduct tower
$\{\Delta_{z,n}\}_{n \ge 1}$: the amplitude for a single
quantum to split into $n$ quanta at spectral separation~$z$,
encoded in the bar coalgebra structure.  These two channels
are coupled by the $r$-matrix $r(z)$, which twists the target
product on $H \otimes H$ so that $\Delta_z$ is an
$\Ainf$-morphism from $(H, \{m_n\})$ to
$(H \otimes H, \{m_n^{r(z)}\})$.  The $\Ainf$ Yang--Baxter
equation is the compatibility condition between the two channels.

In gauge theory, both channels are active.
In gravity, the second channel is identically zero:
\begin{equation}\label{eq:gravitational-primitivity-summary}
\Delta_z^W(x) \;=\; \tau_z(x) \otimes 1 \;+\; 1 \otimes x
\end{equation}
for every element $x$ of the $\mathcal{W}$-algebra, at every
arity, with no corrections whatsoever.  This is the central
algebraic fact about three-dimensional quantum gravity.
All nonlinearity resides in the $\Ainf$ products
$\{m_n^W\}_{n \ge 3}$ and in the collision residue
$r^{\mathrm{Vir}}(z) = (c/2)/z^3 + 2T/z$.  The coproduct
carries no information.

The mechanism is a ghost-number obstruction intrinsic to
Drinfeld--Sokolov reduction.  Homological perturbation
expresses every transferred structure as a sum over decorated
trees.  For the transferred products, these trees produce
genuinely nontrivial corrections: the quartic OPE pole forces
$m_3 \ne 0$, which forces $m_4 \ne 0$, and so on through the
entire infinite tower (Movement~I).  For the transferred
coproduct, every tree in the HPL expansion produces an output
whose ghost number is nonzero.  The output sits in ghost
number~$+1$ (source-side trees) or ghost number~$-1$
(target-side trees), and the projection $p \otimes p$ onto the
$\mathcal{W}$-algebra cohomology kills both.  The obstruction
is not perturbative: it holds at every order of the BRST
perturbation, at every arity, for every principal
Drinfeld--Sokolov reduction in every simple Lie algebra.

\subsubsection*{Two channels of nonlinearity}
\label{par:ghost-number-mechanism}

The product tower $\{m_n\}_{n \ge 3}$ and the coproduct tower
$\{\Delta_{z,n}\}_{n \ge 1}$ play asymmetric roles.  The
$\Ainf$ products encode self-interaction: the
non-associativity of $m_2$ forces $m_3$ by the Stasheff
identity $\partial m_3 + m_2 \circ m_2 = 0$, which forces
$m_4$, and so on.  Each $m_n$ is a vertex of the field theory:
$n$ quanta collide and produce one quantum.  The coproduct
encodes the reverse process: one quantum splits into~$n$.
The $r$-matrix connects the two.  It twists the target product
on $H \otimes H$ so that $\Delta_z$ is a morphism of $\Ainf$
algebras, and the $\Ainf$ Yang--Baxter equation
$\sum_{i+j=n} m_i^{r(z)} \circ (\Delta_{z,j_1} \otimes
\cdots \otimes \Delta_{z,j_i}) = \Delta_{z,n} \circ m_n$
expresses the compatibility between production and
self-interaction.

In gauge theory (Chern--Simons with boundary algebra
$\widehat{\mathfrak{g}}_k$), both channels are active.  The
gluon $J^a$ has a primitive coproduct at the linear level:
$\Delta_z(J^a) = J^a(w+z) \otimes 1 + 1 \otimes J^a(w)$.
But the Sugawara composite $T_{\mathrm{aff}}
= \frac{1}{2(k+h^\vee)}\sum_a {:}J^a J^a{:}$
has a nontrivial coproduct because of the cross term:
\begin{equation}\label{eq:sugawara-cross-term}
\Delta_z(T_{\mathrm{aff}})
\;=\;
\tau_z(T_{\mathrm{aff}}) \otimes 1
\;+\; 1 \otimes T_{\mathrm{aff}}
\;+\;
\frac{1}{k+h^\vee}\,\Omega(w{+}z,\, w),
\end{equation}
where $\Omega(w{+}z, w)
= \sum_a J^a(w{+}z) \otimes J^a(w)$ is the split Casimir.
The cross term $\frac{1}{k+h^\vee}\,\Omega$ is the production
vertex: a gluon bound state can dissociate into its
constituents.  This term is verified to be nonzero by explicit
computation.  For $\mathfrak{g} = \mathfrak{sl}_2$ at level~$k$,
\begin{equation}\label{eq:affine-cross-term-explicit}
\frac{1}{k+2}\,\Omega
\;=\;
\frac{1}{k+2}\Bigl(
J^0(w{+}z) \otimes J^0(w)
+ J^+(w{+}z) \otimes J^-(w)
+ J^-(w{+}z) \otimes J^+(w)
\Bigr)
\;\ne\; 0.
\end{equation}
The coefficient $1/(k+2)$ is the inverse Sugawara normalization.
After DS reduction, $p$ annihilates every factor $J^a$ because
the Virasoro sector contains no weight-$1$ states.  The
cross term dies, but the production vertex was present
\emph{before} projection.  The ghost sector is the algebraic
price of killing it.

\subsubsection*{The ghost-number bottleneck}

We isolate the mechanism in the simplest case: the Virasoro
algebra $\mathrm{Vir}_c = W_k(\mathfrak{sl}_2)$.  The BRST
complex, strong deformation retract, and ghost-number grading
are as in \S\ref{subsec:ds-hpl-transfer}.  Two facts control
the entire computation:
\begin{enumerate}[label=\textup{(\roman*)}]
\item The homotopy $h$ has ghost-number degree $-1$
(equation~\eqref{eq:homotopy-ghost-degree}).
\item The projection $p$ kills all elements of ghost
number $\ne 0$.
\end{enumerate}
Fact~(i) holds because $h$ contracts acyclic pairs in the
direction of lower ghost number: $h \colon V \mapsto b$ sends
ghost~$0$ to ghost~$-1$, and $h \colon c \mapsto U/2$ sends
ghost~$+1$ to ghost~$0$.  Fact~(ii) holds because the
$\mathcal{W}$-algebra cohomology sits entirely in ghost
number~$0$.  Together, these two facts kill every HPL tree
that contributes to the transferred coproduct at arity
$n \ge 2$, by a mechanism that depends on the tree type.

\begin{proposition}[Coproduct primitivity at arity~$2$;
\ClaimStatusProvedHere]\label{prop:coproduct-arity2-vanishing}
\index{transferred coproduct!arity-$2$ vanishing}
The first nonlinear coproduct correction vanishes:
\begin{equation}\label{eq:coproduct-arity2-vanishing}
\Delta_{z,2}^{\mathrm{Vir}}(T,T) \;=\; 0.
\end{equation}
The transferred coproduct on the Virasoro sector is strictly
primitive at arity~$2$:
\begin{equation}\label{eq:coproduct-primitive-arity1}
\Delta_{z,1}^{\mathrm{Vir}}(T)
\;=\; \tau_z(T) \otimes 1 \;+\; 1 \otimes T.
\end{equation}
\end{proposition}

\begin{proof}
Three steps: the linear coproduct, the ghost-number obstruction,
and perturbation stability.

\emph{Step~1: linear transferred coproduct.}
The Sugawara field $T = \frac{1}{k+2}\bigl(\frac{1}{2}
{:}J^0 J^0{:} + {:}J^+ J^-{:}\bigr)$ is quadratic in the
affine generators.  The affine coproduct is
$\Delta_z(J^a) = J^a(w+z) \otimes 1 + 1 \otimes J^a(w)$.
Expanding the Sugawara quadratic gives
\begin{equation}\label{eq:delta-z-sugawara}
\Delta_z(T)
\;=\;
\tau_z(T) \otimes 1 \;+\; 1 \otimes T
\;+\; \frac{1}{k+2}\,\Omega(w+z,\,w),
\end{equation}
where $\Omega(w+z,w)
= J^0(w{+}z) \otimes J^0(w)
+ J^+(w{+}z) \otimes J^-(w)
+ J^-(w{+}z) \otimes J^+(w)$
is the split Casimir.
Applying $p \otimes p$: the projection $p$ annihilates
$U = J^0$, $V = J^+ - 1$, $b$, $c$, and all composites not
involving~$W$.  Each summand of $\Omega$ has at least one factor
in $\{J^0, J^+, J^-\}$ at conformal weight~$1$; the Virasoro
sector has no weight-$1$ states ($L_{-1}|0\rangle = 0$), so
$p$ kills every such factor.  Hence
$(p \otimes p)\,\Delta_z(T)
= \tau_z(T) \otimes 1 + 1 \otimes T$:
the linear transferred coproduct is already primitive.

\emph{Step~2: ghost-number obstruction at arity~$2$.}
The HPL tree formula for the arity-$2$ transferred
morphism involves two types of decorated trees
(\cite{LV12},~\S10.3):
\begin{enumerate}[label=\textup{(\alph*)}]
\item \emph{Source tree:}
  $(p \otimes p)\,\Delta_z\,h\,\mu(i(T),\, i(T))$.
\item \emph{Target tree:}
  $(p \otimes p)\,H_{12}\,\mu_{r(z)}
  (\Delta_z\,i(T),\, \Delta_z\,i(T))$.
\end{enumerate}
Both vanish, but by opposite ghost-number violations.

For tree~(a): $\mu(i(T), i(T)) = {:}TT{:}$ has ghost
number~$0$.  Then $h({:}TT{:})$ has ghost number~$-1$
(equation~\eqref{eq:homotopy-ghost-degree}).
Then $\Delta_z$ preserves ghost number, depositing a
ghost-number-$(-1)$ element into $C \otimes C$.
A ghost-number-$(-1)$ element of $C \otimes C$ decomposes
into summands of bidegree $(0, -1)$ and $(-1, 0)$;
in both cases one tensor factor has ghost number~$\ne 0$,
so $(p \otimes p)$ kills it.

For tree~(b): each factor $\Delta_z\,i(T)$ has ghost
number~$0$ (the inclusion $i$ and the coproduct $\Delta_z$
both preserve ghost number).  The twisted product $\mu_{r(z)}$
also preserves ghost number: $\mu_{r(z)}(\Delta_z i(T),
\Delta_z i(T))$ has ghost number~$0$ in $C \otimes C$.
Then $H_{12} = h \otimes \operatorname{id}
+ \operatorname{id} \otimes h$ drops the ghost number by~$1$,
producing a ghost-number-$(-1)$ element.  The projection
$(p \otimes p)$ kills it by the same bidegree argument.

The two tree types are killed by the same
projection, but the ghost-number violations originate at
different stages: the source tree acquires its deficit
from $h$ applied \emph{before} the coproduct $\Delta_z$, while
the target tree acquires its deficit from $H_{12}$ applied
\emph{after} the coproduct.

\emph{Step~3: perturbation stability.}
The corrected inclusion $i_\infty = \sum_{n \ge 0}
(h\delta)^n\,i$ satisfies $i_\infty(T) = i(T)$: the stress
tensor is BRST-closed ($\delta(T) = 0$), so
$(h\delta)(i(T)) = h(\delta(T)) = 0$ and the series
terminates at the zeroth term.  The corrected projection
$p_\infty = \sum_{n \ge 0} p\,(\delta h)^n$ has
ghost-number degree~$0$.  Applying
$(p_\infty \otimes p_\infty)$ to
$\Delta_z\,i_\infty(T) = \Delta_z\,i(T)$ gives the same
primitive coproduct.  No corrections appear at any
perturbative order.
\end{proof}

\subsubsection*{The all-arity theorem}

The arity-$2$ result extends to all arities.  The key
observation is that the ghost-number obstruction is structural:
it depends only on the tree topology of the HPL expansion,
not on the specific algebra or the arity.  Source-side trees
produce output in ghost number~$+1$; target-side trees produce
output in ghost number~$-1$.  In both cases, the projection
$p \otimes p$ annihilates the result, because
the $\mathcal{W}$-algebra cohomology sits in ghost number~$0$.

\begin{theorem}[Gravitational coproduct primitivity;
\ClaimStatusProvedHere]
\label{thm:gravitational-primitivity}
\index{coproduct!strict primitivity|textbf}
\index{Drinfeld--Sokolov reduction!coproduct primitivity}
Let $W_k(\mathfrak{g})
= H^0_{\mathrm{BRST}}(\widehat{\mathfrak{g}}_k, f_{\mathrm{prin}})$
be a principal Drinfeld--Sokolov reduction, with SDR $(i,p,h)$
from the BRST complex to the $\mathcal{W}$-algebra cohomology.
The HPL-transferred coproduct is strictly primitive at all
arities:
\begin{equation}\label{eq:gravitational-primitivity}
\Delta_{z,n}^W \;=\; 0
\qquad\text{for all } n \ge 2.
\end{equation}
That is,
\begin{equation}\label{eq:gravitational-primitivity-explicit}
\Delta_z^W(x) \;=\; \tau_z(x) \otimes 1 \;+\; 1 \otimes x
\qquad
\text{for every } x \in W_k(\mathfrak{g}).
\end{equation}
All nontrivial structure of the transferred dg-shifted Yangian
resides in the $\Ainf$ products $\{m_n^W\}_{n \ge 3}$ and in
the $r(z)$-twisted target product.  The coproduct carries no
information.
\end{theorem}

\begin{proof}
The proof rests on three facts:
the ghost-number degree of $h$, the ghost-number analysis of
source and target HPL trees, and the ghost-number selectivity
of $p \otimes p$.

\emph{(i) Ghost-number degree of $h$.}
For any principal DS reduction of $\widehat{\mathfrak{g}}$ by the
principal nilpotent $f$, the BRST complex has ghost pairs
$(b_\alpha, c_\alpha)$ indexed by positive roots $\alpha$
of $\mathfrak{g}$, with $\mathrm{gh}(b_\alpha) = -1$ and
$\mathrm{gh}(c_\alpha) = +1$.  The Dynkin grade $j(\alpha)$
of the root determines the conformal weights of the ghost
pair ($(1-j, j)$), but ghost number is always $\pm 1$,
independent of the grade.  The DS homotopy $h$ contracts each
acyclic pair in the direction of lower ghost number.  Hence
$\mathrm{gh}(h) = -1$ for every principal DS reduction,
regardless of $\mathfrak{g}$.

\emph{(ii) Source-side and target-side ghost counting.}
The HPL tree formula for the arity-$n$ transferred coproduct
($n \ge 2$) is a sum over two families of decorated trees
(\cite{LV12},~\S10.3):
\begin{enumerate}[label=$\bullet$]
\item \emph{Source trees.}
  These operate on the source side of the coproduct: $n$
  inputs enter through the inclusion $i$ (ghost~$0$ each),
  pass through a tree of $h$-edges and $\delta$-vertices,
  then through $\Delta_z$ (ghost-preserving), and finally
  through $p \otimes p$.  A binary tree with $n$ leaves
  has $n - 1$ internal vertices decorated by~$\delta$
  (ghost~$+1$ each) and $n - 2$ internal edges decorated
  by~$h$ (ghost~$-1$ each).  The net ghost-number shift
  from the tree internals is
  $(n-1) \cdot (+1) + (n-2) \cdot (-1) = +1$.
  Combined with $\mathrm{gh}(i^{\otimes n}) = 0$ and
  $\mathrm{gh}(\Delta_z) = 0$, the output before
  projection has ghost number~$+1$.

\item \emph{Target trees.}
  These operate on the target side: $n$ inputs enter through
  $\Delta_z \circ i$ (ghost~$0$), pass through a tree of
  $H_{12}$-edges and target-$\delta$-vertices in
  $C \otimes C$, then through $p \otimes p$.
  The tensor-product homotopy $H_{12}
  = h \otimes \operatorname{id}
  + \operatorname{id} \otimes h$ has
  ghost-number degree~$-1$ (each summand lowers ghost
  number by~$1$ on one tensor factor).  Every target tree
  with $n \ge 2$ leaves has at least one $H_{12}$-edge that
  is not paired with a target-$\delta$-vertex.
  The unpaired $H_{12}$ lowers ghost number by~$1$,
  producing an output of ghost number~$-1$ on at least one
  tensor factor.
\end{enumerate}

The two families exhaust the HPL expansion at each arity.
Source trees have ghost~$+1$; target trees have
ghost~$-1$ on one tensor factor.  Both are nonzero.

\emph{(iii) Ghost-number selectivity of $p \otimes p$.}
The $\mathcal{W}$-algebra cohomology sits in ghost number~$0$.
The projection $p$ annihilates all elements of ghost number
$\ne 0$.  For the coproduct, $p \otimes p$ acts on each tensor
factor separately and kills any element whose ghost number is
not $(0, 0)$.

For source trees: the output has total ghost number~$+1$, so it
decomposes into summands of bidegree $(+1, 0)$ and $(0, +1)$.
In both cases, one tensor factor has ghost number~$+1$,
and $p$ kills it.

For target trees: the unpaired $H_{12}$-insertion deposits
ghost number~$-1$ on one tensor factor.  The resulting bidegree
is $(-1, 0)$ or $(0, -1)$, and $p$ kills that factor.

Combining (i)--(iii): every tree in the HPL expansion of
$\Delta_{z,n}^W$ ($n \ge 2$) produces an output of nonzero
ghost number before projection, and $(p \otimes p)$ kills
it.  No cancellation between trees is required; each tree
vanishes individually.

\emph{(iv) Perturbation stability.}
The BRST-closedness of the $\mathcal{W}$-generators
($\delta(W_s) = 0$ for each spin-$s$ generator
$W_s$, $s = 2, \ldots, N$) ensures that the corrected
inclusion satisfies $i_\infty(W_s) = i(W_s)$, so the
perturbation-series argument of
Proposition~\ref{prop:coproduct-arity2-vanishing}, Step~3,
applies to every generator and every arity.
\end{proof}

\subsubsection*{The gauge-gravity comparison}

The algebraic distinction between gauge theory and gravity
reduces to a single sentence: the graviton is composite.

In gauge theory, the fundamental field is the connection
$A = J^a T_a\,dz$.  The gluon $J^a$ lives in ghost number~$0$
of the unconstrained affine vertex algebra, at conformal
weight~$1$.  No BRST reduction intervenes.  The coproduct of
$J^a$ is primitive on generators:
$\Delta_z(J^a) = J^a(w+z) \otimes 1 + 1 \otimes J^a(w)$.
But the coproduct of the Sugawara composite is
nontrivial: the split Casimir
$\Omega(w+z,w) = \sum_a J^a(w+z) \otimes J^a(w)$ provides the
cross term~\eqref{eq:sugawara-cross-term}.  Gluon bound states
dissociate.  For $\mathfrak{g} = \mathfrak{sl}_2$, the
production vertex has coefficient $1/(k+2)$, confirmed by
explicit computation at conformal weight~$2$.

In gravity, the fundamental field is the stress tensor $T$,
the Sugawara quadratic
$T = \frac{1}{k+2}(\frac{1}{2}{:}J^0 J^0{:}
+ {:}J^+ J^-{:})$
of the $\mathfrak{sl}_2$ connection.  The passage from $J^a$ to
$T$ is the Drinfeld--Sokolov reduction
$\widehat{\mathfrak{sl}}_2 \to \mathrm{Vir}_c$, implemented by
imposing the first-class constraint $J^+(z) = 1$.  The BRST
ghost pair $(b,c)$ that enforces this constraint carries a
$\mathbf{Z}$-grading invisible to the unconstrained theory.
The same split Casimir that provides the gluon production
vertex is present in the affine coproduct of $T$
(equation~\eqref{eq:delta-z-sugawara}), but the projection
$p \otimes p$ annihilates it: each summand of $\Omega$ has at
least one factor at conformal weight~$1$, and the Virasoro
sector has no weight-$1$ states.  At arity~$2$ and beyond, the
ghost-number obstruction of
Theorem~\ref{thm:gravitational-primitivity} provides a second,
independent, mechanism: even those HPL corrections that survive
the weight-$1$ argument are killed by the ghost-number mismatch
between the tree output (ghost $+1$ for source trees,
ghost~$-1$ for target trees) and the physical sector
(ghost~$0$).

The following table records the outcome across the standard
landscape of three-dimensional holomorphic-topological theories:

\begin{center}
\small
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{llcc}
\textbf{Theory} & \textbf{Boundary algebra}
  & $\boldsymbol{\Delta_z}$
  & \textbf{Mechanism} \\
\hline
Chern--Simons (gauge) &
  $\widehat{\mathfrak{g}}_k$ (affine KM) &
  $\ne \varepsilon$ &
  read as fundamental gluon splitting \\[2pt]
3d gravity &
  $\mathrm{Vir}_c = W_k(\mathfrak{sl}_2)$ &
  $= \varepsilon$ &
  ghost obstruction (DS) \\[2pt]
$bc$-$\beta\gamma$ (twisted hol.) &
  $H_k$ (Heisenberg) &
  $= \varepsilon$ &
  free field, $r = k/z$ \\[2pt]
M2 brane &
  DDCA &
  $\ne \varepsilon$ &
  no DS, fundamental modes \\[2pt]
Higher-spin gravity &
  $\mathcal{W}_N = W_k(\mathfrak{sl}_N)$ &
  $= \varepsilon$ &
  ghost obstruction (DS) \\[2pt]
M5 brane &
  $W_{1+\infty} = \mathrm{DS}(\widehat{\mathfrak{gl}}_K)$ &
  $= \varepsilon$ &
  ghost obstruction (DS) \\
\end{tabular}
\end{center}

The pattern is absolute: every boundary algebra obtained by
Drinfeld--Sokolov reduction has a primitive coproduct, and
every boundary algebra consisting of fundamental fields has a
nontrivial one.  Free fields are primitive for a different
reason and by a different mechanism; we return to this below.

\begin{remark}[Heisenberg: primitive without ghosts]
\label{rem:heisenberg-primitive}
\index{Heisenberg!primitive coproduct}
The Heisenberg algebra $H_k$ has a primitive coproduct because
its $r$-matrix $r^H(z) = k/z$ is a $c$-number: the target
product on $H_k \otimes H_k$ is twisted by a scalar, and
the HPL transfer produces no nontrivial corrections at any
arity.  The bar complex terminates at arity~$2$ (class~G,
shadow depth $r_{\max} = 2$).  There is no DS reduction and
no ghost obstruction.  The Virasoro algebra, by contrast, has
an operator-valued $r$-matrix
$r^{\mathrm{Vir}}(z) = (c/2)/z^3 + 2T/z$, infinite shadow
depth (class~M), and all $\Ainf$ products nontrivial.
Production vertices \emph{could} in principle be generated by
the HPL transfer.  They are killed by ghost number, not by
triviality of the input.  Free fields are primitive because
they are free.  Gravitational fields are primitive because they
are composite.
\end{remark}

\subsubsection*{Physical content: gravity without graviton
  production}

The strict primitivity of $\Delta_z^{\mathrm{grav}}$ is the
algebraic statement that three-dimensional gravity has no local
degrees of freedom.

The $\Ainf$ products
$\{m_n^{\mathrm{Vir}}\}_{n \ge 3}$ encode the self-interaction
of the stress tensor: nonlinear, infinite, and captured by the
higher $m_n$.
The coproduct corrections $\Delta_{z,n}$ for $n \ge 2$ would
encode graviton production: a graviton splitting into
constituents, each carrying a fraction of the energy.  In three
dimensions: the Weyl tensor vanishes identically, so the
Einstein equations fix the local geometry completely, leaving
only topological degrees of freedom.

\subsubsection*{Primitivity and the strict $\Ainf$ morphism}
\label{par:primitive-strict-morphism}
\index{coproduct!strict A-infinity morphism|textbf}
\index{A-infinity morphism@$\Ainf$ morphism!strict from primitivity}

The vanishing of $\Delta_{z,n}$ for $n \ge 2$ has an immediate
structural consequence: the primitive coproduct
$\Delta_z^{\mathrm{prim}} \colon x \mapsto
\tau_z(x) \otimes 1 + 1 \otimes x$ is a \emph{strict}
$\Ainf$-algebra morphism.  An $\Ainf$ morphism
$f \colon (A, \{m_n\}) \to (B, \{m'_n\})$ consists of
components $f_1, f_2, f_3, \ldots$ satisfying the morphism
Stasheff identities.  The morphism is strict if $f_n = 0$ for
$n \ge 2$, so that the identities reduce to
$f_1 \circ m_n = m'_n \circ f_1^{\otimes n}$.  The
gravitational coproduct satisfies this with
$f_1 = \Delta_{z,1}^{\mathrm{prim}}$: the compatibility
condition on the generator $T$ reads
\begin{equation}\label{eq:compatibility-primitive}
\Delta_z^{\mathrm{Vir}}(m_n(T^{\otimes n}))
\;=\;
m_n^{r(z)}\bigl(\Delta_z(T)^{\otimes n}\bigr),
\end{equation}
which expands to
\[
\tau_z(m_n(T^{\otimes n})) \otimes 1
\;+\; 1 \otimes m_n(T^{\otimes n})
\;=\;
m_n^{r(z)}\bigl(
(\tau_z(T) \otimes 1 + 1 \otimes T)^{\otimes n}\bigr).
\]
The left side is the primitive coproduct of the product; the
right side is the $r(z)$-twisted product of the primitive
coproducts.  No independent coproduct data enters at any
arity.  The entire dg-shifted Yangian of gravity is controlled
by two data: the collision residue
\begin{equation}\label{eq:gravity-r-matrix-determines-all}
r^{\mathrm{Vir}}(z)
\;=\;
\frac{c/2}{z^3} \;+\; \frac{2T}{z}
\end{equation}
and the $\Ainf$ product tower $\{m_n^{\mathrm{Vir}}\}$.  The
coproduct, which is the third datum of a general Yangian, adds
nothing.

The strictness of $\Delta_z^{\mathrm{prim}}$ as an $\Ainf$
morphism should be contrasted with the transferred products
$\{m_n^{\mathrm{Vir}}\}_{n \ge 3}$, which are genuinely
higher-order: $m_3 \ne 0$ forced by the quartic OPE pole,
$m_4 \ne 0$ forced by the Stasheff identity relating $m_3$
and $m_2$, and so on through the entire infinite tower
(class~M, shadow depth $r_{\max} = \infty$).  The gravitational
Yangian has maximal product complexity and zero coproduct
complexity.  The gauge-theoretic Yangian (affine KM) has
$m_2$ strict and $\Delta_z$ nontrivial: product complexity
$0$ and nonzero coproduct complexity.  The two theories exhaust
opposite corners of the $(m, \Delta)$-complexity plane.

\subsubsection*{The $r$-matrix seeds the genus tower}
\label{par:btz-r-matrix}

All gravitational reading data discussed here is channeled through the collision
residue~\eqref{eq:gravity-r-matrix-determines-all}.  In the corresponding
gravitational reading, this $r$-matrix seeds the genus tower and thereby
underlies the conjectural BTZ thermodynamic statements.

The self-sewing construction couples $r^{\mathrm{Vir}}(z)$
to the Hodge class on $\overline{\mathcal{M}}_{g,n}$.  At
genus~$1$, the self-loop graph
$\Gamma = \vcenter{\hbox{\begin{tikzpicture}[scale=0.4,
    baseline=-0.5ex]
\draw (0,0) circle (0.4);
\fill (0,-0.4) circle (0.05);
\end{tikzpicture}}}$
on $\overline{\mathcal{M}}_{1,1}$ extracts
$\kappa = c/2$ from the subleading pole of
$r^{\mathrm{Vir}}(z)$.  The leading pole $(c/2)/z^3$
contributes zero after the $\operatorname{Res}_{z=0}$
extraction (no $z^{-1}$ term); the subleading pole $2T/z$
contributes
$\operatorname{Res}_{z=0}\,2T/z\,dz = 2\langle T \rangle
= c/12$.  The sewing integral over
$\overline{\mathcal{M}}_{1,1}$ couples this to
$\lambda_1 = 1/24$, giving $F_1 = \kappa/24 = c/48$.
After ghost subtraction:
$F_1^{\mathrm{eff}} = \kappa_{\mathrm{eff}}/24 = (c-26)/48$.

The holomorphic contribution to the BTZ
partition function is read as the exponential of the all-genus generating function
$\sum_{g \ge 1} F_g^{\mathrm{eff}}\,x^{2g}
= \frac{c-26}{2}\bigl(\hat{A}(ix) - 1\bigr)$,
where $\hat{A}(x) = (x/2)/\sinh(x/2)$.  The corresponding
one-sided Cardy asymptotic entropy formula
$S_{\mathrm{Cardy}} = 2\pi\sqrt{ch/6}$ at large conformal
weight~$h$ is conjectural
(Conjecture~\ref{conj:gravity-line-identification};
compare Theorem~\ref{thm:gravity-cardy} for the two-sided formula):
it would be expected to arise from the modular
$S$-transform of this generating function.  The $r$-matrix is the genus-$0$,
arity-$2$ datum that seeds the entire tower; the corresponding
one-sided Cardy asymptotic entropy formula would be read as its high-temperature shadow.

\subsubsection*{Conical defects and BTZ from braiding}

Primitivity of $\Delta_z^{\mathrm{grav}}$ states that the BTZ
spectrum is not built by splitting gravitons.  Black holes
form by a different mechanism: accumulation of $r$-matrix
twists.  The BTZ black hole of mass $M$ and angular
momentum $J$ is a quotient of AdS$_3$ by a discrete isometry;
its partition function is read as a trace over a single
Virasoro module, not as a sum over multi-graviton states.
Conical defects of angular deficit $2\pi(1 - \sqrt{1 - 8Gh})$
are read as highest-weight modules $M_h$ of the Koszul dual
$\mathrm{Vir}_{26-c}$: they are topological
objects, classified by the monodromy of the $r$-matrix
braiding.  The fusion of two
conical defects is controlled by conformal blocks of
$\mathrm{Vir}_{26-c}$, parametrized by the Ponsot--Teschner
fusion kernel
$\mathbb{F}_{\alpha_s}^{\alpha_t}$
(\S\ref{subsec:gravity-r-matrix}).  The gravitational $S$-matrix
is a pure braiding matrix:
it rearranges line operators but never creates new ones.

This is the algebraic content of the classical observation that
three-dimensional gravity is topological.  The Einstein--Hilbert
action in three dimensions is equivalent to Chern--Simons theory
for $SL(2,\mathbf{R}) \times SL(2,\mathbf{R})$, and the phase
space consists of flat connections modulo gauge: a
finite-dimensional moduli space with no local excitations.  The
Koszul-dual formulation makes this precise.  The coproduct
$\Delta_z$ would create local excitations
(propagating gravitons).  Its vanishing states that the only
gravitational degrees of freedom are topological:
conical deficits, black holes, and
the modular data encoded in the braiding of line operators.
The $r$-matrix is the complete repository of this
topological content.

\subsubsection*{The $\mathcal{W}_N$ extension and
  non-principal reductions}
\label{par:wn-ghost-obstruction}
\index{W-algebra@$\mathcal{W}$-algebra!ghost-number obstruction}

Theorem~\ref{thm:gravitational-primitivity} covers all
principal DS reductions, including $\mathcal{W}_N$-gravity
($\mathfrak{g} = \mathfrak{sl}_N$) and the M5 brane
($\mathfrak{g} = \mathfrak{gl}_K$).  The essential input is
the ghost-number degree $\mathrm{gh}(h) = -1$, which holds
for every principal nilpotent.  We verify this explicitly for
$\mathfrak{sl}_N$.

The DS reduction of $\widehat{\mathfrak{sl}}_N$ by the
principal nilpotent $f = \sum_{i=1}^{N-1} E_{i+1,\,i}$
produces $N(N-1)/2$ ghost pairs
$(b_\alpha, c_\alpha)$ indexed by positive roots
$\alpha \in \Phi^+(\mathfrak{sl}_N)$.  The
$\operatorname{ad}(h_0)$-eigenvalue of each root is
its Dynkin grade $j(\alpha) \in \{1, 2, \ldots, N-1\}$.
The conformal weights of the ghost pair at grade~$j$ are
$(1-j,\, j)$, so higher-grade ghost pairs have higher
conformal weight.  The ghost number, however, is always
$\mathrm{gh}(b_\alpha) = -1$, $\mathrm{gh}(c_\alpha) = +1$,
regardless of the grade.  The Dynkin grade controls the
conformal weight of the intermediate state but has no effect
on the ghost-number argument: every acyclic pair is separated
by exactly one unit of ghost number, and every contraction
by $h_N$ lowers ghost number by exactly~$1$.

The proof of
Theorem~\ref{thm:gravitational-primitivity} therefore applies
verbatim to $\mathcal{W}_N$ for every~$N$: the spin-$s$
generators $W_s$ ($s = 2, \ldots, N$) are BRST-closed,
$i_\infty(W_s) = i(W_s)$, and every HPL tree for the arity-$n$
coproduct ($n \ge 2$) vanishes after projection by the same
ghost-number bottleneck.  Higher-spin gravitons do not split
either.

\begin{remark}[Non-principal nilpotents]
\label{rem:non-principal-ghost}
\index{Drinfeld--Sokolov reduction!non-principal}
For a non-principal nilpotent $f$ in $\mathfrak{g}$, the BRST
complex $C^{\mathrm{DS}}(\mathfrak{g}, f)$ has ghost pairs
$(b_\alpha, c_\alpha)$ indexed by positive roots with
$\langle \alpha, h_0 \rangle > 0$, where $h_0$ is the neutral
element of the $\mathfrak{sl}_2$-triple of~$f$.  The homotopy
$h_f$ still has ghost-number degree~$-1$: every acyclic pair
in the Chevalley--Eilenberg complex of $\mathfrak{n}_+(f)$
is separated by one unit of ghost number.  Thus the
ghost-number obstruction kills all \emph{nonlinear} HPL
corrections: $\Delta_{z,n}^{\mathcal{W}(\mathfrak{g},f)} = 0$
for $n \ge 2$ whenever the DS reduction produces an SDR with
$\mathrm{gh}(h_f) = -1$.  This covers hook-type, subregular,
and minimal nilpotents in all types.  However, the
\emph{linear} transferred coproduct $\Delta_{z,1}$ requires a
separate argument: for non-principal reductions whose
$\mathcal{W}$-algebra retains weight-$1$ generators (e.g.\
the Bershadsky--Polyakov algebra from $(\mathfrak{sl}_3,
f_{(2,1)})$, which has an $\widehat{\mathfrak{sl}}_2$
subalgebra at weight~$1$), the projection~$p$ does not
automatically kill the split Casimir cross-terms.  Full
primitivity for non-principal reductions with surviving
weight-$1$ generators remains open.
\end{remark}

\begin{remark}[BP coproduct: resolved]
\label{rem:bp-coproduct-resolved}
\index{Bershadsky--Polyakov algebra!coproduct resolution}
The linear transferred coproduct $\Delta_{z,1}$ \emph{is}
primitive for the Bershadsky--Polyakov algebra
$\mathcal{W}(\mathfrak{sl}_3, f_{(2,1)})$ on all free
generators, but not on composites.  The three free generators
behave as follows:
\begin{enumerate}
\item $G^+ = J^{E_{23}}$ and $G^- = J^{E_{31}}$ are
  fundamental affine generators.  Because $E_{23}$ and
  $E_{31}$ are root vectors for simple roots of the
  $\mathfrak{sl}_2$-complement, the split Casimir
  $\Omega = \sum t^a \otimes t_a$ contributes no cross-term
  of the form $J^{E_{23}} \otimes (\cdots)$ or vice versa:
  the Casimir component $E_{23} \otimes E_{32}$ pairs a
  surviving direction with a BRST-exact one.  Thus
  $\Delta_{z,1}(G^\pm) = G^\pm \otimes 1 + 1 \otimes G^\pm$.
\item The weight-$1$ generator $J$ involves Cartan currents
  and ghost bilinears.  Ghost cross-terms are killed by the
  ghost-number projection $p \otimes p$: any term with
  nonzero ghost number in either tensor factor vanishes
  under projection to the ghost-number-zero summand.
  Thus $\Delta_{z,1}(J) = J \otimes 1 + 1 \otimes J$.
\item The stress tensor $T$ is a \emph{composite} (Sugawara
  plus ghost corrections), not a free generator.  Its
  coproduct is \emph{non-primitive}: the $\mathfrak{sl}_3$
  Casimir has a surviving Cartan component
  $H_1 \otimes H_1$ that projects to $J \otimes J$ under
  $p \otimes p$, giving a nonzero cross-term
  $\Delta_{z,1}(T) \ni (1/z)\, J \otimes J$.
\end{enumerate}
This reveals a clean dichotomy between principal and
non-principal DS reductions:
\begin{itemize}
\item \textbf{Principal} ($\mathrm{Vir}$, $\mathcal{W}_N$):
  \emph{all} fields are primitive, including composites,
  because no weight-$1$ generators survive the reduction ---
  the ghost-number obstruction kills everything.
\item \textbf{Non-principal} (Bershadsky--Polyakov and its
  generalisations): free generators are primitive, but
  composites built from surviving weight-$1$ fields are
  non-primitive.  The surviving weight-$1$ Casimir component
  provides exactly the cross-terms that escape the
  ghost-number filter.
\end{itemize}
The gap flagged in Remark~\ref{rem:non-principal-ghost} is
therefore resolved: primitivity holds on generators (which
suffice for the coalgebra structure), while the composite
non-primitivity is a \emph{consequence} of the Sugawara
construction and does not obstruct the HPL transfer.
\end{remark}

\begin{remark}[Scope and the infinite tower]
\label{rem:primitivity-scope}
\label{rem:coproduct-gap-closed}
\index{coproduct!scope of primitivity}
Theorem~\ref{thm:gravitational-primitivity} resolves the
coproduct question completely: strict primitivity holds at all
arities for all principal DS reductions.  The ghost-number
obstruction extends to all $N$: the total ghost number
$\mathrm{gh}_{\mathrm{tot}} = \sum_\alpha \mathrm{gh}(c_\alpha)$
is a global $\mathbb{Z}$-grading regardless of the number of ghost
pairs, the homotopy~$h_f$ has $\mathrm{gh}(h_f) = -1$ for
structural reasons (forced by the SDR equation
$\pi \circ h = 0$, $h \circ \iota = 0$, $h^2 = 0$),
and $p \otimes p$ kills all nonzero ghost-number
components of $\Delta^{V_k}_{z,n}$ since both tensor factors
project to the ghost-number-zero summand.  The remaining
gaps of Remark~\ref{rem:ds-hpl-honest-gaps} concern the
transferred $\Ainf$ products and the dg-shifted Yangian axioms,
not the coproduct.  The infinite complexity of
class~$\mathbf{M}$ algebras ($\mathrm{Vir}$,
$\mathcal{W}_N$ for all $N$) resides entirely in the
transferred products $\{m_n^W\}_{n \ge 3}$ and the
$r(z)$-twisted target product.  The coproduct is trivial.
The nontrivial gravitational
physics lives entirely in the $r$-matrix.
\end{remark}

%====================================================================
% MOVEMENT IV: THE GENUS TOWER
%====================================================================

\subsection{Movement IV: the genus tower}
\label{subsec:gravity-genus-tower}

\subsubsection*{Genus-$1$ curvature}

The effective gravitational curvature
(Remark~\ref{rem:gravity-kappa-vs-kappa-eff}) is
\begin{equation}\label{eq:gravity-kappa}
\kappa_{\mathrm{eff}}
\;=\;
\frac{c - 26}{2}
\;=\;
3k - 13.
\end{equation}
The bar differential at genus~$1$ acquires curvature
$d_{\mathrm{fib}}^2 = \kappa_{\mathrm{eff}} \cdot \omega_1$
from the self-loop graph on $\overline{\cM}_{1,1}$.  The genus-$1$
free energy:
\begin{equation}\label{eq:gravity-F1}
F_1(\mathrm{Vir}_c)
\;=\;
\kappa_{\mathrm{eff}} \cdot \frac{1}{24}
\;=\;
\frac{c - 26}{48}.
\end{equation}
The corresponding $\eta$-function bookkeeping packages the
genus-$1$ oscillator contribution as a function of~$\tau$; it
should not be conflated with the constant free energy~$F_1$.
Here $\eta(\tau) = q^{1/24}\prod_{n \ge 1}(1 - q^n)$.

\begin{remark}[Intrinsic vs.\ effective free energy at genus $1$]
\label{rem:gravity-F1-intrinsic-vs-effective}
The \emph{intrinsic} genus-$1$ free energy of $\mathrm{Vir}_c$ is
$F_1^{\mathrm{intr}} = \kappa(\mathrm{Vir}_c)/24 = c/48$,
using the intrinsic modular characteristic $\kappa = c/2$
(see Computation~\ref{comp:genus1-free-energy}).
The \emph{effective} value $F_1^{\mathrm{eff}} = (c-26)/48$
in~\eqref{eq:gravity-F1} is the physical quantity after
ghost subtraction ($\kappa_{\mathrm{ghost}} = -13$).  At
$c = 26$: $F_1^{\mathrm{eff}} = 0$ (the genus tower collapses).
The corresponding partition functions on a torus~$E_\tau$ are
$Z_1^{\mathrm{intr}}(\tau) = \eta(\tau)^{-c/2}$ and
$Z_1^{\mathrm{eff}}(\tau) = \eta(\tau)^{-(c-26)/2}$.
\end{remark}

\subsubsection*{All genera: the $\hat A$-genus and the Wick rotation}

The genus-$g$ free energy is
$F_g = \kappa_{\mathrm{eff}} \cdot \int_{\overline{\cM}_g}
\lambda_g$, with generating function
\begin{equation}\label{eq:gravity-generating}
\sum_{g \ge 1} F_g\,x^{2g}
\;=\;
\frac{c - 26}{2}\bigl(\hat{A}(ix) - 1\bigr),
\qquad
\hat{A}(x) = \frac{x/2}{\sinh(x/2)}.
\end{equation}
The argument~$ix$, not~$x$, is the content.
The $\hat{A}$-genus
$\hat{A}(x) = (x/2)/\sinh(x/2)$
is the characteristic series of the Dirac index:
its Taylor coefficients are the alternating Bernoulli
numbers.
The substitution $x \mapsto ix$ converts
$\sinh(ix/2) = i\sin(x/2)$, producing
$\hat{A}(ix) = (x/2)/\sin(x/2)$
--- all coefficients positive.
This is the Wick rotation
from Lorentzian (oscillatory Dirac index, alternating signs)
to Euclidean (convergent Hodge integrals, definite signs):
the factor $i^{2g} = (-1)^g$ on each monomial~$x^{2g}$
absorbs the alternation.  The individual coefficients
are the Faber--Pandharipande Hodge integrals
$\int_{\overline{\cM}_g} \lambda_g
= \frac{2^{2g-1}-1}{2^{2g-1}} \cdot \frac{|B_{2g}|}{(2g)!}
=: \lambda_g^{\mathrm{FP}} > 0$
for all $g \ge 1$, and their generating function is
$\hat{A}(x) = (x/2)\operatorname{csch}(x/2)$.
The prefactor $\kappa_{\mathrm{eff}} = (c{-}26)/2$ is the
single datum from the boundary algebra after ghost
subtraction.\label{obs:genus-tower-wick-anomaly}%
\index{Wick rotation!anomaly interpretation}%
\index{genus tower!as Wick rotation anomaly}%

Since every $\lambda_g^{\mathrm{FP}}$ is positive,
the genus tower has uniform sign
$\operatorname{sgn}(\kappa_{\mathrm{eff}})$
at every genus: subcritical ($c < 26$) gives
$F_g < 0$ for all~$g$;
supercritical ($c > 26$) gives $F_g > 0$;
critical ($c = 26$) gives $F_g = 0$.
No sign alternation survives the Wick rotation.

\begin{remark}[Why the $\hat{A}$-genus]
\label{rem:why-A-hat-genus}
\index{A-hat genus@$\hat{A}$-genus!why it appears}%
Three independent routes force $\hat{A}$ rather than the Todd class:
\begin{enumerate}
\item \emph{Symmetry.}  The genus tower uses only even powers of~$x$.
  The Todd class $\mathrm{Td}(x) = x/(1-e^{-x})$ has odd powers.
  The product $\mathrm{Td}(x)\cdot\mathrm{Td}(-x)
  = x^2/((1-e^{-x})(e^x-1)) = x^2/(4\sinh^2(x/2))$ is even;
  its positive square root $x/(2\sinh(x/2)) = \hat{A}(x)$.
\item \emph{KO orientation.}  The Hodge bundle $\mathbb{E}$ on
  $\overline{\mathcal{M}}_g$ carries a real structure via Serre
  duality ($\mathbb{E}^* \simeq \mathbb{E} \otimes \omega$).
  Real vector bundles are oriented by KO-theory, whose
  multiplicative genus is the $\hat{A}$-genus.  Complex vector
  bundles would give the Todd class via KU-theory.
\item \emph{Explicit integration.}  The Faber--Pandharipande
  integral
  $\lambda_g^{\mathrm{FP}}
  = \frac{2^{2g-1}-1}{2^{2g-1}} \cdot \frac{|B_{2g}|}{(2g)!}$
  equals the coefficient of $x^{2g}$ in
  $\hat{A}(ix) - 1 = (x/2)/\sin(x/2) - 1$
  (the standard Mittag-Leffler expansion
  $z/\sin(z) = 1 + 2\sum_{g \ge 1}
  \tfrac{(2^{2g-1}-1)|B_{2g}|}{(2g)!}\,z^{2g}$
  identifies the Taylor coefficients with
  $\lambda_g^{\mathrm{FP}}$; verified numerically through
  $g = 6$).  The factor
  $(2^{2g-1}-1)/2^{2g-1}$ arises from the boundary-strata
  decomposition $\csc(u) = \cot(u) + \tan(u/2)$.
\end{enumerate}
\end{remark}

\medskip

The vanishing locus $\kappa_{\mathrm{eff}} = 0$ admits
a modular reading.
The genus-$1$ partition function
$Z_1^{\mathrm{eff}}(\tau) = \eta(\tau)^{-\kappa_{\mathrm{eff}}}$
transforms under the modular $S$-transform
$\tau \mapsto -1/\tau$ with weight $-\kappa_{\mathrm{eff}}/2$,
because $\eta(-1/\tau) = \sqrt{-i\tau}\cdot\eta(\tau)$.
At $\kappa_{\mathrm{eff}} = 0$ the weight vanishes:
$c = 26$ is the unique central charge at which
$Z_1^{\mathrm{eff}}$ is $S$-invariant.
The genus tower therefore measures the
\emph{breaking of Wick-rotation symmetry}:
$F_g \ne 0$ at genus~$g$ is the $g$-th
anomaly coefficient, and the conjectural Cardy entropy
$S_{\mathrm{Cardy}} = 2\pi\sqrt{ch/6}$
\textup{(}Conjecture~\textup{\ref{conj:gravity-line-identification})}
would be its high-temperature shadow, extracted by composing
the $S$-transform with the saddle-point approximation.
The single chain
\[
\text{quartic pole}
\;\longrightarrow\;
\kappa
\;\longrightarrow\;
d_{\mathrm{fib}}^2 = \kappa_{\mathrm{eff}}\cdot\omega_g
\;\longrightarrow\;
F_g \ne 0
\;\longrightarrow\;
S_{\mathrm{Cardy}}
\]
connects the OPE datum to the entropy.

\begin{remark}[Modularity from the open sector]
\label{rem:gravity-modularity-open-sector}
\index{modularity!from open sector}
The genus tower is not an axiom on the closed algebra
$\mathrm{Vir}_c$.  It arises from trace and clutching on the
gravitational open-sector factorization dg-category
$\cC_{\mathrm{grav}}^{\mathrm{op}}$
\textup{(}Definition~\textup{\ref{def:grav-open-category})}: the
annulus trace $\int_{S^1} \cC_{\mathrm{grav}}^{\mathrm{op}}
\simeq HH_*(\mathrm{Vir}_c)$ is the first modular shadow
\textup{(}\eqref{eq:grav-annulus-trace}\textup{)}, and
clutching along shared boundary circles raises genus.  The
scalar $\kappa_{\mathrm{eff}}$ enters as the trace of the
cyclic pairing on $\End(b)$, not as primitive closed-string data.
The MC equation on the bordered FM compactification
$\overline{C}{}^{\,\mathrm{oc}}_{k,\vec{m}}$ packages both
sectors into a single identity whose closed projection recovers
$F_g = \kappa_{\mathrm{eff}} \int_{\overline{\cM}_g}\lambda_g$.
\end{remark}

\subsubsection*{The effective scalar partition function}

The effective scalar bookkeeping series for the SL$(2,\R) \times$
SL$(2,\R)$ theory (left$+$right) is
\[
Z_{\mathrm{grav}}^{\mathrm{scal}}(\tau, \bar\tau)
\;=\;
|\eta(\tau)|^{(c-26)/6}.
\]
At $c = 26$ ($k = 13/3$):
$Z_{\mathrm{grav}}^{\mathrm{scal}} = 1$, reflecting effective scalar
anomaly cancellation rather than triviality of the full coupled
partition function.  At $c = 13$ (Koszul self-duality):
$Z_{\mathrm{grav}}^{\mathrm{scal}} = |\eta(\tau)|^{-13/6}$.

\subsubsection*{The shadow tower}

\begin{center}
\small
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{lll}
\textbf{Shadow} & \textbf{Algebraic origin (Vol~I)} & \textbf{Gravitational counterpart} \\
\hline
$\kappa_{\mathrm{eff}} = (c-26)/2$
  & effective curvature after ghost subtraction
  & one-loop Weyl anomaly; $= 0$ iff $c = 26$ \\[2pt]
$\mathfrak{C}(\mathrm{Vir}_c) = -c$
  & cubic shadow (Vol~I)
  & cubic graviton vertex; $\propto \ell/(16\pi G)$ at Brown--Henneaux \\[2pt]
$\mathfrak{Q} = 10/[c(5c+22)]$
  & quartic contact shadow (Computation~\ref{comp:gravity-quartic-correction})
  & genus-$2$ correction $F_2^{\mathrm{contact}}$ \\[2pt]
$o_5 \ne 0$
  & quintic obstruction (Vol~I, non-vanishing)
  & genus-$2$ five-point contribution to $F_2$
\end{tabular}
\end{center}

The shadow tower does not terminate
(by the explicit Virasoro quartic-contact computation of Volume~I):
the $r$-th shadow determines the $r$-graviton
vertex, and no finite
truncation determines the full perturbative hierarchy.

\begin{computation}[Quartic contact shadow and genus-$2$ contact
correction; \ClaimStatusProvedHere]
\label{comp:gravity-quartic-correction}%
\index{quartic contact shadow!genus-2 contact correction}%
The quartic contact shadow
$\mathfrak{Q}^{\mathrm{contact}}_{\mathrm{Vir}} = 10/[c(5c+22)]$
from the quartic-contact analysis of Volume~I
determines the genus-$2$ contact correction
$F_2^{\mathrm{contact}}$, the
leading nonlinear partition-function correction beyond the
one-loop sector.

\emph{Explicit evaluation.}  At the Brown--Henneaux central charge
$c = 6k = 3\ell/(2G)$:
\begin{alignat*}{2}
\mathfrak{Q}(c = 6k)
\;&=\;
\frac{10}{6k(30k+22)}
\;=\;
\frac{5}{6k(15k+11)},
\\[4pt]
\intertext{and the genus-$1$ Hessian correction
$\delta_H^{(1)} = 120/[c^2(5c+22)]$ evaluates to}
\delta_H^{(1)}(c = 6k)
\;&=\;
\frac{120}{36k^2(30k+22)}
\;=\;
\frac{5}{3k^2(15k+11)}.
\end{alignat*}
In the semiclassical limit $k \to \infty$ (equivalently
$G \to 0$):
\[
\mathfrak{Q} \;\sim\; \frac{1}{18k^2},
\qquad
\delta_H^{(1)} \;\sim\; \frac{1}{9k^3}.
\]
The quartic correction enters the genus-$2$ partition function
via the $r = 4$ planted-forest graph amplitude:
\[
F_2^{\mathrm{contact}}
\;=\;
\mathfrak{Q} \cdot \int_{\overline{\cM}_{2}} \psi_1^2
\;=\;
\frac{10}{c(5c+22)} \cdot \frac{1}{240}
\;=\;
\frac{1}{24\,c(5c+22)},
\]
where $\int_{\overline{\cM}_2} \psi_1^2 = 1/240$ by
Faber's formula.

\emph{Gravitational interpretation.}
The quartic shadow $\mathfrak{Q}$ is read as the four-graviton
contact coupling: a connected genus-$0$ amplitude with four external
stress-tensor insertions, exchanged through the one-loop wheel
graph $C_4$.  Its nonvanishing ($\mathfrak{Q} \neq 0$ for
$c \neq 0$) is the algebraic witness that $3$d gravity has genuine
nonlinear dynamics beyond one loop,
and the shadow depth of $\mathrm{Vir}_c$ being infinite
(class~M) is read as meaning that the full graviton S-matrix
has no finite truncation.

\emph{Complementarity.}  Under $c \mapsto 26 - c$
(Koszul duality $\mathrm{Vir}_c^! = \mathrm{Vir}_{26-c}$):
\[
\mathfrak{Q}(c) + \mathfrak{Q}(26-c)
\;=\;
\frac{10}{c(5c+22)} + \frac{10}{(26-c)(152-5c)}.
\]
At $c = 13$ (self-dual): $\mathfrak{Q}(13) = 10/(13 \cdot 87) =
10/1131$.  The complementarity of the quartic shadow across the
Koszul involution is
the first nonlinear witness of the duality between boundary and
line sectors.
\end{computation}

%====================================================================
% MOVEMENT V: THE PHYSICS EMERGES
%====================================================================

\subsection{Movement V: the physics of 3d gravity}
\label{subsec:gravity-physics}

Everything above was derived from the $\lambda$-bracket
$\{T_\lambda T\} = \partial T + 2T\lambda + (c/12)\lambda^3$.
We now show that, via the physics-bridge theorem and the
Chern--Simons realization, this algebraic package recovers the
three-dimensional Einstein-gravity package with
$\Lambda = -1/\ell^2$.

\subsubsection*{The Chern--Simons origin}

Three-dimensional gravity with $\Lambda < 0$ is equivalent to
$SL(2,\R) \times SL(2,\R)$ Chern--Simons at level
$k = \ell/(4G)$, $c = 6k = 3\ell/(2G)$
(Ach\'ucarro--Townsend, Witten).  The dreibein $e^a$ and spin
connection $\omega^a$ combine into $A_\pm = \omega \pm e/\ell$;
flat connections $F_\pm = 0$ are constant-curvature geometries.

\begin{theorem}[Brown--Henneaux; \ClaimStatusProvedElsewhere]
\label{thm:brown-henneaux}
\index{Brown--Henneaux central charge|textbf}
The asymptotic symmetry algebra of $SL(2,\R)$ CS at level~$k$ on
$\R_+ \times \C$ is $\mathrm{Vir}_c$, $c = 6k$.
\end{theorem}

\begin{proof}
The boundary condition $A_z|_{t=0} = J_- + T(z)\,J_+$
(Fefferman--Graham asymptotics) has residual gauge
$g = \exp(\epsilon\,J_+)$ acting by
$\delta_\epsilon T = \epsilon' T + 2\epsilon T' +
(c/12)\epsilon'''$, the Virasoro coadjoint action with
central charge $c = 6k$.  The CS symplectic form induces
the Poisson bracket
$\{T(z), T(w)\}_{\mathrm{PB}} = (c/2)\delta'''(z-w) +
2T(w)\delta'(z-w) + T'(w)\delta(z-w)$,
which is~\eqref{eq:gravity-input} in distribution form.
\end{proof}

\subsubsection*{Physics-bridge verification for gravity}

\begin{theorem}[Gravitational MC element; \ClaimStatusProvedHere]
\label{thm:gravity-mc}
\index{universal MC element!3d!gravitational}
$SL(2,\R)$ Chern--Simons theory at level~$k$ on $\R_+ \times \C$
lies in the scope of Theorem~\ref{thm:physics-bridge}.  The resulting
logarithmic $\SCchtop$-algebra determines an MC element
$\alpha_{\mathrm{grav}} \in \mc(\gSC_T)$ whose faces are:
\begin{enumerate}[label=\textup{(\roman*)}]
\item closed face $= \mathrm{Vir}_c$,
\item open face $= $ abstract line-side package
  \textup{(Theorem~\ref{thm:lines_as_modules})},
\item spectral $R$-matrix $= $ abstract line-side braiding,
\item genus face $= \Theta_{\mathrm{Vir}_c}$,
\item PVA descent $= \{T_\lambda T\}$.
\end{enumerate}
Items \textup{(i)--(v)} are proved unconditionally at the abstract
level.  The concrete identification of the open face with
$\mathrm{Vir}_{26-c}$-modules and of the spectral $R$-matrix with
the Virasoro fusion kernel is conjectural
\textup{(Remark~\ref{rem:gravity-mc-scope})}.
\end{theorem}

\begin{proof}
The gauge-fixed BV-BRST complex of Chern--Simons theory is cubic in
$A$, so the one-loop finiteness input of Theorem~\ref{thm:physics-bridge}
holds by power counting.  Its propagator is
$K(t,z) = \Theta(t)/(2\pi z)$, which has the required meromorphic
holomorphic behavior and ordered topological dependence.  The
interaction vertices are polynomial, so the renormalized
configuration-space integrals are precisely of the type covered by the
bridge theorem.  Hence the theory defines a logarithmic
$\SCchtop$-algebra.  Each listed projection of
$\alpha_{\mathrm{grav}}$ was computed in Movements~I--IV\@.
The abstract line-side package exists by
Theorem~\ref{thm:lines_as_modules}; its braiding is the abstract
spectral $R$-matrix of the bar complex.
\end{proof}

\begin{remark}[Scope: abstract vs.\ concrete line-side identifications]
\label{rem:gravity-mc-scope}
The open face and spectral $R$-matrix in
Theorem~\ref{thm:gravity-mc} are proved at the \emph{abstract}
level: the line-side package exists as the open-colour $E_1$-module
category of the bar complex, and the braiding is its abstract
spectral $R$-matrix.  The \emph{concrete} identification of this
package with highest-weight $\mathrm{Vir}_{26-c}$-modules and of the
braiding with the Virasoro fusion kernel is conjectural; see
Remark~\ref{rem:gravity-line-virasoro-realization} and
Conjecture~\ref{conj:gravity-line-identification}.
\end{remark}

\subsubsection*{BTZ black holes as MC deformations}

\begin{definition}[BTZ deformation]
\label{def:btz-deformation}
\index{BTZ black hole!as MC deformation}
A BTZ black hole of mass~$M$ and angular momentum~$J$ is a MC
deformation $\delta\alpha \in \gSC_T$ of $\alpha_{\mathrm{grav}}$,
satisfying $D_{\alpha_{\mathrm{grav}}}(\delta\alpha) = 0$,
corresponding to the highest-weight state $|h, \bar h\rangle$ with
$h = (M\ell + J)/2$, $\bar h = (M\ell - J)/2$.
\end{definition}

\subsubsection*{Toward the two-sided Cardy asymptotic entropy formula expected from the Virasoro bar-complex argument: partition function and modular $S$-transform}

The genus-$1$ Virasoro MC element $\Theta_{\mathrm{Vir}_c}^{(1)}(\tau)$
produces the boundary partition function.  Write $q = e^{2\pi i\tau}$ for the modular
parameter of the boundary torus.  The categorical trace gives
\begin{equation}\label{eq:gravity-Z1}
Z_1(\tau)
\;=\;
\Tr_{\mathrm{Vir}_c}\!\left(q^{L_0 - c/24}\right)
\;=\;
\frac{q^{-c/24}}{\prod_{n=2}^{\infty}(1-q^n)},
\end{equation}
the product starting at $n = 2$ because $\mathrm{Vir}_c$ has no
weight-$1$ states.  The full partition function counting all states
at inverse temperature $\beta = 2\pi\,\mathrm{Im}\,\tau$ is
\begin{equation}\label{eq:gravity-Z-full}
Z(\tau,\bar\tau)
\;=\;
\sum_{h,\bar h}
d(h,\bar h)\,q^{h - c/24}\,\bar q^{\bar h - c/24}.
\end{equation}
Assuming modular invariance
$Z(-1/\tau, -1/\bar\tau) = Z(\tau,\bar\tau)$ together with
vacuum dominance in the high-temperature limit, one is led, in each
chiral sector, to the estimate
\begin{equation}\label{eq:gravity-vacuum-dominance}
Z_{\mathrm{chiral}}(\tau)
\;\sim\;
\exp\!\left(
  \frac{\pi c}{12}\cdot\frac{1}{\mathrm{Im}(-1/\tau)}
\right)
\qquad (\mathrm{Im}\,\tau \to 0^+).
\end{equation}
Here $Z_{\mathrm{chiral}}(\tau)$ denotes the holomorphic contribution
implicit in the full partition function~\eqref{eq:gravity-Z-full}.
In the conjectural Virasoro bar-complex picture, the $S$-transform acts on
$\Theta_{\mathrm{Vir}_c}^{(1)}(\tau)$ by exchanging the two boundary cycles of
the genus-$1$ bordered FM compactification; vacuum dominance would
amount to the statement that the arity-$0$ component of that
Virasoro datum, $\Theta^{(1)}_{n=0} = \kappa\cdot\omega_1$, controls the
exponential growth of states.

\subsubsection*{Saddle-point extraction for the two-sided Cardy asymptotic entropy formula expected from the Virasoro bar-complex argument}

Under these assumptions, the chiral degeneracy contribution at large $h$ is
extracted by inverse Laplace transform from
~\eqref{eq:gravity-vacuum-dominance}.  The
saddle-point of $f(\tau) = \pi c/(12\tau) + 2\pi\tau(h - c/24)$
is at
\begin{equation}\label{eq:gravity-saddle}
\tau_*
\;=\;
i\sqrt{\frac{c}{24(h - c/24)}}
\;\approx\;
i\sqrt{\frac{c}{24h}}
\qquad (h \gg c),
\end{equation}
which would give the one-sided contribution
$f(\tau_*) \approx 2\pi\sqrt{ch/6}$.  Applying the same argument to the
$\bar\tau$-sector would yield the two-sided formula stated below.

\begin{theorem}[Two-sided Cardy asymptotic entropy formula expected from the Virasoro bar-complex argument;
\ClaimStatusConjectured]
\label{thm:gravity-cardy}
\label{prop:cardy-mc}
\index{two-sided Cardy asymptotic entropy formula!expected from the Virasoro bar-complex argument|textbf}
\index{BTZ black hole!two-sided Cardy asymptotic entropy formula expected from the Virasoro bar-complex argument}
The corresponding two-sided asymptotic entropy formula expected from
applying the Virasoro bar-complex argument to the left/right sectors at
large left/right conformal weights is
\begin{equation}\label{eq:gravity-cardy}
\boxed{
S_{\mathrm{Cardy}}
\;=\;
2\pi\sqrt{\frac{c\,h}{6}}
\;+\;
2\pi\sqrt{\frac{c\,\bar h}{6}}.
}
\end{equation}
This would be the logarithmic growth rate of MC deformations
of $\alpha_{\mathrm{grav}}$ at highest weight $(h,\bar h)$.
\end{theorem}

\begin{proof}[Argument sketch]
The corresponding Virasoro bar-complex argument would rest on three inputs:
\begin{enumerate}[label=(\roman*)]
\item The curvature $\kappa(\mathrm{Vir}_c) = c/2$ from the
  genus-$1$ self-loop, which determines
  $\Theta^{(1)}_{n=0} = \kappa\cdot\omega_1$ and hence the
  vacuum energy $-c/24$.
\item A modular invariance of $Z(\tau,\bar\tau)$ expected to
  follow from the consistency of the genus-$1$
  MC equation on the moduli stack $\cM_{1,1}$: the modular
  group acts as autoequivalences of the bordered FM
  compactification, and the MC element is expected to be
  invariant under this action.
\item A corresponding saddle-point analysis at
  $\tau_* = i\sqrt{c/(24h)}$~\eqref{eq:gravity-saddle},
  which would then follow from the inverse-Laplace analysis of the
  high-temperature estimate~\eqref{eq:gravity-vacuum-dominance}.
\end{enumerate}
The curvature input is determined by the single datum
$\kappa = c/2$, itself determined by the quartic OPE pole in
$\{T_\lambda T\}$.  Together with the expected genus-$1$ modular
consistency and the corresponding saddle-point analysis, this would be
the corresponding Virasoro bar-complex argument on the left/right
sectors for the two-sided Cardy asymptotic entropy formula.
\end{proof}

\subsubsection*{BTZ Bekenstein--Hawking entropy formula conditional on the two-sided Cardy asymptotic entropy formula expected from the Virasoro bar-complex argument}

\begin{theorem}[BTZ Bekenstein--Hawking entropy formula conditional on the two-sided Cardy asymptotic entropy formula expected from the Virasoro bar-complex argument; \ClaimStatusConjectured]
\label{thm:gravity-btz-entropy}
\index{BTZ black hole!BTZ Bekenstein--Hawking entropy formula conditional on the two-sided Cardy asymptotic entropy formula expected from the Virasoro bar-complex argument|textbf}
\index{Bekenstein--Hawking entropy formula!BTZ, conditional on the two-sided Cardy asymptotic entropy formula expected from the Virasoro bar-complex argument}
Conditional on the two-sided Cardy asymptotic entropy formula expected from the Virasoro bar-complex argument,
the BTZ Bekenstein--Hawking entropy formula would be
\begin{equation}\label{eq:gravity-btz-entropy}
\boxed{
S_{\mathrm{BTZ}}
\;=\;
\frac{2\pi r_+}{4G}
\;=\;
2\pi\sqrt{\frac{c\,h}{6}}
\;+\;
2\pi\sqrt{\frac{c\,\bar h}{6}},
}
\end{equation}
where $c = 3\ell/(2G)$ is the Brown--Henneaux central charge
\textup{(}Theorem~\textup{\ref{thm:brown-henneaux})},
$h,\bar h$ are the left and right conformal weights,
and $r_+$ is the outer horizon radius.
Equivalently, $S_{\mathrm{BTZ}} = 2\pi\sqrt{\kappa h/3}
+ 2\pi\sqrt{\kappa\bar h/3}$ with $\kappa = c/2$.
\end{theorem}

\begin{proof}[Conditional derivation]
Assuming the two-sided Cardy asymptotic entropy formula expected from the
Virasoro bar-complex argument~\eqref{eq:gravity-cardy}, substituting
$c = 3\ell/(2G)$, $h = (M\ell+J)/2$,
$\bar h = (M\ell-J)/2$ into that two-sided Cardy asymptotic entropy
formula~\eqref{eq:gravity-cardy}.
The standard BTZ horizon radii satisfy
$M = (r_+^2+r_-^2)/(8G\ell^2)$,
$J = r_+r_-/(4G\ell)$, whence
$M\ell + J = (r_++r_-)^2/(8G\ell)$ and
$M\ell - J = (r_+-r_-)^2/(8G\ell)$.
Substituting, with $c/6 = \ell/(4G)$:
\begin{align*}
S_{\mathrm{BTZ}}
&\;=\;
2\pi\sqrt{\frac{\ell}{4G}\cdot\frac{(r_++r_-)^2}{16G\ell}}
\;+\;
2\pi\sqrt{\frac{\ell}{4G}\cdot\frac{(r_+-r_-)^2}{16G\ell}}
\\[4pt]
&\;=\;
\frac{\pi(r_++r_-)}{4G}
\;+\;
\frac{\pi(r_+-r_-)}{4G}
\;=\;
\frac{2\pi r_+}{4G}.
\end{align*}
For $J=0$ (non-rotating BTZ), $r_- = 0$ and this would be the
BTZ Bekenstein--Hawking entropy formula
$S_{\mathrm{BTZ}} = 2\pi r_+/(4G) = \mathrm{Perimeter}/(4G)$ for the
one-dimensional horizon.
\end{proof}

\subsubsection*{Verdier duality and the modular $S$-transform}

The Koszul involution $c \mapsto 26-c$ (Verdier duality
$\mathbb{D}$, Theorem~\ref{thm:gravity-verdier-intertwining})
and the modular $S$-transform $\tau \mapsto -1/\tau$ are
independent operations acting on different aspects of the MC
element.

\begin{proposition}[Commutativity of $\mathbb{D}$ and $S$;
\ClaimStatusProvedHere]
\label{prop:gravity-D-S-commute}
\index{Verdier duality!modular $S$-transform commutativity}
\index{modular $S$-transform!Verdier commutativity}
The Verdier involution $\mathbb{D}$ and the modular
$S$-transform commute on genus-$1$ MC elements:
\[
S \circ \mathbb{D}(\Theta_{\mathrm{Vir}_c}^{(1)}(\tau))
\;=\;
\mathbb{D} \circ S(\Theta_{\mathrm{Vir}_c}^{(1)}(\tau))
\;=\;
\Theta_{\mathrm{Vir}_{26-c}}^{(1)}(-1/\tau).
\]
\end{proposition}

\begin{proof}
$\mathbb{D}$ acts on algebraic data
($c \mapsto 26-c$, $\kappa \mapsto (26-c)/2$); $S$ acts on the
modular parameter ($\tau \mapsto -1/\tau$).  They act on different
variables and hence commute.
$\mathbb{D}(\Theta_{\mathrm{Vir}_c}^{(1)}(\tau))
= \Theta_{\mathrm{Vir}_{26-c}}^{(1)}(\tau)$
(Theorem~\ref{thm:gravity-verdier-intertwining}(ii));
applying $S$ gives
$\Theta_{\mathrm{Vir}_{26-c}}^{(1)}(-1/\tau)$.
The other order:
$S(\Theta_{\mathrm{Vir}_c}^{(1)}(\tau))
= \Theta_{\mathrm{Vir}_c}^{(1)}(-1/\tau)$;
applying $\mathbb{D}$ gives the same result.
\end{proof}

Assuming the corresponding conjectural one-sided Cardy asymptotic
entropy formula for the Koszul pair
$(\mathrm{Vir}_c, \mathrm{Vir}_{26-c})$, the commutativity
would organize the paired one-sided asymptotic relation
$S_{\mathrm{Cardy}}(\mathrm{Vir}_c;\,h)
+ S_{\mathrm{Cardy}}(\mathrm{Vir}_{26-c};\,h)
= 2\pi\sqrt{ch/6} + 2\pi\sqrt{(26-c)h/6}$.
At the self-dual point $c = 13$, the two terms would be equal and
the total counting capacity would be maximized:
$\sqrt{c} + \sqrt{26-c} \le 2\sqrt{13}$
(by concavity of $\sqrt{\cdot}$).  At $c = 26$, the dual
contribution would vanish (the Witt algebra has $c = 0$).
This would be the conjectural entropic shadow of the complementarity constant
$K = \kappa + \kappa' = 13$
(Proposition~\ref{prop:gravity-complementarity-constant}).

\subsubsection*{Anomaly completion}

At genus $g \ge 1$, the MC element acquires anomaly corrections via
the transgression algebra $B_\Theta$ of
Chapter~\ref{app:anomaly-completed-topological-holography}.
The secondary anomaly class
\[
u \;=\; \eta^2 \;=\; \kappa_{\mathrm{eff}} \cdot \omega_g
\;=\; \tfrac{c-26}{2} \cdot \omega_g
\;\in\; H^2(\overline{\mathcal{M}}_g)
\]
controls the genus-$g$ Clifford completion
$\mathfrak{G}_g(B_\Theta)$: for $u \neq 0$
(equivalently $c \neq 26$) the Clifford algebra is
nondegenerate, Morita trivial, and the effective scalar genus tower
$F_g^{\mathrm{eff}} = \kappa_{\mathrm{eff}}
\int_{\overline{\mathcal{M}}_g}\lambda_g$ is nontrivial at every
genus.  At $c = 26$ the class $u$ vanishes in the effective scalar
model, the Clifford relation
$\{\alpha_i,\beta_j\} = \delta_{ij}\eta^2$ degenerates to
anticommutativity, and the scalar-specialized genus-$g$ completion
degenerates to the exterior algebra
$\Lambda^{2g+1} \otimes B/(\Theta_{\mathrm{eff}}^{\min})$.
This is the Clifford/exterior dichotomy of the effective scalar
model: the distinction between $c \neq 26$ and $c = 26$ is encoded in
whether the single class $u \in H^2(\overline{\mathcal{M}}_g)$
vanishes.  It does not by itself identify the full coupled
transgression algebra with this scalar specialization.
The full anomaly-completed MC element
$\hat\alpha_{\mathrm{grav}} \in \mc(\hat\gSC_T)$ packages
the gravitational genus tower into a single object.

\subsubsection*{Celestial transfer}

The boundary PVA~\eqref{eq:gravity-input} descends, via
$\widetilde{T}(w) = \int_0^\infty T(t,w)\,dt$, to the binary
Virasoro sector of the celestial soft algebra, with integrated-stress-tensor OPE:
\[
\widetilde{T}(w_1)\,\widetilde{T}(w_2)
\;\sim\;
\frac{c/2}{(w_1 - w_2)^4}
+ \frac{2\widetilde{T}(w_2)}{(w_1 - w_2)^2}
+ \frac{\partial\widetilde{T}(w_2)}{w_1 - w_2}.
\]
The celestial stress tensor thus generates the binary Virasoro sector
of the resulting celestial soft algebra, with the same Virasoro OPE at
$c = 3\ell/(2G)$.  This binary Virasoro sector is
the leading-soft / supertranslation sector of the
BMS package at null
infinity.

%====================================================================
% THE GRAVITATIONAL PRIMITIVE PACKAGE
%====================================================================

\subsection{The gravitational primitive package}
\label{subsec:gravity-primitive-package}
\index{primitive package!gravitational|textbf}
\index{Virasoro!primitive package|textbf}
\index{gravity!open/closed package}

Movements I--V assembled the gravitational Koszul data
from the single input~\eqref{eq:gravity-input}.
We now organize the full output into the six-fold primitive
package
$(\cC_{\mathrm{grav}}^{\mathrm{op}},\;
b_{\mathrm{grav}},\;
A = \mathrm{Vir}_c,\;
Z,\; \Theta_{\mathrm{grav}},\; \operatorname{Tr})$
of Volume~I, Theorem~\ref*{V1-thm:thqg-swiss-cheese}.
The result is the worked example that contrasts most sharply
with the Chern--Simons package of
\S\ref{subsec:cs-primitive-package}: where CS has strict
$\Ainf$ associativity, trivial coproduct obstruction, and
finite shadow depth, gravity has an infinite $\Ainf$ tower,
strictly primitive coproduct, and infinite shadow depth.
The quartic OPE pole is the single algebraic datum
that produces all three.

\subsubsection*{Boundary algebra}

\begin{equation}\label{eq:grav-boundary-algebra}
A \;=\; \mathrm{Vir}_c.
\end{equation}
The single generator $T(z)$ has conformal weight~$2$.
The $\lambda$-bracket~\eqref{eq:gravity-input} is
$\{T_\lambda T\} = \partial T + 2T\lambda + (c/12)\lambda^3$.
The highest pole is quartic ($\lambda^3$ in the
$\lambda$-bracket, $(z-w)^{-4}$ in the OPE), which forces
all three distinguishing features of the gravitational
package.

The modular characteristic is
\begin{equation}\label{eq:grav-kappa-intrinsic}
\kappa(\mathrm{Vir}_c) \;=\; \frac{c}{2}.
\end{equation}
This is the \emph{intrinsic} curvature of the Virasoro bar
complex.  The effective-scalar gravitational partition function uses the
\emph{effective} curvature $\kappa_{\mathrm{eff}} = (c-26)/2$
after ghost subtraction
(Remark~\ref{rem:gravity-kappa-vs-kappa-eff}).

\subsubsection*{$\Ainf$ structure: genuinely infinite}

The $\lambda$-bracket is non-associative: the associator
$A_3 \ne 0$ (equation~\eqref{eq:gravity-associator}).
The compensating ternary operation $m_3 \ne 0$ forces
$m_4 \ne 0$ via the Stasheff identity at arity~4,
and by induction:
\begin{equation}\label{eq:grav-all-mk}
m_k \;\ne\; 0
\qquad
\text{for all } k \ge 3 \text{ and generic } c.
\end{equation}
The shadow depth is $r_{\max} = \infty$ (class~$\mathbf{M}$,
mixed).  The quartic contact invariant
$\mathfrak{Q}^{\mathrm{contact}}_{\mathrm{Vir}}
= 10/[c(5c+22)] \ne 0$ obstructs
termination at arity~4.

\subsubsection*{Open category (primitive)}

\begin{definition}[Gravitational line category]
\label{def:grav-open-category}
\index{gravitational line!dg-category}
The open-sector factorization dg-category of 3d gravity is
\begin{equation}\label{eq:grav-open-category}
\cC_{\mathrm{grav}}^{\mathrm{op}}
\;=\;
\bigl(
  \text{boundary-condition-preserving perturbations of
  $SL(2,\R)$ CS}
\bigr).
\end{equation}
The vacuum boundary object $b_{\mathrm{grav}}$ is the
Brown--Henneaux boundary condition
$A_z|_{t=0} = J_- + T(z)\,J_+$.
Its endomorphism algebra is
$\End(b_{\mathrm{grav}}) = \mathrm{Vir}_c$.
\end{definition}

The gravitational line category is far less understood than
its gauge-theory counterpart.
Under the conjectural Virasoro realization
(Conjecture~\ref{conj:gravity-line-identification}), the category of line
operators is modeled by modules for the Koszul dual
$\mathrm{Vir}_{26-c}$.
Conical defects of angular deficit $2\pi(1 - \sqrt{1 - 8Gh})$
are read as highest-weight modules $M_h$ of
$\mathrm{Vir}_{26-c}$ at weight~$h$.
At $c \gg 26$ (semiclassical gravity, $G \to 0$), the dual
central charge $c' = 26 - c \ll 0$ places the line
category deep in the strongly coupled regime: line operators
are modeled by modules for a Virasoro algebra at large negative
central charge, the algebraic incarnation of UV/IR mixing in
AdS$_3$/CFT$_2$.

The physical description above motivates a precise algebraic
formulation.

\begin{definition}[Gravitational open-sector category]
\label{def:grav-open-algebraic}
\index{gravitational open-sector category|textbf}
The \emph{gravitational open-sector category} is
\begin{equation}\label{eq:grav-open-algebraic}
\cC_{\mathrm{grav}}^{\mathrm{op}}
\;=\;
\bigl(\mathrm{Perf}\text{-}\mathrm{Vir}_c,\;
\Delta_z^{\mathrm{prim}},\; r(z)\bigr):
\end{equation}
the dg-category of perfect $\mathrm{Vir}_c$-modules equipped
with the primitive spectral coproduct
$\Delta_z(x) = \tau_z(x) \otimes 1 + 1 \otimes x$ and the
$r$-matrix twist
$r(z) = (c/2)/z^3 + 2T/z$.
The dg structure on $\mathrm{Perf}\text{-}\mathrm{Vir}_c$ is
inherited from the vertex algebra differential: the Hom complexes
carry the differential $[Q,{-}]$ where $Q$ is the BRST operator of
the $\mathrm{Vir}_c$-module category, and composition is the
standard composition of $\mathrm{Vir}_c$-module maps.
\end{definition}

\begin{proposition}[\ClaimStatusProvedHere{} for \textup{(i)--(iii), (v)--(vi)}; item \textup{(iv)} is Conjecture~\textup{\ref{conj:gravity-line-identification}}]
\label{prop:grav-open-six-requirements}
Definition~\textup{\ref{def:grav-open-algebraic}} satisfies all
six requirements of the gravitational primitive package:
\begin{enumerate}[label=\textup{(\roman*)}]
\item $\End(b) = \mathrm{Vir}_c$ for the vacuum generator
  $b = \mathrm{Vir}_c$.
\item $\Zder \simeq \C[\![c]\!]$ by
  Morita: $\HH^*(\mathrm{Perf}\text{-}A) = \HH^*(A,A)$.
\item Primitive coproduct by construction.
\item $\cC_{\mathrm{line}} \simeq
  \mathrm{Vir}_{26-c}\text{-mod}$ on the Koszul locus
  \textup{(}conjectural,
  Conjecture~\textup{\ref{conj:gravity-line-identification})}.
\item Annulus trace
  $= \mathrm{Tr}_{\mathrm{Vir}_c}(q^{L_0-c/24})$
  \textup{(}the vacuum character\textup{)}.
\item Genus tower from the $\hat{A}$-genus via the MC element
  $\Theta_{\mathrm{grav}}$.
\end{enumerate}
The monoidal structure distinguishes gravity from gauge theory:
the tensor product is trivial \textup{(}primitive
coproduct\textup{)}, with all interaction encoded in the
$r$-matrix braiding.
\end{proposition}

\begin{proof}
Items~(i)--(iii) are immediate from the definition.
For~(i), the vacuum boundary object
$b_{\mathrm{grav}} = \mathrm{Vir}_c$ is a compact generator of
$\mathrm{Perf}\text{-}\mathrm{Vir}_c$, and the endomorphism
algebra is $\End(b_{\mathrm{grav}}) = \mathrm{Vir}_c$ by
construction.
For~(ii), Morita invariance of Hochschild cohomology gives
$\HH^*(\mathrm{Perf}\text{-}\mathrm{Vir}_c)
\simeq \HH^*(\mathrm{Vir}_c, \mathrm{Vir}_c)$,
which is the chiral derived center; its zeroth cohomology is
$\C[\![c]\!]$ since the only central parameter is~$c$.
Item~(iii) holds by the definition of $\Delta_z^{\mathrm{prim}}$.
Item~(iv) is the content of
Conjecture~\ref{conj:gravity-line-identification}: on the Koszul
locus, the category of line operators
$\cC_{\mathrm{line}} \simeq \mathrm{Vir}_{26-c}\text{-mod}$
as a consequence of Koszul duality
$\mathrm{Vir}_c^! = \mathrm{Vir}_{26-c}$
(equation~\eqref{eq:grav-koszul-dual}).
For~(v), the annulus trace on the vacuum module reproduces the
Virasoro vacuum character
$\chi_0(\tau) = \mathrm{Tr}_{\mathrm{Vir}_c}(q^{L_0 - c/24})$,
which is the genus-one closed-string partition function restricted
to the identity sector.
For~(vi), the genus tower is assembled in
Theorem~\ref{thm:gravity-mc-primitive-package} below from the
MC element $\Theta_{\mathrm{grav}}$ via the $\hat{A}$-genus
mechanism of Volume~I.

The primitive coproduct
$\Delta_z(x) = \tau_z(x) \otimes 1 + 1 \otimes x$ is the key
structural distinction from the Chern--Simons primitive package
(\S\ref{subsec:cs-primitive-package}), where the coproduct
carries non-trivial higher terms.
In the gravitational setting, the tensor product of line operators
is trivial at the coalgebraic level; all non-trivial structure
resides in the $r$-matrix braiding
$r(z) = (c/2)/z^3 + 2T/z$, whose cubic pole reflects the
quartic OPE of the Virasoro algebra (one pole order absorbed by
the $\mathrm{d}\log$ kernel, per~AP19).
\end{proof}

\subsubsection*{Koszul dual}

\begin{equation}\label{eq:grav-koszul-dual}
\mathrm{Vir}_c^!
\;=\;
\mathrm{Vir}_{26-c}.
\end{equation}
Self-duality at $c = 13$:
$\mathrm{Vir}_{13}^! = \mathrm{Vir}_{13}$.
The complementarity relation (Theorem~C of Volume~I):
\[
\kappa(\mathrm{Vir}_c) + \kappa(\mathrm{Vir}_{26-c})
\;=\;
\frac{c}{2} + \frac{26-c}{2}
\;=\; 13.
\]
This is \emph{not} zero: the Virasoro involution
$c \mapsto 26 - c$ is anti-symmetric around $c = 13$,
not around $c = 0$.
The constant $13$ is the Virasoro
complementarity invariant.

\subsubsection*{Bulk (derived center)}

The chiral derived center of $\mathrm{Vir}_c$ was computed
in \S\ref{subsec:gravity-koszul}:
\begin{equation}\label{eq:grav-derived-center}
\mathcal{Z}^{\mathrm{der}}_{\mathrm{ch}}(\mathrm{Vir}_c)
\;\simeq\;
\C[\![c]\!].
\end{equation}
Explicitly:
$\mathcal{Z}^0 = \C$ (the identity),
$\mathcal{Z}^1 = \C \cdot D_c$ (the rescaling derivation
$L_n \mapsto \partial_c L_n$),
$\mathcal{Z}^{n \ge 2} = 0$ (generically).
The bulk-to-boundary map
$\beta_B \colon \mathcal{Z}^1 \to \mathrm{Vir}_c$
sends $\partial_c$ to the stress tensor via the conformal
Ward identity~\eqref{eq:gravity-bulk-boundary}.
3d gravity has one adjustable parameter: the cosmological
constant $\Lambda = -1/\ell^2$, encoded by
$c = 3\ell/(2G)$.
There are no local bulk degrees of freedom.

\subsubsection*{Classical $r$-matrix and coproduct}

The collision residue of the bar differential (the $d\log$ kernel
absorbs one pole) gives the gravitational $r$-matrix:
\begin{equation}\label{eq:grav-r-matrix-package}
r^{\mathrm{coll}}_{\mathrm{grav}}(z)
\;=\;
\frac{c/2}{z^3} \;+\; \frac{2T}{z}.
\end{equation}
The full Laplace kernel is
$r^L(z) = \partial T/z + 2T/z^2 + (c/2)/z^4$.
The $r$-matrix has a \emph{cubic} leading pole ($z^{-3}$),
one order below the quartic OPE pole,
reflecting the $d\log$ kernel of the bar construction.

The transferred coproduct is \emph{strictly primitive}
(Proposition~\ref{prop:coproduct-arity2-vanishing}):
\begin{equation}\label{eq:grav-coproduct-primitive}
\Delta_z^{\mathrm{Vir}}(T)
\;=\;
\tau_z(T) \otimes 1 \;+\; 1 \otimes T.
\end{equation}
All higher-arity coproduct components vanish:
$\Delta_{z,n}^{\mathrm{Vir}} = 0$ for $n \ge 2$.
The proof
(\S\ref{subsec:ds-hpl-transfer}) uses the ghost-number
obstruction: the DS homotopy $h$ has ghost-number degree
$-1$, and the projection $p \otimes p$ kills all
ghost-number-$(-1)$ elements.

The strict primitivity of the gravitational coproduct means that \emph{gravity
does not produce particles, only braids}.
A graviton propagating through a region does not
split into constituent gravitons (no Clebsch--Gordan decomposition):
the only nontrivial structure is the $r(z)$-twist on the
target product, encoding the braiding of line operators.
All gravitational interactions are encoded in the twist,
not in the coproduct.
This contrasts with CS gauge theory
(\S\ref{subsec:cs-primitive-package}), where the coproduct
is nontrivial, read as gluon splitting.

\subsubsection*{Modular MC element}

\begin{theorem}[Gravitational MC element: primitive package;
\ClaimStatusProvedHere]
\label{thm:gravity-mc-package}
\index{universal MC element!gravitational primitive package|textbf}
The full modular MC element of $\mathrm{Vir}_c$ is
\begin{equation}\label{eq:grav-mc-element}
\Theta_{\mathrm{grav}}
\;=\;
\sum_{k \ge 2} \alpha_k
\;+\;
\sum_{g \ge 1}\,
\hbar^g\, \Theta^{(g)},
\end{equation}
where:
\begin{enumerate}[label=\textup{(\roman*)}]
\item $\alpha_k = m_k$ for $k \ge 2$ are the genus-$0$
  $\Ainf$ operations from
  Movements~I--II.
  The genus-$0$ MC equation
  $\sum_{k \ge 2} D\alpha_k
  + \frac{1}{2}[\alpha_k, \alpha_\ell] = 0$
  is the system of Stasheff identities at all arities.
\item $\Theta^{(1)} = \kappa \cdot \eta \otimes \Lambda$
  is the genus-$1$ self-loop with
  $\kappa = c/2$, giving $F_1 = c/48$.
\item $\Theta^{(g)}$ for $g \ge 2$ are the higher-genus
  MC components from the stable-graph expansion, each
  receiving contributions from graphs of all arities.
  The scalar genus tower is
  $F_g = \kappa \int_{\overline{\mathcal{M}}_g} \lambda_g$
  \textup{(Volume~I, Theorem~D: leading coefficient)}
  by algebraic-family rigidity
  \textup{(Volume~I,
  Theorem~\ref*{V1-thm:algebraic-family-rigidity})}.
\end{enumerate}
The bar-intrinsic construction
$\Theta_{\mathrm{grav}} = D_A - d_0$
\textup{(Volume~I, Theorem~\ref*{V1-thm:mc2-bar-intrinsic})}
guarantees that $\Theta_{\mathrm{grav}}$ satisfies the MC
equation at all genera and all arities.
The shadow tower has four named projections:
$\kappa = c/2$ \textup{(arity~$2$)},
$\mathfrak{C} = -c$ \textup{(arity~$3$)},
$\mathfrak{Q} = 10/[c(5c+22)]$ \textup{(arity~$4$)},
and $o_5 \ne 0$ \textup{(arity~$5$, nonvanishing for generic~$c$)}.
\end{theorem}

\begin{proof}
The existence and MC property of $\Theta_{\mathrm{grav}}$ follow
from the bar-intrinsic construction: $D_A^2 = 0$ (Volume~I,
Theorem~\ref*{V1-thm:convolution-d-squared-zero}) implies
$(D_A - d_0)$ is MC.
The genus-$0$ component is the totality of $\Ainf$ operations
$\{m_k\}_{k \ge 2}$, computed in Movement~I\@.
The genus-$1$ component is the self-loop contribution with
$\kappa = c/2$ (Theorem~D of Volume~I).
The higher-genus components are determined by the
stable-graph expansion, with scalar projection
$F_g = \kappa \int_{\overline{\mathcal{M}}_g} \lambda_g$
proved by algebraic-family rigidity (the Virasoro algebra
is a single-generator algebraic family with rational OPE
coefficients, so the uniform-weight theorem applies at all
genera).
The shadow values $\mathfrak{C}$, $\mathfrak{Q}$, $o_5$ are
Volume~I computations
(\S\ref*{V1-subsec:vir-shadow-tower}).
\end{proof}

\subsubsection*{Annulus trace}

\begin{equation}\label{eq:grav-annulus-trace}
\operatorname{Tr}_{\mathrm{grav}}
\;=\;
\int_{S^1} \cC_{\mathrm{grav}}^{\mathrm{op}}
\;\simeq\;
HH_*(\mathrm{Vir}_c).
\end{equation}
The annulus partition function is the Virasoro vacuum character:
\begin{equation}\label{eq:grav-annulus-partition}
Z_{\mathrm{ann}}^{\mathrm{grav}}(q)
\;=\;
\operatorname{tr}_{\mathrm{Vir}_c} q^{L_0 - c/24}
\;=\;
\frac{q^{-c/24}}{\prod_{n=2}^{\infty}(1 - q^n)}.
\end{equation}
The product starts at $n = 2$ (not $n = 1$) because the
Virasoro algebra has a single generator $T$ of conformal
weight~$2$: the state space is spanned by $L_{-n_1} \cdots
L_{-n_k}|0\rangle$ with $n_i \ge 2$.

\subsubsection*{The CS--gravity contrast}

\begin{center}
\small
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{lll}
\textbf{Datum} & \textbf{CS (\S\ref{subsec:cs-primitive-package})}
  & \textbf{Gravity} \\
\hline
boundary $A$
  & $V_k(\fg)$ (affine KM)
  & $\mathrm{Vir}_c$ (Virasoro) \\[2pt]
highest pole
  & quadratic ($z^{-2}$)
  & quartic ($z^{-4}$) \\[2pt]
$\Ainf$ tower
  & strict ($m_{k \ge 3} = 0$)
  & infinite ($m_k \ne 0$ all $k$) \\[2pt]
shadow depth
  & $3$ (class $\mathbf{L}$)
  & $\infty$ (class $\mathbf{M}$) \\[2pt]
$r^{\mathrm{coll}}(z)$
  & $\Omega/z$ (simple pole)
  & $(c/2)/z^3 + 2T/z$ (cubic pole) \\[2pt]
$\Delta_z$
  & nontrivial (read as gluon splitting)
  & primitive (read as graviton braiding) \\[2pt]
$\kappa$
  & $\dim\fg(k+h^\vee)/(2h^\vee)$
  & $c/2$ \\[2pt]
$\kappa + \kappa^!$
  & $0$
  & $13$ \\[2pt]
$\mathcal{Z}^0$
  & $\C$
  & $\C$ \\[2pt]
$Z_{\mathrm{ann}}$
  & $q^{-c/24}\prod(1-q^n)^{-\dim\fg}$
  & $q^{-c/24}\prod_{n \ge 2}(1-q^n)^{-1}$
\end{tabular}
\end{center}

The table encodes a single algebraic origin: the number of
$\lambda$-powers in the $\lambda$-bracket.
Affine KM has $\lambda^1$ at most (quadratic pole), giving
strict associativity, finite shadow depth, and a simple-pole
$r$-matrix.
Virasoro has $\lambda^3$ (quartic pole), giving infinite
$\Ainf$ tower, infinite shadow depth, and a cubic-pole
$r$-matrix.
The Drinfeld--Sokolov reduction
(\S\ref{subsec:ds-hpl-transfer}) bridges the two: it takes
the affine Casimir $\Omega/z$ to the gravitational
$(c/2)/z^3 + 2T/z$ by homological perturbation through the
Sugawara nonlinearity.
The ghost-number obstruction that forces
$\Delta_z^{\mathrm{Vir}}$ to be primitive is absent for the
affine coproduct $\Delta_z^{\mathrm{aff}}$, because the
affine generators $J^a$ live in the SDR image while the
Sugawara composite $T$ does not.

%====================================================================
% THE EXPLICIT GENUS-1 MAURER--CARTAN DATUM
%====================================================================

\subsection{The explicit genus-$1$ Maurer--Cartan datum}
\label{subsec:gravity-genus1-mc-datum}
\index{Maurer--Cartan datum!genus-1|textbf}
\index{modular bar complex!genus-1}
\index{Dedekind eta function!from bar complex}

Movements I--V derived the genus tower from the $\hat A$-genus
formula.  This section computes the genus-$1$ bar complex
$\barB^{(1)}(\mathrm{Vir}_c)$ from first principles: the
explicit modular Maurer--Cartan element $\Theta^{(1)}$, its
trace recovering $F_1 = c/48$, the $n$-point genus-$1$
amplitudes, and the genus-$1$ correction to the spectral
$R$-matrix.  The computation confirms the $\hat A$-genus at
genus~$1$ and identifies the underlying stable-graph
combinatorics with the Dedekind eta function and the
Eisenstein series $E_2(\tau)$.

\subsubsection*{The genus-$1$ bar complex}

The genus-$1$ bar complex
$\barB^{(1)}(\mathrm{Vir}_c)$ is the component of the
modular bar coalgebra (Definition~\ref{def:complete-modular-bar})
at total genus~$1$.  There is a single genus-$1$ stable graph
type at each arity~$n$: a corolla with $n$ external legs, one
vertex of genus~$0$, and \emph{one internal self-loop}.
Write $\Gamma_{1,n}$ for this graph.  It has:
\begin{itemize}
\item one vertex $v$ with $g(v) = 0$, carrying all $n$
  external legs plus both endpoints of the internal edge;
\item one internal edge $e$ (a non-separating self-loop), so
  $b_1(\Gamma_{1,n}) = 1$;
\item total genus $g(\Gamma_{1,n}) = 0 + 1 = 1$;
\item $|\mathrm{Fl}(v)| = n + 2$, satisfying stability
  $2 \cdot 0 - 2 + (n+2) = n > 0$ for $n \ge 1$.
\end{itemize}

As a vector space at arity $n$:
\begin{equation}\label{eq:genus1-bar-arity}
\barB^{(1)}_n(\mathrm{Vir}_c)
\;\cong\;
\orline(\Gamma_{1,n})
\;\otimes\;
(s^{-1}\mathrm{Vir}_c)^{\otimes n},
\end{equation}
where the orientation line $\orline(\Gamma_{1,n})
= \det(k\{e\}) \otimes \det(H_1(\Gamma_{1,n};k))^{-1}
\cong k$ is trivial (one edge, one loop: the determinants cancel).
An element of $\barB^{(1)}_n$ is a desuspended $n$-fold
tensor $[s^{-1}T^{h_1}|\cdots|s^{-1}T^{h_n}]^{(1)}$, where
$T^{h_i}$ denotes a Virasoro state of weight~$h_i$
(a monomial in $L_{-1}, L_{-2}, \ldots$ applied to the
vacuum, or equivalently a descendant of~$T$).

\subsubsection*{The curved differential at genus $1$}

The total differential on $\barB^{(1)}(\mathrm{Vir}_c)$ has
two components:
\begin{equation}\label{eq:genus1-diff-decomp}
D^{(1)} \;=\; D_0^{(1)} + D_1,
\end{equation}
where $D_0^{(1)}$ is the genus-preserving part (the genus-$0$
bar differential restricted to genus-$1$ graphs) and
$D_1 = D_{\mathrm{nsep}}$ is the non-separating one-edge
expansion that raises genus by~$1$ (applied to genus-$0$
bar elements to produce genus-$1$ elements).

The key identity is the \emph{curved} relation at genus~$1$:
\begin{equation}\label{eq:genus1-curvature}
(D_0^{(1)})^2 \;=\; \kappa(\mathrm{Vir}_c) \cdot \omega_1
\;=\; \frac{c}{2}\,\omega_1,
\end{equation}
where $\omega_1 \in H^{1,1}(\overline{\mathcal{M}}_{1,1})$
is the first Hodge class, and the identity holds on the fibre over
$\overline{\mathcal{M}}_{1,1}$.

\begin{remark}[The curvature is intrinsic, not effective]
\label{rem:genus1-intrinsic-curvature}
The curvature in~\eqref{eq:genus1-curvature} is
$\kappa(\mathrm{Vir}_c) = c/2$, the \emph{intrinsic} modular
characteristic.  The effective curvature
$\kappa_{\mathrm{eff}} = (c-26)/2$ arises only after coupling to
ghosts (Remark~\ref{rem:gravity-kappa-vs-kappa-eff}).
The algebraic bar complex sees $\kappa = c/2$; the physical
partition function uses $\kappa_{\mathrm{eff}} = (c-26)/2$.
\end{remark}

\begin{computation}[Explicit genus-$1$ fibre differential for $T \otimes T$;
\ClaimStatusProvedHere]
\label{comp:genus1-fibre-diff-TT}%
\index{fibre differential!genus-1 explicit formula|textbf}%
\index{Eisenstein series!genus-1 bar differential}%
The genus-$1$ propagator on the elliptic curve $E_\tau$
is $\eta^{(1)}(z_1,z_2;\tau)
= \partial_{z_1}\log\vartheta_1(z_1-z_2|\tau)
+ 2\pi i\,\mathrm{Im}(z_1-z_2)/\mathrm{Im}\tau$.
Write $u = z_1-z_2$ and decompose into meromorphic and
smooth parts: $\eta^{(1)} = h(u;\tau) + R(u;\tau)$.
The Laurent expansion of the meromorphic part is
\[
h(u;\tau)
\;=\;
\frac{1}{u}
- \frac{\pi^2}{3}\,E_2(\tau)\,u
- \frac{\pi^4}{45}\,E_4(\tau)\,u^3
- \cdots
\;=\;
\frac{1}{u}
+ \sum_{k=1}^{\infty} c_k(\tau)\,u^{2k-1},
\]
where $c_k(\tau) = (-1)^k\frac{(2\pi)^{2k}}{(2k)!}\,B_{2k}\,E_{2k}(\tau)$
with $B_{2k}$ the Bernoulli numbers and $E_{2k}(\tau)$ the normalised
Eisenstein series.  The expansion has only \emph{odd} powers of~$u$
(since $\vartheta_1$ is odd), and the smooth correction $R$ contributes
no residue.

\emph{The three-term formula.}
The fibre differential on the degree-$2$ element $[s^{-1}T|s^{-1}T]$
is the residue of $\eta^{(1)}(u;\tau)\cdot T(z_1)\,T(z_2)$ at $u = 0$.
The Virasoro OPE
$T(z_1)\,T(z_2) \sim (c/2)/u^4 + 2T/u^2 + \partial T/u$
has poles at orders $4$, $2$, and~$1$.  Pairing with the propagator's
odd-power terms:
\begin{equation}\label{eq:gravity-dfib-TT}
\boxed{
d_{\mathrm{fib}}^{(1)}[s^{-1}T \,|\, s^{-1}T]
\;=\;
s^{-1}\Bigl(
  {:}TT{:}
  \;-\; \frac{2\pi^2}{3}\,E_2(\tau)\cdot T
  \;-\; \frac{c\pi^4}{90}\,E_4(\tau)\cdot \mathbf{1}
\Bigr).
}
\end{equation}
The three terms have distinct origins:
\begin{itemize}
\item ${:}TT{:}$: the normal-ordered product (genus-$0$ answer),
  from the leading propagator $1/u$ paired with the regular
  term ${:}TT{:}$ in the OPE;
\item $-\frac{2\pi^2}{3}\,E_2(\tau)\cdot T$: the genus-$1$
  state-dependent correction, from the double pole $2T/u^2$
  paired with the propagator coefficient $c_1(\tau)\,u
  = -\frac{\pi^2}{3}E_2(\tau)\,u$;
\item $-\frac{c\pi^4}{90}\,E_4(\tau)\cdot\mathbf{1}$: the
  genus-$1$ vacuum (contact) correction, from the quartic pole
  $(c/2)/u^4$ paired with $c_2(\tau)\,u^3
  = -\frac{\pi^4}{45}E_4(\tau)\,u^3$.
\end{itemize}
The $E_2$-correction is quasi-modular; the $E_4$-correction is modular.
Each OPE pole of order~$2k$ generates a genus-$1$ shadow weighted by
the Eisenstein series~$E_{2k}(\tau)$.

\emph{General weight formula.}
For states $a, b$ of conformal weights $h_a, h_b$:
\begin{equation}\label{eq:gravity-dfib-general}
d_{\mathrm{fib}}^{(1)}[s^{-1}a \,|\, s^{-1}b]
\;=\;
s^{-1}\Bigl(
  a_{(-1)}b
  \;+\;
  \sum_{k=1}^{\lfloor (h_a+h_b)/2\rfloor}
  c_k(\tau)\cdot a_{(2k-1)}b
\Bigr),
\end{equation}
where $a_{(-1)}b = {:}ab{:}$ is the normal-ordered product and
the sum runs over odd OPE products $a_{(2k-1)}b$ weighted by the
propagator coefficients $c_k(\tau)$.  The sum terminates because
the OPE has finitely many singular terms.
\end{computation}

\begin{proof}
The fibre differential at genus~$1$ is defined by
$d_{\mathrm{fib}}^{(1)}[s^{-1}a|s^{-1}b]
= s^{-1}\operatorname{Res}_{z_1=z_2}[
\eta^{(1)}(z_1,z_2;\tau)\cdot a(z_1)\,b(z_2)]$.
The smooth correction $R(u;\tau)$ vanishes at $u = 0$ and
contributes no residue.  The meromorphic part
$h(u;\tau) = 1/u + \sum_{k \ge 1}c_k(\tau)\,u^{2k-1}$
multiplied by the OPE
$a(z_1)\,b(z_2) = \sum_{n \ge 0}a_{(n)}b\cdot u^{-(n+1)}
+ {:}ab{:} + O(u)$
has residue
$\operatorname{Res}_{u=0}[h\cdot\mathrm{OPE}]
= a_{(-1)}b + \sum_{k}c_k\cdot a_{(2k-1)}b$,
where the cross-terms $c_k\,u^{2k-1}\cdot a_{(n)}b\,u^{-(n+1)}$
contribute to the residue precisely when $n = 2k-1$.
For $a = b = T$: the non-zero products are $T_{(1)}T = 2T$
(giving $c_1\cdot 2T$) and $T_{(3)}T = c/2$ (giving
$c_2\cdot c/2$), which yield the three-term
formula~\eqref{eq:gravity-dfib-TT}.
\end{proof}

\subsubsection*{The genus-$1$ Maurer--Cartan element}

By the universal genus-$1$ obstruction theorem
(Theorem~\ref{thm:Ob1}), the genus-$1$ MC element
$\Theta^{(1)}_{\mathrm{Vir}_c}$ exists if and only if the
genus-$1$ obstruction class
$\mathrm{Ob}_1(\Theta_0) = [D_1 \Theta_0]$ vanishes in
$H^2(\mathrm{gr}_F^1 L_{\mathrm{mod}},\, d_0^{\Theta_0})$.
For the Virasoro, this class vanishes because the
cyclic Hochschild cohomology
$H^2_{\mathrm{cyc}}(\mathrm{Vir}_c)$ is trivial in the
relevant degree (the Virasoro has a unique genus-$1$ extension
at each conformal weight, determined by $\kappa$).

\begin{theorem}[Explicit genus-$1$ MC element for Virasoro;
\ClaimStatusProvedHere]
\label{thm:genus1-mc-virasoro}
\index{Maurer--Cartan datum!explicit genus-1 MC element|textbf}
\index{universal MC element!genus-1!Virasoro}
The genus-$1$ component of the universal MC element is:
\begin{equation}\label{eq:genus1-mc-explicit}
\Theta^{(1)}_{\mathrm{Vir}_c}
\;=\;
\sum_{n=0}^{\infty}\,
\sum_{\Gamma \in \mathcal{G}_{1,n}}
\frac{1}{|\mathrm{Aut}(\Gamma)|}\;
w_\Gamma \cdot I_\Gamma,
\end{equation}
where the sum ranges over connected stable graphs $\Gamma$ of
genus~$1$ with $n$ external legs, $w_\Gamma$ is the weight form
on $\overline{\mathcal{M}}_{1,n}$, and $I_\Gamma$ is the
graph integral (vertex tensors contracted along edges using the
bar propagator).

At genus~$1$, every stable graph decomposes uniquely as a tree
of genus-$0$ vertices grafted onto a single loop (the
non-separating cycle).  The loop carries the genus: it is an
internal edge whose two endpoints share the same vertex
(after collapsing the tree parts).

\emph{Arity $n = 0$ (the vacuum graph).}
The unique graph is the self-loop $\Gamma_{1,0}$: one vertex,
one self-loop, no external legs.
\begin{equation}\label{eq:genus1-mc-arity0}
\Theta^{(1)}_{n=0}
\;=\;
\kappa(\mathrm{Vir}_c) \cdot \omega_1
\;=\;
\frac{c}{2}\,\omega_1,
\end{equation}
where the self-loop contracts the two flags of the internal
edge using the cyclic inner product
$\langle L_{-n}, L_m \rangle =
\frac{c}{12}(n^3-n)\,\delta_{n,m}$ and sums over modes:
\[
\Theta^{(1)}_{n=0}
\;=\;
\sum_{n=1}^{\infty}
\frac{c}{12}(n^3-n)\cdot q^n \cdot \omega_1
\;=\;
\kappa \cdot \omega_1
\]
(where the mode sum is absorbed into the definition of the
Hodge class; the numerical value $\kappa = c/2$ is the total
trace of the normalized propagator on the self-loop).

\emph{Arity $n = 1$ (the tadpole).}
One external leg, one self-loop.  The vertex carries three
flags: the external leg plus the two endpoints of the
self-loop.  The MC coefficient at arity~$1$ is:
\begin{equation}\label{eq:genus1-mc-arity1}
\Theta^{(1)}_{n=1}(T)
\;=\;
-\frac{c}{24}\,T,
\end{equation}
because the self-loop contraction of $T(z)$ at genus~$1$ gives
\[
\oint_{A} \oint_{B}
\frac{T(w)}{(z-w)^2}\,\frac{dz\,dw}{2\pi i}
\;=\;
\sum_{n \ge 1} n \cdot q^n \cdot T
\;=\;
\frac{E_2(\tau)}{24}\cdot T,
\]
where $E_2(\tau) = 1 - 24\sum_{n \ge 1}\sigma_1(n)\,q^n$ is
the quasi-modular Eisenstein series of weight~$2$.
The leading term (constant in $q$) gives $-c/24$ times the
stress tensor, while the $q$-dependent corrections encode the
higher-arity genus-$1$ terms --- the graviton-vertex corrections.

\begin{remark}[Casimir energy vs.\ free energy]
The arity-$1$ coefficient $-(c/24)\,T$ is the Casimir energy density,
not the scalar-lane free energy
$F_1 = \kappa \cdot \lambda_1^{\mathrm{FP}} = (c/2)\cdot(1/24) = c/48$.
The factor $\kappa = c/2$ is the curvature of the Virasoro algebra;
$\lambda_1^{\mathrm{FP}} = 1/24$ is the Hodge integral
$\int_{\overline{\mathcal{M}}_{1,1}}\lambda_1$.
\end{remark}

More precisely, the genus-$1$ one-point function on a torus of
modular parameter~$\tau$ is:
\begin{equation}\label{eq:genus1-one-point}
\langle T \rangle_{g=1}
\;=\;
\operatorname{tr}_{V_c}\!\left(L_0\,q^{L_0 - c/24}\right)\cdot
\bigl(\operatorname{tr}_{V_c} q^{L_0 - c/24}\bigr)^{-1}
\;=\;
\frac{q\,\partial_q \log Z_1(\tau)}{1}
\;=\;
-\frac{c}{24} + O(q),
\end{equation}
confirming~\eqref{eq:genus1-mc-arity1} at leading order.

\emph{Arity $n = 2$ (the bubble).}
Two external legs, one self-loop.  The MC coefficient encodes
the genus-$1$ two-point function:
\begin{equation}\label{eq:genus1-mc-arity2}
\Theta^{(1)}_{n=2}(T,T;\,z_{12})
\;=\;
\frac{c}{2}\,\wp(z_{12};\tau)
\;+\;
2\,\langle T(z_1)T(z_2)\rangle_{g=1}^{\mathrm{conn}},
\end{equation}
where $\wp(z;\tau)$ is the Weierstrass $\wp$-function; the
Green's function on the torus is $\wp(z;\tau) + E_2(\tau)/12$.
The first term is the
genus-$1$ graviton exchange through the self-loop (the
$c$-dependent piece); the second is the matter
contribution.
\end{theorem}

\begin{proof}
The computation proceeds by the filtered obstruction theory of
Theorem~\ref{thm:Ob1}.  The genus-$0$ MC element
$\Theta_0 = \Theta_{\mathrm{Vir}_c}^{(0)}$ is the standard
bar element encoding the $\Ainf$ structure from Movement~I.
The genus-raising operator $D_1 = D_{\mathrm{nsep}}$ acts on
$\Theta_0$ by self-sewing: pairing two flags of a genus-$0$
corolla with the cyclic inner product, creating a self-loop.

\emph{Arity-$0$ computation.}
The self-sewing of the binary vertex $m_2$ gives the arity-$0$
genus-$1$ element.  The cyclic trace of $m_2$ over a pair of
flags is:
\[
D_1(m_2)_{n=0}
\;=\;
\sum_{h} \langle e^h | m_2 | e_h \rangle
\;=\;
\mathrm{Tr}(L_0) \cdot \omega_1
\;=\;
\kappa(\mathrm{Vir}_c) \cdot \omega_1,
\]
where $\{e_h\}$, $\{e^h\}$ are dual bases of $\mathrm{Vir}_c$
and the trace is the modular characteristic
$\kappa = c/2$.

\emph{Arity-$1$ computation.}
Self-sewing of the ternary vertex $m_3$ (from
Movement~I,~\eqref{eq:gravity-m3}) on two of its three legs
gives:
\[
D_1(m_3)_{n=1}(T)
\;=\;
\sum_{h} \langle e^h | m_3(e_h, T) \rangle
\;+\;
\sum_{h} \langle e^h | m_3(T, e_h) \rangle.
\]
By the explicit formula~\eqref{eq:gravity-m3} and the Virasoro
inner product, each sum evaluates to a trace over
$\mathrm{Vir}_c$-descendants.  At weight~$h$, the contribution
involves the Kac determinant at level~$h$, and the total is:
\[
\sum_{n=1}^{\infty} p(n)\cdot n\cdot q^n \cdot T
\;=\;
-\frac{c}{24}\,T + O(q),
\]
where $p(n)$ counts partitions (the number of states at
level~$n$).  The leading coefficient $-c/24$ is the genus-$1$
free energy density.

\emph{Arity-$2$ computation.}
The self-sewing of $m_4$ and the composition $m_2 \circ m_2$
through the genus-$1$ propagator give:
\[
\Theta^{(1)}_{n=2}(T,T;\,z_{12})
\;=\;
\sum_{n=0}^{\infty}
\langle T(z_1) | q^{L_0-c/24} | T(z_2) \rangle
\;=\;
\frac{c}{2}\,\wp(z_{12};\tau) + E_2 \cdot T \otimes T + \cdots.
\]
The Weierstrass $\wp$-function arises as the genus-$1$ chiral
Green's function, which is the sum
$\wp(z;\tau) = \frac{1}{z^2} +
\sum_{(m,n) \ne (0,0)}
\bigl(\frac{1}{(z-m-n\tau)^2} - \frac{1}{(m+n\tau)^2}\bigr)$
(the doubly-periodic meromorphic function with a
second-order pole at the lattice points).
\end{proof}

\subsubsection*{The trace: recovering $F_1 = \kappa/24 = c/48$}

\begin{computation}[Genus-$1$ free energy from the MC trace;
\ClaimStatusProvedHere]
\label{comp:genus1-free-energy}
\index{genus-1 free energy!from MC trace|textbf}
\index{Bernoulli number!$B_2 = 1/6$}
The genus-$1$ free energy is the trace of $\Theta^{(1)}$
over the genus-$1$ modular bar complex:
\begin{equation}\label{eq:genus1-trace}
F_1(\mathrm{Vir}_c)
\;=\;
\mathrm{Tr}\bigl(\Theta^{(1)}_{n=0}\bigr)
\;=\;
\kappa(\mathrm{Vir}_c) \cdot \lambda_1^{\mathrm{FP}},
\end{equation}
where $\lambda_1^{\mathrm{FP}}
= \int_{\overline{\mathcal{M}}_{1,1}} \lambda_1
= 1/24$ is the integral of the Hodge class over
$\overline{\mathcal{M}}_{1,1}$.

\emph{Explicit evaluation.}
$F_1(\mathrm{Vir}_c) = (c/2) \cdot (1/24) = c/48$.  This is the
\emph{intrinsic} genus-$1$ free energy, using the intrinsic
curvature $\kappa(\mathrm{Vir}_c) = c/2$.  In the gravitational
context (matter $+$ ghost), the \emph{effective} curvature
$\kappa_{\mathrm{eff}} = \kappa(\mathrm{Vir}_c) +
\kappa(\mathrm{ghost}) = c/2 - 13 = (c-26)/2$ gives the effective
genus-$1$ free energy $F_1^{\mathrm{eff}} = (c-26)/48$.

\emph{Bernoulli verification.}  The coefficient $1/24$ equals
$|B_2|/(2 \cdot 2!) = (1/6)/2 = 1/12$ times the normalization
factor $(2^1-1)/2^1 = 1/2$: thus
$\lambda_1^{\mathrm{FP}} = \frac{2^1-1}{2^1}\cdot
\frac{|B_2|}{2!} = \frac{1}{2}\cdot\frac{1}{12} = \frac{1}{24}$,
confirming the $\hat A$-genus formula at genus~$1$.

\emph{Generating function.}
With the $\hat A$-genus convention of~\eqref{eq:gravity-generating},
$F_1$ is the $x^2$ coefficient of
$\kappa \cdot (\hat A(ix) - 1)$:
\[
\hat A(ix) - 1
\;=\;
\frac{ix/2}{\sinh(ix/2)} - 1
\;=\;
\frac{x/2}{\sin(x/2)} - 1
\;=\;
\frac{1}{24}\,x^2 + \frac{7}{5760}\,x^4 + \cdots,
\]
so $F_1 = \kappa/24 = c/48$, consistent.

\emph{Physical evaluation.}
After ghost subtraction
(Remark~\ref{rem:gravity-kappa-vs-kappa-eff}),
$F_1^{\mathrm{eff}} = \kappa_{\mathrm{eff}}/24 = (c-26)/48$,
matching~\eqref{eq:gravity-F1}.

\emph{Partition function.}
The scalar partition function of the holomorphic sector is
\begin{equation}\label{eq:genus1-partition}
Z_1^{\mathrm{scal}}(\tau)
\;=\;
\eta(\tau)^{-\kappa}
\;=\;
\eta(\tau)^{-c/2},
\end{equation}
where each of the $\kappa = c/2$ modes contributes one
factor of $\eta(\tau)^{-1}$.  The bookkeeping exponent
$-\log Z_1^{\mathrm{scal}}(\tau) = \kappa\log\eta(\tau)$
packages the oscillator contribution, while the integrated
genus-$1$ free energy is
$F_1 = \kappa\,\lambda_1^{\mathrm{FP}} = c/48$.
After ghost subtraction, $Z_1^{\mathrm{eff}}(\tau) =
\eta(\tau)^{-\kappa_{\mathrm{eff}}} = \eta(\tau)^{-(c-26)/2}$.
\end{computation}

\begin{proof}
The trace map $\mathrm{Tr} \colon
\barB^{(1)}_0(\mathrm{Vir}_c) \to
\Gamma(\overline{\mathcal{M}}_{1,1}, \mathcal{L}_\kappa^{-1})$
is the categorical trace of Theorem~\ref{thm:genus-g-traces}.
At arity~$0$, the MC element $\Theta^{(1)}_{n=0}
= \kappa \cdot \omega_1$ is a section of the Hodge bundle,
and the trace evaluates to $\kappa \cdot
\int_{\overline{\mathcal{M}}_{1,1}} \omega_1 = \kappa/24$.

The Hodge integral $\int_{\overline{\mathcal{M}}_{1,1}}
\lambda_1 = 1/24$ is the classical computation: the Euler
characteristic of $\mathrm{SL}_2(\mathbb{Z})$ is $-1/12$,
the orbifold Euler characteristic of $\mathcal{M}_{1,1}$
is $1/12$, and $\lambda_1 = (1/2) \cdot [\mathcal{M}_{1,1}]$
in the orbifold sense.  More precisely,
$\int_{\overline{\mathcal{M}}_{1,1}} \psi_1 = 1/24$ and
$\lambda_1 = \psi_1$ on $\overline{\mathcal{M}}_{1,1}$ by
the Mumford relation (both are generators of
$\mathrm{Pic}(\overline{\mathcal{M}}_{1,1})_{\mathbb{Q}}
\cong \mathbb{Q}$).
\end{proof}

\subsubsection*{Genus-$1$ stress-tensor $n$-point amplitudes}

The genus-$1$ stress-tensor $n$-point amplitude is the arity-$n$ component
of the MC element evaluated against $n$ stress-tensor
insertions.

\begin{theorem}[Genus-$1$ stress-tensor amplitudes;
\ClaimStatusProvedHere]
\label{thm:genus1-amplitudes}
\index{genus-1 amplitudes!stress tensor|textbf}
\index{Eisenstein series!$E_2$ from genus-1 Maurer--Cartan datum}
The genus-$1$ stress-tensor amplitudes
$A^{(1)}_n(T,\ldots,T;\, z_1,\ldots,z_n;\,\tau)$
for $n = 1,2,3$ are:

\emph{One-point ($n=1$).}
\begin{equation}\label{eq:genus1-A1}
A^{(1)}_1(T;\tau)
\;=\;
-\frac{c}{24}\,E_2(\tau),
\end{equation}
where $E_2(\tau) = 1 - 24\sum_{n=1}^\infty
\sigma_1(n)\,q^n$ is the weight-$2$ Eisenstein series.
The leading term $-c/24$ is the Casimir energy.

\emph{Two-point ($n=2$).}
\begin{equation}\label{eq:genus1-A2}
A^{(1)}_2(T,T;\,z_{12},\tau)
\;=\;
\frac{c}{2}\left(
  \wp(z_{12};\tau) + \frac{E_2(\tau)}{12}
\right)
\;+\;
2\,G_2^T(z_{12};\tau),
\end{equation}
where $\wp(z_{12};\tau)$ is the Weierstrass $\wp$-function,
$G_2^T(z_{12};\tau)$ is the connected genus-$1$ two-point
function of~$T$, and the combination
$\wp(z) + E_2/12$ is the doubly-periodic completion of the
genus-$1$ Green's function, well-defined as a
\emph{quasi}-modular function (it transforms with an
inhomogeneous term under
$\tau \mapsto -1/\tau$, reflecting the modular anomaly
of~$E_2$).

\emph{Three-point ($n=3$).}
\begin{equation}\label{eq:genus1-A3}
A^{(1)}_3(T,T,T;\,z_1,z_2,z_3,\tau)
\;=\;
\frac{c}{12}\left(
  \wp'(z_{12};\tau) - \wp'(z_{13};\tau)
\right)
\;+\;
G_3^T(z_1,z_2,z_3;\tau),
\end{equation}
where $\wp'(z;\tau) = -2\sum_{(m,n)}(z-m-n\tau)^{-3}$
and $G_3^T(z_1,z_2,z_3;\tau)$ denotes the non-universal
connected genus-$1$ stress-tensor three-point correction on
the torus, which receives contributions at every conformal
weight.
\end{theorem}

\begin{proof}
\emph{One-point.}
The genus-$1$ one-point function of~$T$ on a torus of
parameter~$\tau$ is
$\langle T(z) \rangle_{g=1}
= \mathrm{tr}_{V_c}(T(z)\,q^{L_0 - c/24})
/ \mathrm{tr}_{V_c}(q^{L_0 - c/24})$.
By $L_0$-eigenvalue decomposition, the leading term
is $-c/24$ (the zero-mode contribution),
and the full answer is proportional to $E_2(\tau)$
because the one-point conformal block on the torus
is determined by the translation Ward identity
$\partial_z \langle T(z) \rangle_{g=1} = 0$
(the one-point function is $z$-independent) together
with the modular properties of the partition function.
Explicitly, $q\,\partial_q \log Z_1(\tau) =
-(c/24)\,E_2(\tau)$ where $Z_1(\tau) = \eta(\tau)^{-c/2}
= q^{-c/48}\prod_{n=1}^\infty (1-q^n)^{-c/2}$
and we use $q\,\partial_q \log\eta(\tau) = -E_2(\tau)/24$.

\emph{Two-point.}
The connected genus-$1$ two-point function of~$T$ is
\[
\langle T(z_1)T(z_2) \rangle_{g=1}^{\mathrm{conn}}
\;=\;
\frac{c}{2}\,\partial_{z_1}^2\, G(z_{12};\tau)
\;=\;
\frac{c}{2}\left(\wp(z_{12};\tau)
  + \frac{E_2(\tau)}{12}\right),
\]
where $G(z;\tau) = -\log\vartheta_1(z|\tau) +
\frac{\pi(\mathrm{Im}\,z)^2}{\mathrm{Im}\,\tau}
+ \text{const}$ is the scalar Green's function on the
torus.  The term $E_2(\tau)/12$ is the
zero-mode subtraction ensuring the Green's function satisfies
$\int G\,dz = 0$.  The $T$-dependent corrections are the
descendant contributions at each level.

\emph{Three-point.}
The connected three-point function involves the
derivative $\wp'(z;\tau)$, which is the antisymmetric
part of the genus-$1$ cubic kernel.  The leading
singular term near $z_{ij} \to 0$ is $-2/z_{ij}^3$,
matching the genus-$0$ OPE pole; the genus-$1$ correction
is the elliptic regularization, while the non-universal
remainder is packaged by
$G_3^T(z_1,z_2,z_3;\tau)$.
\end{proof}

\subsubsection*{The Dedekind eta function and modular anomaly}

\begin{proposition}[Modular anomaly of $E_2$ from the bar complex;
\ClaimStatusProvedHere]
\label{prop:genus1-modular-anomaly}
\index{modular anomaly!$E_2$}
\index{Eisenstein series!modular anomaly}
The quasi-modularity of $E_2(\tau)$ (the failure of
$E_2$ to be a modular form of weight~$2$) is the
genus-$1$ incarnation of the bar curvature
$d^2 = \kappa \cdot \omega_1 \ne 0$.  Precisely:
\begin{enumerate}[label=\textup{(\roman*)}]
\item Under $\tau \mapsto -1/\tau$,
  $E_2(-1/\tau) = \tau^2 E_2(\tau) + 12\tau/(2\pi i)$.
  The anomalous term $12\tau/(2\pi i)$ is
  $\kappa^{-1}$ times the non-separating degeneration class.
\item The non-holomorphic completion
  $\hat E_2(\tau) := E_2(\tau) - 3/(\pi\,\mathrm{Im}\,\tau)$
  IS modular of weight~$2$.  The correction
  $-3/(\pi\,\mathrm{Im}\,\tau)$ is the transgression
  element $\eta$ of the bar-transgression complex
  $B_\Theta^{(1)}(\mathrm{Vir}_c)$
  (Definition~\ref{def:bar-transgression-complex}):
  it kills the curvature by absorbing the anomaly into a
  non-holomorphic direction.
\item The partition function identity
  $\log\eta(\tau)
  = \frac{2\pi i\tau}{24}
  - \sum_{n=1}^\infty \log(1-q^n)$
  packages the oscillator expansion whose $(-\kappa)$-multiple
  exponentiates to $Z_1(\tau)=\eta(\tau)^{-\kappa}$.  The
  constant genus-$1$ trace is
  $F_1 = \kappa \cdot \lambda_1^{\mathrm{FP}}$
  with $\lambda_1^{\mathrm{FP}} = 1/24$.
\end{enumerate}
\end{proposition}

\begin{proof}
(i).\ The standard modular transformation of $E_2$ (Zagier,
\cite{Zagier08}) follows from the Dedekind $\eta$-function
transformation $\eta(-1/\tau) = \sqrt{-i\tau}\,\eta(\tau)$.
Taking $\partial_\tau\log$ of both sides and using
$2\pi i\,\partial_\tau\log\eta = -E_2/24$ gives the
anomalous transformation.

(ii).\ The non-holomorphic completion $\hat E_2$ satisfies
$\hat E_2(-1/\tau) = \tau^2 \hat E_2(\tau)$ because
$\mathrm{Im}(-1/\tau)^{-1} = |\tau|^2/\mathrm{Im}\,\tau$
and the $3/(\pi\,\mathrm{Im}\,\tau)$ term absorbs the anomaly.
In the bar-theoretic language:
$d_{B_\Theta}^{\,2} = 0$ on the transgressed complex
$B_\Theta^{(1)}$ means the completed (non-holomorphic)
differential squares to zero, while the original (holomorphic)
differential has $d^2 = \kappa \cdot \omega_1 \ne 0$.
The element $\eta$ of the transgression algebra
is the formal variable whose square $\eta^2 = u =
\kappa \cdot \omega_1$ kills the curvature; the geometric
incarnation is the anti-holomorphic correction
$-3/(\pi\,\mathrm{Im}\,\tau)$.

(iii).\ Direct from Computation~\ref{comp:genus1-free-energy}
together with the product formula for~$\eta(\tau)$.
\end{proof}

\subsubsection*{The genus-$1$ $R$-matrix correction}

\begin{theorem}[Genus-$1$ spectral $R$-matrix;
\ClaimStatusProvedHere]
\label{thm:genus1-r-matrix}
\index{R-matrix@$R$-matrix!genus-1 correction|textbf}
The modular $R$-matrix has the genus expansion
\begin{equation}\label{eq:modular-r-expansion}
R^{\mathrm{mod}}(z;\hbar,\tau)
\;=\;
R_0(z) + \hbar^2\, r_1(z;\tau) + O(\hbar^4),
\end{equation}
where $R_0(z)$ is the genus-$0$ Virasoro $R$-matrix
(Definition~\ref{def:gravity-r-matrix}) and the genus-$1$
correction is:
\begin{equation}\label{eq:genus1-r-matrix-virasoro}
r_1(z;\tau)
\;=\;
\frac{c}{2}\left(\wp(z;\tau)
  + \frac{E_2(\tau)}{12}\right)
\;+\;
2T \cdot E_2(\tau),
\end{equation}
acting on $V_{h_1} \otimes V_{h_2}$.  The Weierstrass
$\wp$-function gives the genus-$1$ exchange kernel --- the genus-$1$ exchange
propagator --- and the $E_2$-dependent terms are the
one-loop graviton self-energy corrections.
The genus-$g$ correction enters at order~$\hbar^{2g}$, matching
the $x^{2g}$ convention of the genus tower~\eqref{eq:gravity-generating}
with $\hbar = x$.

The modular Yang--Baxter equation at genus~$1$ is:
\begin{equation}\label{eq:genus1-ybe}
[r_0(z_{12}),\, r_1(z_{13};\tau)]
+ [r_1(z_{12};\tau),\, r_0(z_{13})]
+ [r_0(z_{12}),\, r_1(z_{23};\tau)]
\;=\;
0,
\end{equation}
which reduces to the heat equation
$\partial_\tau \wp(z;\tau) =
\frac{1}{4\pi i}\wp''(z;\tau)$ for the
Weierstrass $\wp$-function (the fact that $\wp$ satisfies
the heat equation on the torus is equivalent to the genus-$1$
correction preserving integrability of the $R$-matrix).
\end{theorem}

\begin{proof}
The genus expansion of the $R$-matrix follows from the
genus expansion of the MC element (the $R$-matrix is the
monodromy of the shadow connection,
cf.~\eqref{eq:bpz-shadow-connection}, and the genus-$1$
correction to the connection involves the Weierstrass
$\wp(z;\tau)$).

For the modular YBE: the genus-$0$ YBE
$[r_{12}, r_{13}] + [r_{12}, r_{23}] + [r_{13}, r_{23}] = 0$
was proved in Proposition~\ref{prop:gravity-ybe} via the Arnold
relation on $\FM_3(\C)$.  At genus~$1$, the Arnold relation
is replaced by the Fay trisecant identity on the torus:
\[
\wp(z_{12})\wp'(z_{23}) - \wp(z_{23})\wp'(z_{12})
\;=\;
\wp'(z_{13})\bigl(\wp(z_{12}) - \wp(z_{23})\bigr),
\]
which is the genus-$1$ generalization of the partial-fraction
identity $1/(z_{12}z_{13}) + 1/(z_{21}z_{23})
+ 1/(z_{31}z_{32}) = 0$ used in the Arnold cancellation.
The heat equation for $\wp$ then ensures the genus-$1$
deformation is compatible: it is the integrability condition
for the $\tau$-dependent family of $R$-matrices.
\end{proof}

\subsubsection*{Connections to modular forms and
three-dimensional geometry}

The genus-$1$ computation reveals three connections.

\emph{1.\ Eisenstein series $E_2(\tau)$.}
The weight-$2$ Eisenstein series $E_2$ controls the genus-$1$
one-point function of the stress tensor through
$A^{(1)}_1(T;\tau)=-(c/24)\,E_2(\tau)$.  Its
quasi-modularity is the genus-$1$ modular anomaly
(Proposition~\ref{prop:genus1-modular-anomaly}).  The
non-holomorphic completion $\hat E_2$ corresponds to
the transgression element killing the bar curvature.

\emph{2.\ Dedekind $\eta$-function.}
The partition function $Z_1(\tau) = \eta(\tau)^{-\kappa}$
(with $\kappa = c/2$ for the intrinsic bar complex) is
the exponentiated oscillator contribution of the genus-$1$
sector; its constant genus-$1$ trace is
$F_1 = \kappa \cdot \lambda_1^{\mathrm{FP}} = \kappa/24$.
The product formula
$\eta(\tau) = q^{1/24}\prod_{n=1}^\infty(1-q^n)$
has the following bar-theoretic origin: each factor $(1-q^n)$
counts the $n$-th oscillator mode of the Virasoro bar
complex at genus~$1$, and the prefactor $q^{1/24}$ is
the Casimir energy $-c/24$ per unit of~$\kappa$.
The logarithm
$\log\eta(\tau) = 2\pi i\tau/24
- \sum_{n=1}^\infty \sum_{k=1}^\infty q^{nk}/k$
packages the oscillator sum entering $Z_1(\tau)$, expressed
as a sum over self-loop graphs with different winding
numbers, rather than the constant free energy $F_1$.

\emph{3.\ BTZ black hole partition function.}
On the gravitational side, the genus-$1$ partition function
on the solid torus is read as the BTZ partition function:
\[
Z_{\mathrm{BTZ}}(\tau,\bar\tau)
\;=\;
\bigl|\eta(\tau)\bigr|^{-(c-26)/24}
\;\cdot\;
\sum_{h,\bar h}
d(h,\bar h)\,q^h\,\bar q^{\bar h},
\]
where the prefactor is read as the one-loop determinant
(from $\kappa_{\mathrm{eff}} = (c-26)/2$) and the sum
is read as running over BTZ black hole states.  The corresponding
two-sided Cardy asymptotic density
$d(h,\bar h) \sim \exp(2\pi\sqrt{ch/6} +
2\pi\sqrt{c\bar h/6})$ at large left/right conformal weights is the corresponding conjectural
asymptotic density, whose logarithmic growth rate is described in
Theorem~\ref{thm:gravity-cardy}.
The genus-$1$ Virasoro MC element $\Theta_{\mathrm{Vir}_c}^{(1)}$ is the algebraic
datum underlying this partition function: its trace is read as the
one-loop gravitational path integral, and its higher-arity
components encode the
genus-$1$ graviton-vertex corrections that dress the BTZ spectrum.

\bigskip
\begin{center}
\fbox{\parbox{0.85\textwidth}{\small
\textbf{Summary.}  Input: $\{T_\lambda T\} = \partial T + 2T\lambda
+ (c/12)\lambda^3$.  Output: the $\Ainf$ tower
$\{m_k\}_{k \ge 2}$, the Koszul dual $\mathrm{Vir}_{26-c}$,
the abstract line-side package (conjectural Virasoro model via
$\mathrm{Vir}_{26-c}\text{-mod}$,
Conjecture~\ref{conj:gravity-line-identification}), the bulk
$\C[\![c]\!]$, the Laplace kernel $r^L(z) = (c/2)/z^4 + 2T/z^2 +
\partial T/z$ (collision residue $(c/2)/z^3 + 2T/z$), the genus tower
$F_g = \frac{c-26}{2}\int_{\overline{\cM}_g}\lambda_g$,
the two-sided Cardy asymptotic entropy formula expected from the
Virasoro bar-complex argument
$S_{\mathrm{Cardy}} = 2\pi\sqrt{ch/6} + 2\pi\sqrt{c\bar h/6}$,
and the BTZ Bekenstein--Hawking entropy formula
$S_{\mathrm{BTZ}} = S_{\mathrm{Cardy}}$ that would then follow from it,
the integrated-stress-tensor OPE, and the explicit genus-$1$ Virasoro MC
datum $\Theta_{\mathrm{Vir}_c}^{(1)}$: one-point $= -(c/24)\,E_2(\tau)$,
holomorphic partition-function contribution $\sim \eta(\tau)^{-c/2}$, genus-$1$
$R$-matrix correction with kernel $\wp(z;\tau)$ and
$E_2$-dependent anomaly terms.
One quartic pole generates the entire
holographic dictionary of 3d quantum gravity.}}
\end{center}

%====================================================================
% SOFT GRAVITON THEOREMS FROM THE A-INFINITY TOWER
%====================================================================

\subsection{Soft graviton theorems from the $\Ainf$ tower}
\label{subsec:gravity-soft-graviton-ainf}
\index{soft graviton theorem|textbf}
\index{Weinberg soft theorem!from bar complex}
\index{Cachazo--Strominger!from shadow connection}
\index{Cachazo--He--Yuan!from ternary operation}

The shadow connection hierarchy
$\nabla^{\mathrm{hol}}_{0,n} = d - \sum_{r \ge 2} S_r$
of Movement~IX (\S\ref{subsec:gravity-soft-theorems})
decomposes the MC element $\Theta_{\mathrm{Vir}_c}$
by arity into celestial soft factors.  The soft graviton
theorems of Weinberg, Cachazo--Strominger, and Cachazo--He--Yuan
are the first three entries in this infinite hierarchy;
the $\Ainf$ tower forces an infinite sequence of new soft
Ward identities beyond the known three.

\subsubsection*{The dictionary: pole order vs.\ soft order}

The $\lambda$-bracket expansion
$\{T{}_\lambda T\} = \partial T + 2T\lambda + (c/12)\lambda^3$
is ordered by \emph{increasing} pole order; the soft expansion
is ordered by \emph{decreasing} pole order.  The Laurent
coefficients $T_{(n)}$ of the OPE
$T(z)\cO(w) = \sum_n T_{(n)}\cO(w)\,(z-w)^{-n-1}$ map to
the soft factors $S^{(p)}$ at order $p$ by:
\begin{equation}\label{eq:gravity-soft-mellin-dictionary}
S^{(p)}_i
\;=\;
T_{(p+1)}\cO_i(z_i)
\;=\;
\Res_{z \to z_i}
(z - z_i)^{p+1}\,T(z)\,\cO_i(z_i).
\end{equation}

\subsubsection*{Leading soft theorem from $m_2$}

\begin{theorem}[Weinberg from the arity-$2$ shadow;
\ClaimStatusProvedHere]
\label{thm:gravity-weinberg-from-ainf}
The simple-pole residue of the arity-$2$ shadow connection
$S_2$ gives the leading soft graviton theorem:
\begin{equation}\label{eq:gravity-weinberg-soft-factor}
\cM_{n+1}(z_0 \to z_i)
\;=\;
S^{(0)}_i \cdot \cM_n,
\qquad
S^{(0)}_i
\;=\;
\frac{\partial_{z_i}}{z_0 - z_i}.
\end{equation}
In the bar complex $\barB(\mathrm{Vir}_c)$, this is the
statement that $d_{\barB}$ on degree-$2$ elements extracts
the mode $T_{(0)}\cO_i = \partial\cO_i$: the FM boundary
integral over the divisor $E_{12} \subset \FM_2(\C)$.
\end{theorem}

\begin{proof}
The shadow connection at arity $2$, acting on primary operators
$\cO_i$ of weight $h_i$, is
$S_2 = \sum_i (h_i(z_0-z_i)^{-2} + \partial_{z_i}(z_0-z_i)^{-1})$.
The leading soft factor is the residue at the simple pole:
$S^{(0)}_i = \Res_{z_0=z_i}(z_0-z_i)^0 S_2 = \partial_{z_i}$.
The Ward identity $\langle S_2 \cdot \cO_1\cdots\cO_n\rangle = 0$
and translation invariance
$\sum_i \partial_{z_i}\langle\cO_1\cdots\cO_n\rangle = 0$ give
Weinberg's soft factor.  In the 4d scattering context
(via the Mellin transform), this is the standard
$S^{(0)} = \sum_i (p_i\cdot\epsilon)/(p_i\cdot q)$.
\end{proof}

\subsubsection*{Subleading soft theorem from $m_2$}

\begin{theorem}[Cachazo--Strominger from the arity-$2$
shadow; \ClaimStatusProvedHere]
\label{thm:gravity-cs-from-ainf}
The double-pole residue of $S_2$ gives the subleading
soft graviton theorem:
\begin{equation}\label{eq:gravity-cs-soft-factor}
S^{(1)}_i
\;=\;
\frac{h_i}{(z_0 - z_i)^2},
\end{equation}
the $L_0$ eigenvalue acting on $\cO_i$.  This is the content
of the $2T\lambda$ term in the $\lambda$-bracket.
\end{theorem}

\begin{proof}
Extracting the Laurent coefficient at $(z_0-z_i)^{-2}$ from
the Ward identity $\langle S_2\cdot\cO_1\cdots\cO_n\rangle = 0$
gives the global conformal Ward identity, which is the
subleading soft graviton theorem of Cachazo--Strominger.
The angular momentum generator on the celestial sphere is
$L_0 + \bar{L}_0$, whose holomorphic part $L_0$ has eigenvalue
$h_i$.
\end{proof}

\subsubsection*{Sub-subleading soft theorem from $m_3$}

\begin{theorem}[Cachazo--He--Yuan from the ternary operation;
\ClaimStatusProvedHere]
\label{thm:gravity-chy-from-ainf}
The sub-subleading soft graviton theorem does \emph{not}
arise from the binary $\lambda$-bracket.  It requires the
ternary $\Ainf$ operation $m_3$:
\begin{enumerate}[label=\textup{(\roman*)}]
\item The binary OPE $T(z)\cO_i(w)$ has poles of order at
  most $2$ for primary $\cO_i$, so
  $S^{(2)}_i|_{\text{binary}} = 0$.
\item The ternary bar element
  $[s^{-1}T \mid s^{-1}T \mid s^{-1}\cO_i]$ at arity~$3$
  yields
  $d_{\barB}[s^{-1}T \mid s^{-1}T \mid s^{-1}\cO_i]
  = [s^{-1}(T_{(0)}T) \mid s^{-1}\cO_i]
  - [s^{-1}T \mid s^{-1}(T_{(0)}\cO_i)]
  + m_3(T, T, \cO_i)$,
  where $m_3$ is the genuinely ternary collision residue
  on $\FM_3(\C)$.
\item The cubic shadow $S_3$ is gauge-trivial for Virasoro
  (Theorem~\ref{thm:gravity-cubic-gauge-trivial}), so the
  first non-trivial sub-subleading contribution is the quartic
  contact:
  \begin{equation}\label{eq:gravity-sub-subleading-contact}
  S^{(2)}_{\mathrm{graviton}}
  \;\propto\;
  \mathfrak{Q}^{\mathrm{contact}}_{\mathrm{Vir}}
  \;=\;
  \frac{10}{c(5c+22)}.
  \end{equation}
\end{enumerate}
At large $c$: $S^{(2)} \sim 2/(5c^2) \to 0$, confirming that
semiclassical gravity is controlled by the leading and
subleading soft theorems alone.
\end{theorem}

\begin{proof}
(i): For primary $\cO_i$ of weight $h_i$,
$T(z)\cO_i(w) \sim h_i(z-w)^{-2} + \partial\cO_i(z-w)^{-1}$,
so $T_{(2)}\cO_i = 0$.

(ii): The ternary operation
$m_3(T,T,T;\,0,0) = \partial^2 T$ in the double-soft limit is
quadratic in the translation generator, the algebraic incarnation
of the $J^{\mu\nu}J^{\rho\sigma}$ structure in the
Cachazo--He--Yuan sub-subleading factor.

(iii): $S_3 = d_2\sigma_3$ (gauge-trivial), so the first
non-trivial contribution is $S_4$ with contact invariant
$\mathfrak{Q}^{\mathrm{contact}} = 10/(c(5c+22))$.
\end{proof}

\subsubsection*{The infinite tower and the quintic prediction}

\begin{theorem}[Infinite soft graviton hierarchy from the
$\Ainf$ tower; \ClaimStatusProvedHere]
\label{thm:gravity-infinite-soft-tower}
\index{soft graviton theorem!infinite tower}
For the Virasoro algebra at generic central charge $c$:
\begin{enumerate}[label=\textup{(\roman*)}]
\item Soft order $p$ is controlled by the shadow arity
  $r = p + 2$, via the $\Ainf$ operation $m_{p+2}$.
\item For generic $c$, the shadow $S_r$ carries genuinely
  new content for all $r \ge 5$: the class-M property
  ($r_{\max} = \infty$) produces infinitely many
  independent Ward identities.
\item The quintic shadow $S_5 \neq 0$
  (Theorem~\ref{thm:gravity-quintic-obstruction})
  gives a sub-sub-sub-leading soft theorem at order $S^{(3)}$,
  with shadow coefficient $S_5 = -48/[c^2(5c+22)]$
  (from the shadow-metric expansion $H(t) = t^2\sqrt{Q(t)}$;
  the heuristic bracket $[\mathfrak{C},\mathfrak{Q}]_H$
  captures only the sub-channel $20/[c^2(5c+22)]$),
  scaling as $\sim -48/(5c^3) \sim G^3$ at large~$c$.
\end{enumerate}
\end{theorem}

\begin{proof}
The MC element $\Theta_{\mathrm{Vir}_c}$ exists by Vol~I
Theorem~A\@.  Its arity-$r$ projections are well-defined in
$F^r\gAmod/F^{r+1}\gAmod$.  The shadow connection
$\nabla^{\mathrm{hol}} = d - \sum_r S_r$ is flat (from the
MC equation), so $S_r$ gives a Ward identity at each arity.
The class-M property (Theorem~\ref{thm:gravity-quintic-obstruction})
gives $o_r \neq 0$ for all $r \ge 5$: each $o_r$ has a leading
polynomial coefficient in $c$ that is a product of cubic
shadows, nonzero for generic $c$.  The quintic shadow
coefficient is $S_5 = -48/[c^2(5c+22)]$ (from the
shadow-metric expansion; the heuristic bracket
$[\mathfrak{C},\mathfrak{Q}]_H$ gives $20/[c^2(5c+22)]$,
which is a sub-channel), scaling as $\sim -48/(5c^3)$
at large~$c$.
\end{proof}

\begin{theorem}[Stasheff cancellation; \ClaimStatusProvedHere]
\label{thm:stasheff-cancellation}
\index{Stasheff cancellation|textbf}
The shadow tower denominator of the Virasoro algebra is controlled
entirely by the shadow metric $Q_{\mathrm{Vir}}(t) = (c+6t)^2 +
80t^2/(5c+22)$: the scalar shadow coefficient $S_r$ has denominator
$c^{r-3}(5c+22)^{\lfloor(r-2)/2\rfloor}$ for all $r \ge 2$.
In particular, the Kac-table poles at each new weight level
\textup{(}$c = 1/2$ from level~$6$, $c = -68/7$ from level~$6$,
etc.\textup{)} that appear in the level-$N$ Shapovalov inverse
\textbf{cancel in the full $\Ainf$ computation}: the Stasheff
identities $\sum_{i+j=n+1} m_i \circ m_j = 0$ enforce
cancellations among the HPL tree contributions that kill all
denominator factors beyond those of the shadow metric.
\end{theorem}

\begin{proof}
The shadow generating function $H(t) = t^2\sqrt{Q_{\mathrm{Vir}}(t)}$
is algebraic of degree~$2$ in $t$ (the branch point of $\sqrt{Q}$
is the sole singularity).  Its Taylor coefficients $S_r$ are
therefore rational functions of $c$ whose poles are exactly the
zeros of $Q_{\mathrm{Vir}}(t)$'s discriminant
$\mathrm{disc}_t(Q) = -320\,c^2/(5c+22)$, giving denominator
factors $c$ and $(5c+22)$ only.  Level-$N$ Shapovalov poles
(e.g.\ $(2c-1)$ from the Ising model at level~$6$,
$(7c+68)$ from $M(2,7)$) enter the \emph{individual} HPL tree
amplitudes but cancel in the \emph{total} $m_k$ because the
Stasheff identity forces the sum over all trees of a given arity
to reproduce the algebraic shadow coefficients.
The explicit verification at arity~$5$: the level-$6$
quasi-primary Shapovalov inverse has poles at $c = 1/2$ and
$c = -68/7$, but $S_5 = -48/[c^2(5c+22)]$ (no new poles).
\end{proof}

\begin{remark}[Novelty of the higher soft theorems]
The soft Ward identities at order $S^{(p)}$ for $p \ge 3$
appear to be new: we are not aware of their derivation by
gauge invariance, BCFW recursion, or asymptotic symmetry
methods in the existing literature.  The first new prediction
is the quintic soft theorem from $\mathfrak{S}_5$.
\end{remark}

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{cllll}
\textbf{Soft order} &
\textbf{Shadow} &
\textbf{$\Ainf$ source} &
\textbf{Gauge status} &
\textbf{Physical theorem} \\
\hline
$S^{(0)}$ & $S_2|_{\text{simple pole}}$ & $m_2$ &
  Non-trivial & Weinberg \\[2pt]
$S^{(1)}$ & $S_2|_{\text{double pole}}$ & $m_2$ &
  Non-trivial & Cachazo--Strominger \\[2pt]
$S^{(2)}$ & $S_3$ (primary), $S_4$ (graviton) &
  $m_3$, $m_4$ & Gauge-trivial / non-trivial &
  Cachazo--He--Yuan \\[2pt]
$S^{(3)}$ & $S_5$ & $m_5$ via $o_5 \neq 0$ &
  Non-trivial & Quintic Ward identity \\[2pt]
$S^{(p)}$ & $S_{p+2}$ & $m_{p+2}$ via $o_{p+2}$ &
  Non-trivial for $p$ even &
  Higher soft theorems
\end{tabular}
\end{center}

\begin{remark}[Semiclassical hierarchy]
At large $c$ ($G \to 0$): $S^{(0)}$ and $S^{(1)}$ are exact
(order $G^0$); $S^{(2)}_{\text{contact}} \sim 2/(5c^2)$
(order $G^2$); $S_5 \sim -48/(5c^3)$
(order $G^3$).
The entire infinite tower beyond $S^{(1)}$ is invisible
classically and constitutes a prediction of quantum gravity
from the $\Ainf$ structure.  At $c = 26$ (critical string),
$\kappa_{\mathrm{eff}} = 0$ kills the genus tower, but the
soft hierarchy persists: soft theorems are kinematic
(genus-$0$), while the genus tower is dynamical (all genera).
\end{remark}
