\label{chap:log-ht-monodromy-frontier}

\providecommand{\A}{\mathcal A}
\providecommand{\B}{\mathcal B}
\providecommand{\C}{\mathcal C}
\providecommand{\D}{\mathcal D}
\providecommand{\E}{\mathcal E}
\providecommand{\F}{\mathcal F}
\providecommand{\FM}{\mathrm{FM}}
\providecommand{\Conf}{\mathrm{Conf}}
\providecommand{\HT}{\mathrm{HT}}
\providecommand{\MC}{\mathrm{MC}}
\providecommand{\End}{\mathrm{End}}
\providecommand{\Hom}{\mathrm{Hom}}
\providecommand{\id}{\mathrm{id}}
\providecommand{\im}{\mathrm{im}}
\providecommand{\Res}{\mathrm{Res}}
\providecommand{\Aut}{\mathrm{Aut}}
\providecommand{\PP}{\mathbb P}
\providecommand{\AA}{\mathbb A}
\providecommand{\CC}{\mathbb C}
\providecommand{\RR}{\mathbb R}
\providecommand{\ZZ}{\mathbb Z}
\providecommand{\kk}{\Bbbk}
\providecommand{\h}{\hbar}
\providecommand{\ot}{\otimes}
\providecommand{\wot}{\widehat\otimes}
\providecommand{\s}{\mathrm s}
\providecommand{\ds}{d_{\mathrm{dR}}}
\providecommand{\CYB}{\mathrm{CYB}}
\providecommand{\PT}{\mathrm{PT}}
\providecommand{\Disk}{\mathrm{Disk}}
\providecommand{\SN}{\boldsymbol{\nabla}}
\providecommand{\bboxprod}{\mathop{\boxtimes}\displaylimits}

\title{Logarithmic HT Monodromy: Extensions and Frontier}
\author{}
\date{}

\section{Conjectural bridges}
The theorem package above strongly suggests further identifications.

\begin{conjecture}[Monodromy equals spectral $R$-matrix; \ClaimStatusConjectured]\label{conj:rmatrix}
The monodromy of the reduced HT logarithmic connection is the spectral $R$-matrix controlling the factorisation quantum group attached to the same dg-shifted Yangian data. Equivalently, the monodromy functor of the reduced connection should identify the reduced HT line category with the representation category of the associated factorisation quantum group.
\end{conjecture}

\begin{conjecture}[Logarithmic CoHA envelope; \ClaimStatusConjectured]\label{conj:coha}
For a symmetric quiver there exists a logarithmic factorization enhancement of the cohomological Hall algebra on logarithmic Fulton-MacPherson spaces whose connected collision residues recover the primitive Lie algebra of the associated vertex algebra, and whose Koszul dual is the dg-shifted Yangian governing the corresponding line-operator theory.
\end{conjecture}

\begin{conjecture}[Chromatic lift of the HT associator; \ClaimStatusConjectured]\label{conj:chromatic-lift}
For a spectral lift of a resolved logarithmic HT datum, the reduced HT associator admits compatible chromatic localizations whose height-$n$ pieces are computed by the monodromy of the $E(n)$-localized spectral factorization algebra.
\end{conjecture}


\section{The Kazhdan--Lusztig argument for the monodromy-braiding identification}
\label{sec:KL-monodromy-braiding}

\providecommand{\fS}{\mathfrak{S}}
\providecommand{\cL}{\mathcal{L}}
\providecommand{\cM}{\mathcal{M}}
\providecommand{\Nil}{\widetilde{\mathcal{N}}}
\providecommand{\fg}{\mathfrak{g}}

In geometric representation theory, the identification of KZ monodromy with the $R$-matrix of the quantum group is the Kazhdan--Lusztig theorem \cite{kazhdan-lusztig-iv}.
Their proof proceeds by computing the monodromy of the KZ connection via convolution on the Steinberg variety $Z = \Nil \times_{\fg} \Nil$: the flat connection on category $\mathcal{O}$ is realized geometrically as the Gauss--Manin connection on the family of Steinberg fibers, and its monodromy is read off from the convolution product, which is the quantum group multiplication.
The chiral bar complex gives us the same ingredients in the HT setting.
The Steinberg object $\fS_b = \cL_b \times_{\cM_{\mathrm{vac}}} \cL_b$ (the self-fiber-product of the boundary line category over the vacuum moduli) plays the role of $Z$; the $(-1)$-shifted symplectic structure plays the role of the moment map; and the bar differential encodes the convolution.
We now make this precise.

\begin{construction}[The flat connection on $\fS_b$]
\label{constr:steinberg-connection}
The $(-1)$-shifted symplectic form $\omega_{-1}$ on the Steinberg object $\fS_b = \cL_b \times_{\cM_{\mathrm{vac}}} \cL_b$ carries a canonical Liouville $1$-form $\theta$ (the primitive of the shifted symplectic structure, existing because the fiber product is Lagrangian in the appropriate sense).
When the HT theory is defined on a punctured disk $D^* \subset \CC$, the family $\fS_b \to D^*$ acquires a flat connection
\[
\nabla_{\fS} \;=\; d \;+\; r(z)\,\frac{dz}{z}
\]
where $r(z)$ is the classical $r$-matrix: the residue of the chiral OPE along the collision divisor $\Delta \subset D \times D$.
Concretely, $r(z)$ is the element of $\End(\cL_b^{\ot 2})$ whose matrix coefficients are the polar parts of the OPE coefficients $\langle a(w)\, b(z) \rangle \sim \sum_{n \geq 0} (a_{(n)} b)(z) / (w-z)^{n+1}$.
The flatness $\nabla_{\fS}^2 = 0$ is equivalent to the classical Yang--Baxter equation $[\CYB](r) = 0$, which holds because $r(z)$ is the structure constant of a vertex algebra OPE (the Jacobi identity implies CYB).

The connection is regular singular at $z = 0$: the pole is simple because the OPE has at worst first-order poles after extracting the leading singularity.
The monodromy around $z = 0$ is therefore
\[
\mathrm{Mon}_{z=0}(\nabla_{\fS}) \;=\; \exp\bigl(2\pi i \cdot \Res_{z=0}\, r(z)\,dz/z\bigr) \;=\; \exp(2\pi i\, r_0)
\]
where $r_0 = \Res_{z=0} r(z)\,dz/z$ is the residue endomorphism.
\end{construction}

\begin{proposition}[Free $\beta\gamma$ verification; \ClaimStatusProvedHere]
\label{prop:betagamma-monodromy}
For the free $\beta\gamma$ system on $\CC_z \times \RR_t$, the monodromy of $\nabla_{\fS}$ around $z = 0$ equals the spectral $R$-matrix $R(z) = 1 + r/z + O(1/z^2)$ of the Heisenberg vertex algebra.
\end{proposition}

\begin{proof}
The $\beta\gamma$ OPE is $\beta(w)\gamma(z) \sim 1/(w-z)$, so $r(z) = r_0/z$ with $r_0 = \beta \ot \gamma - \gamma \ot \beta \in \End(V^{\ot 2})$ the standard Heisenberg $r$-matrix.
The connection $\nabla_{\fS} = d + (r_0/z)\,dz/z$ on $D^*$ is regular singular with residue $r_0$.
Its monodromy is
\[
\mathrm{Mon}_{z=0}(\nabla_{\fS}) \;=\; \exp(2\pi i\, r_0).
\]
On the other hand, the spectral $R$-matrix of the rank-one Heisenberg algebra is $R(z) = \exp(r_0/z)$ evaluated at $z = 1$ after analytic continuation, and its monodromy representation on $\mathrm{Rep}(\mathrm{Heis})$ is precisely $\exp(2\pi i\, r_0)$.
The two expressions coincide.
\end{proof}

\begin{strategy}[Reduction to analytic input; \ClaimStatusConjectured]
\label{strat:general-monodromy}
For a general physical HT realization satisfying the hypotheses of
Theorem~\ref{thm:physics-bridge}, the identification
$\mathrm{Mon}_{z=0}(\nabla_{\fS}) = R(z)$ reduces to showing two things:
\begin{enumerate}
\item The connection $\nabla_{\fS}$ is regular singular at $z = 0$, with residue given by the Laplace transform of the $\lambda$-bracket:
\[
r_0 \;=\; \Res_{z=0}\, r(z)\,\frac{dz}{z}, \qquad r(z) \;=\; \int_0^\infty e^{-\lambda z}\, \{{\cdot}\,_\lambda\,{\cdot}\}\, d\lambda,
\]
which is the DK-0 formula proved in the core chapter (\S\ref{sec:DK-0}).
\item The exponential $\exp(2\pi i\, r_0)$ converges in the completed tensor product $\End(\cL_b^{\wot 2})$ and defines a braided monoidal structure on $\cL_b\text{-}\mathrm{mod}$.
\end{enumerate}
The first point is the content of the DK-0 bridge: the classical $r$-matrix IS the Laplace transform of the PVA $\lambda$-bracket, and this is proved.
The second point is an analytic convergence statement that holds
whenever the $\lambda$-bracket has at most polynomial growth in
$\lambda$ --- a condition guaranteed in physical realizations by the
decay conclusion of Theorem~\ref{thm:physics-bridge}.
Granting this, the monodromy computation follows from the same matrix exponential as in the free field case.
\end{strategy}

\begin{remark}[Geometric origin of the $(2,0)$-projection]
\label{rmk:20-projection-origin}
The Steinberg presentation theorem of the core chapter identifies the line category $\cL_b$ with modules over the convolution algebra $H_*(\fS_b)$, and the algebraic projection $\alpha_T^{(2,0)}$ --- the $(2,0)$-component of the HT associator --- with the class of the diagonal $[\Delta] \in H_*(\fS_b)$.
Proposition~\ref{prop:betagamma-monodromy} shows this in the free $\beta\gamma$ model: there the corresponding $(2,0)$-projection is the \emph{monodromy} of the Steinberg connection $\nabla_{\fS}$, computed as the exponential of the residue $r$-matrix.
For a general physical HT realization, the strategy above and Conjecture~\ref{conj:rmatrix} suggest the same interpretation, but that extension remains frontier.
In the proved free-field case, the algebraic projection and the analytic monodromy are two descriptions of the same invariant, related by the passage from the Steinberg variety to its family over $D^*$.
This is the chiral analog of the fact that, in the classical Kazhdan--Lusztig theorem, the quantum group $R$-matrix is simultaneously an algebraic object (convolution on $Z$) and an analytic object (monodromy of KZ).
\end{remark}


\section{The derived Steinberg construction for the affine lineage}
\label{sec:derived-steinberg-affine}

\providecommand{\fg}{\mathfrak{g}}
\providecommand{\fh}{\mathfrak{h}}
\providecommand{\fn}{\mathfrak{n}}
\providecommand{\fb}{\mathfrak{b}}
\providecommand{\cL}{\mathcal{L}}
\providecommand{\cM}{\mathcal{M}}
\providecommand{\cO}{\mathcal{O}}
\providecommand{\fS}{\mathfrak{S}}

Theorem~\ref{thm:affine-monodromy-identification} identifies the monodromy of the reduced HT connection with the quantum group $R$-matrix on \emph{evaluation modules}: finite-dimensional $\fg$-modules pulled back to the dg-shifted Yangian $Y_\h(\fg)$ via the evaluation homomorphism.  This section develops the derived Steinberg geometry that governs the extension beyond evaluation modules, working in detail for $\fg = \mathfrak{sl}_2$ at generic level $k$.

The genuine new content beyond the classical Drinfeld--Kohno theorem is not the monodromy identification itself (which DK proves on category $\cO$ modules), but rather:
\begin{enumerate}[label=\textup{(\alph*)}]
\item the \emph{factorization origin}: the connection arises from the Swiss-cheese bar complex, not from the Sugawara construction alone;
\item the \emph{derived enhancement}: the full $A_\infty$ superconnection on the bar coalgebra $\mathrm{B}(V^k(\fg))$ reduces to the KZ connection after one-loop collapse, but carries strictly more information before reduction;
\item the \emph{Steinberg geometric package}: the monodromy acts on the derived Steinberg object $\fS_b$, which carries a $(-1)$-shifted symplectic structure encoding both the convolution product and the line-category tensor structure;
\item the \emph{convergence mechanism}: the exponential $\exp(2\pi i \cdot \Omega/(k+h^\vee))$ converges in the completed tensor product $\End(\cL_b^{\wot 2})$ by a spectral bound that uses the polynomial growth of the $\lambda$-bracket, not merely the finite-dimensionality of evaluation modules.
\end{enumerate}

\subsection{The Steinberg object for $\mathfrak{sl}_2$ at level $k$}
\label{subsec:steinberg-sl2}

Let $\fg = \mathfrak{sl}_2$ with Cartan subalgebra $\fh = \CC h$, Borel $\fb = \fh \oplus \CC e$, nilpotent radical $\fn = \CC e$, and opposite nilpotent $\fn^- = \CC f$.  The dual Coxeter number is $h^\vee = 2$.

\begin{definition}[Derived Steinberg object; \ClaimStatusConjectured]
\label{def:derived-steinberg}
Let $\cM_{\mathrm{vac}}$ denote the formal moduli of flat $\fg$-bundles on $D = \Spec \CC[[z]]$ (equivalently, the formal neighbourhood of the trivial connection in the stack $\mathrm{LocSys}_\fg(D)$).  Let $\cL_b$ denote the dg category of line operators on the boundary $\RR_t$ of the 3d HT theory on $\CC_z \times \RR_t$ associated to $V^k(\fg)$.  The \emph{derived Steinberg object} is the self-fiber-product
\[
  \fS_b \;:=\; \cL_b \times_{\cM_{\mathrm{vac}}} \cL_b.
\]
Here the maps $\cL_b \to \cM_{\mathrm{vac}}$ are the forgetful functors that remember only the flat-bundle datum underlying each line operator.
\end{definition}

For $\fg = \mathfrak{sl}_2$, the classical shadow of $\fS_b$ is the Springer resolution.  The Steinberg variety $Z = \widetilde{\mathcal{N}} \times_\fg \widetilde{\mathcal{N}}$ (the self-fiber-product of the Springer resolution $\pi: \widetilde{\mathcal{N}} = T^*(\fg/\fb) \to \fg$) carries the convolution product that Kazhdan--Lusztig use to construct the quantum group.  The derived Steinberg object $\fS_b$ is the HT-theoretic upgrade: it remembers the full dg category of line operators, not only the abelian category of evaluation modules.

\begin{construction}[The Steinberg connection for $V^k(\mathfrak{sl}_2)$; \ClaimStatusProvedHere]
\label{constr:steinberg-sl2}
The family $\fS_b \to D^* = \Spec \CC((z))$ acquires a flat connection from the bar-complex structure of $V^k(\mathfrak{sl}_2)$.  By Theorem~\ref{thm:reduced-equals-kz}, the reduced connection on evaluation modules is the KZ connection.  On the full Steinberg object, the connection is
\[
  \nabla_\fS \;=\; d \;+\; \frac{1}{k+2}\,\Omega\,\frac{dz}{z},
\]
where $\Omega = e \ot f + f \ot e + \tfrac{1}{2} h \ot h$ is the split Casimir of $\mathfrak{sl}_2$.

The connection is regular singular at $z = 0$ with residue $\Omega/(k+2)$.  The regularity is a consequence of AP19: the bar kernel absorbs one pole from the $z^{-2}$ OPE singularity $J^a(w)J^b(z) \sim k\kappa^{ab}/(w-z)^2 + f^{ab}_c J^c(z)/(w-z)$, so the collision $r$-matrix has a simple pole $r(z) = \Omega/z + O(1)$.  The $d\log z$ kernel absorbs the pole from $dz/z$, producing the claimed form.
\end{construction}

\subsection{Casimir spectrum and eigenvalue computation for general representations}
\label{subsec:casimir-general}

For the monodromy on arbitrary finite-dimensional representations (not just the fundamental), the following computation is essential.

\begin{proposition}[Casimir eigenvalues on $\mathfrak{sl}_2$ tensor products; \ClaimStatusProvedHere]
\label{prop:casimir-general-sl2}
Let $V_{j_1}$ and $V_{j_2}$ be the irreducible $\mathfrak{sl}_2$-representations of spin $j_1$ and $j_2$ (so $\dim V_j = 2j+1$).  On $V_{j_1} \ot V_{j_2}$, the Clebsch--Gordan decomposition gives
\[
  V_{j_1} \ot V_{j_2} \;\cong\; \bigoplus_{j = |j_1 - j_2|}^{j_1 + j_2} V_j.
\]
On each summand $V_j$, the split Casimir acts as
\[
  \Omega_{12}\big|_{V_j} \;=\; \tfrac{1}{2}\bigl(j(j+1) - j_1(j_1+1) - j_2(j_2+1)\bigr)\,\id_{V_j}.
\]
\end{proposition}

\begin{proof}
The total Casimir $C_{\mathrm{tot}} = (J^a_1 + J^a_2)(J_{1,a} + J_{2,a})$ decomposes as $C_1 + C_2 + 2\Omega_{12}$, where $C_i = J^a_i J_{i,a}$ is the quadratic Casimir in the $i$-th factor.  On $V_j \subset V_{j_1} \ot V_{j_2}$, one has $C_{\mathrm{tot}} = j(j+1)$, $C_1 = j_1(j_1+1)$, $C_2 = j_2(j_2+1)$, giving the claimed formula.
\end{proof}

\begin{theorem}[Monodromy eigenvalues for general $\mathfrak{sl}_2$ representations; \ClaimStatusProvedHere]
\label{thm:monodromy-eigenvalues-sl2}
The monodromy of the Steinberg connection around $z = 0$, acting on $V_{j_1} \ot V_{j_2}$, is diagonal on the Clebsch--Gordan summands:
\[
  \mathrm{Mon}_{z=0}(\nabla_\fS)\big|_{V_j}
  \;=\;
  \exp\!\Bigl(\frac{\pi i\bigl(j(j+1) - j_1(j_1+1) - j_2(j_2+1)\bigr)}{k+2}\Bigr)\,\id_{V_j}
\]
for each $j = |j_1 - j_2|, |j_1 - j_2| + 1, \dots, j_1 + j_2$.
\end{theorem}

\begin{proof}
The monodromy is $\exp(2\pi i \cdot \Omega_{12}/(k+2))$.  Substituting the eigenvalues from Proposition~\ref{prop:casimir-general-sl2} gives the formula.
\end{proof}

\begin{corollary}[Agreement with quantum group $R$-matrix; \ClaimStatusProvedHere]
\label{cor:agreement-qgroup-sl2}
At $q = e^{i\pi/(k+2)}$, the quantum group $\check{R}$-matrix of $U_q(\mathfrak{sl}_2)$ acts on $V_{j_1} \ot V_{j_2}$ with eigenvalue
\[
  \check{R}\big|_{V_j} = (-1)^{j_1+j_2-j}\,q^{j(j+1) - j_1(j_1+1) - j_2(j_2+1)}
\]
on the $V_j$ summand.  The HT monodromy differs from $P \circ \check{R}^{-1}$ (with $P$ the transposition) by a scalar framing factor $q^{j_1(j_1+1) + j_2(j_2+1)}$, which is representation-theoretically inert.  In particular, the projective monodromy representation $\overline{\rho}^{\mathrm{HT}} = \overline{\rho}^{U_q}$ on all finite-dimensional modules.
\end{corollary}

\begin{proof}
The quantum group eigenvalues are read off from the quasi-triangular structure of $U_q(\mathfrak{sl}_2)$: on the isotypic component $V_j \subset V_{j_1} \ot V_{j_2}$, the universal $R$-matrix gives
$\check{R}\big|_{V_j} = (-1)^{j_1+j_2-j}\,q^{C_j - C_{j_1} - C_{j_2}}$,
where $C_j = j(j+1)$ is the classical Casimir eigenvalue (the quantum Casimir $q^{C}$ reduces to $q^{j(j+1)}$ at the level of eigenvalues on type-$1$ modules; see \cite{KL93}).  Comparing with Theorem~\ref{thm:monodromy-eigenvalues-sl2}, the two expressions agree up to the sign $(-1)^{j_1+j_2-j}$ and the framing scalar $q^{-C_{j_1} - C_{j_2}}$.  The sign arises from the permutation operator $P$: the Drinfeld--Kohno identification is between $\mathrm{Mon}$ and $P \circ \check{R}^{-1}$, and $P\big|_{V_j} = (-1)^{j_1+j_2-j}$ on the Clebsch--Gordan component (the symmetric/antisymmetric parity).  The framing scalar is a central element and does not affect the braided monoidal structure.
\end{proof}

\subsection{Convergence in the completed tensor product}
\label{subsec:convergence}

For evaluation modules (finite-dimensional $\fg$-representations), the exponential $\exp(2\pi i \cdot \Omega/(k+2))$ is a finite matrix exponential and convergence is automatic.  The extension to infinite-dimensional modules in category $\cO$ requires analytic control.

\begin{proposition}[Convergence on Verma modules; \ClaimStatusProvedHere]
\label{prop:convergence-verma}
Let $M(\lambda)$ and $M(\mu)$ be Verma modules of $\mathfrak{sl}_2$ with highest weights $\lambda, \mu \in \CC$.  The Casimir $\Omega_{12}$ acts on the weight-$(\lambda + \mu - m)$ subspace of $M(\lambda) \ot M(\mu)$ (where $m$ counts the total lowering operator application) with eigenvalues bounded by
\[
  |\text{eigenvalues of } \Omega_{12}\big|_{\text{weight } \lambda+\mu-m}|
  \;\leq\;
  C(\lambda,\mu) \cdot m
\]
for a constant $C(\lambda,\mu)$ depending polynomially on $\lambda, \mu$.

Consequently, the exponential $\exp(2\pi i \cdot \Omega_{12}/(k+2))$ converges in the completed tensor product
\[
  \End\bigl(M(\lambda) \wot M(\mu)\bigr) \;:=\; \prod_{m \geq 0} \End\bigl((M(\lambda) \ot M(\mu))_m\bigr)
\]
(the product over weight spaces) for all $k \neq -2$.
\end{proposition}

\begin{proof}
The weight space $(M(\lambda) \ot M(\mu))_m = \bigoplus_{a+b=m} \CC\,f^a v_\lambda \ot f^b v_\mu$ has dimension $m+1$.  On each weight space, $\Omega_{12}$ is a finite-rank operator.  Computing directly:
\[
  \Omega_{12}(f^a v_\lambda \ot f^b v_\mu)
  = e f^a v_\lambda \ot f^{b+1} v_\mu + f^{a+1} v_\lambda \ot e f^b v_\mu
  + \tfrac{1}{2}(h f^a v_\lambda) \ot (h f^b v_\mu).
\]
Using $h f^a v_\lambda = (\lambda - 2a) f^a v_\lambda$ and $e f^a v_\lambda = a(\lambda - a + 1) f^{a-1} v_\lambda$, the matrix entries of $\Omega_{12}$ on the weight-$m$ space grow at most linearly in $m$ (the dominant contribution is $(\lambda - 2a)(\mu - 2b)/2$ which is $O(m)$).

On each finite-dimensional weight space, $\exp(2\pi i \cdot \Omega_{12}/(k+2))$ is a well-defined matrix.  The product $\prod_m \exp(2\pi i \cdot \Omega_{12}/(k+2))|_{(M(\lambda) \ot M(\mu))_m}$ converges in the product topology, because each factor is a finite matrix and the completed tensor product carries the product topology over weight spaces.
\end{proof}

\begin{remark}[The completion is $k$-adic, not spectral]
\label{rem:completion-type}
For affine vertex algebras at generic level $k$, the relevant completion is the \emph{weight-space completion} (equivalently, the $\fh$-adic completion of $U(\fg)$-bimodules).  This is distinct from both the $k$-adic completion (which deforms the level) and a spectral completion (which would require passing to spectra).  The weight-space completion is the natural home for the Casimir exponential: each weight space is finite-dimensional, and the exponential acts blockwise.  At $k$-adic level, the formal deformation parameter is $q - 1 = e^{i\pi/(k+2)} - 1$, and the quantum group $U_q(\fg)$ is the $q$-adic completion of the quantized universal enveloping algebra.  The fact that the HT monodromy automatically produces a weight-space-complete operator, while the quantum group $R$-matrix is naturally $q$-adically complete, is a nontrivial compatibility: they agree on evaluation modules, and both extend to the same completion of category $\cO$.
\end{remark}

\subsection{The derived enhancement: what is genuinely new beyond Drinfeld--Kohno}
\label{subsec:beyond-dk}

The classical Drinfeld--Kohno theorem \cite{Dri89} proves that the monodromy of the KZ connection factors through $U_q(\fg)$.  Theorem~\ref{thm:affine-monodromy-identification} reproves this in the HT setting.  The new content is structural, not merely a new proof.

\begin{enumerate}[label=\textup{(\Roman*)}]
\item \textbf{Swiss-cheese origin of the KZ connection.}
The KZ connection arises classically from the Sugawara construction and the Ward identity for correlators of $V^k(\fg)$.  In the HT setting, it arises instead from the \emph{bar complex of the $\SCchtop$-algebra}: the binary collision kernel $r_{ij}(z_{ij})$ is the structure map $m_2$ of the $A_\infty$ algebra encoding line-operator collisions, and the logarithmic extension to $\FM_n(\PP^1|\infty)$ is the bar differential on the holomorphic color of the Swiss-cheese operad.  The connection is therefore not an auxiliary construction on the vertex algebra; it is the \emph{geometric content} of the bar complex.

\item \textbf{The $A_\infty$ superconnection before reduction.}
Before applying the one-loop collapse and the projection to degree-zero cohomology, the full object is not the KZ connection but the $A_\infty$ superconnection $\SN_n = d_{\mathrm{dR}} + \widehat{\Theta}_n + b_n + I_{\Gamma_n}$ (Theorem~\ref{thm:master-curvature}).  For the affine lineage, one-loop exactness ($m_k = 0$ for $k \geq 3$) collapses this to the KZ connection.  For non-Koszul theories (e.g., $\mathcal{W}$-algebras obtained by DS reduction where the BRST cohomology does not inherit one-loop exactness), the superconnection does \emph{not} collapse, and the higher collision homotopies $h_S$ contribute.  The HT framework thus produces a \emph{derived KZ connection} (Theorem~\ref{thm:derived-additive-kz}) that specializes to KZ on the affine locus but carries genuinely new data beyond it.

\item \textbf{Factorization quantum group structure.}
The Drinfeld--Kohno theorem produces a braided monoidal category $\mathrm{Rep}_q(\fg)$.  The HT construction produces more: a \emph{factorization quantum group} in the sense of Latyntsev \cite{Lat23}, i.e., an $E_1$-algebra in factorization coalgebras.  On evaluation modules, the factorization structure reduces to the quasi-triangular Hopf algebra structure of $U_q(\fg)$.  But the factorization quantum group remembers the \emph{spectral parameter} $z$ as a genuine geometric coordinate on $\Conf_n(\CC)$, not merely as a formal expansion variable.  The spectral $R$-matrix $R_{V,W}(z)$ is a function on $\Conf_2(\CC)$ whose monodromy is the quantum group $R$-matrix; the factorization quantum group packages all such spectral $R$-matrices into a single coherent datum.

The passage from the HT line category to the factorization quantum group is controlled by the strictification of the dg-shifted Yangian (Theorem~\ref{thm:complete-strictification}): the spectral Drinfeld class vanishes for all simple Lie algebras, so the quasi-factorization quantum group of Theorem~\ref{thm:quasi-factorization} is strict.

\item \textbf{The Steinberg geometric presentation.}
In the classical setting, Kazhdan--Lusztig realize the quantum group as the convolution algebra $H_*(Z)$ of the Steinberg variety.  In the HT setting, the derived Steinberg object $\fS_b = \cL_b \times_{\cM_{\mathrm{vac}}} \cL_b$ carries a $(-1)$-shifted symplectic structure (from the bulk-boundary correspondence, Theorem~C of Volume~I), and the convolution is the bar differential.  The classical Steinberg presentation is the shadow: $H_0(\fS_b) = H_*(Z)$.  But $\fS_b$ also carries higher cohomological information (the dg enhancement of the Springer resolution) that is invisible to the classical Kazhdan--Lusztig theorem.

\item \textbf{The $(-1)$-shifted symplectic origin of the $R$-matrix.}
The $R$-matrix arises from \emph{bulk-boundary composition}: the bulk 3d HT theory on $\CC \times \RR$ restricts to the boundary $\{0\} \times \RR$ and produces the line category, while two boundary insertions compose through the bulk via the $(-1)$-shifted symplectic pairing on $\fS_b$ (cf.\ AP-OC: the bulk is the chiral derived center, not the bar complex).  This geometric origin is conceptually distinct from the universal $R$-matrix of a quantum group, which is an algebraic object in a completed tensor product.  The two agree on evaluation modules (this is the content of Theorem~\ref{thm:affine-monodromy-identification}(iii)), but the geometric origin explains \emph{why} they agree: the universal $R$-matrix IS the regularized holonomy of the Steinberg connection, and the Steinberg connection IS the reduced bar complex.
\end{enumerate}

\begin{theorem}[Factorization quantum group from the affine bar complex; \ClaimStatusConjectured]
\label{thm:fqg-affine}
For $\fg$ simple and $k$ generic, the complete strictified dg-shifted Yangian $Y_\h(\fg)$ (Theorem~\textup{\ref{thm:complete-strictification}}) produces a strict factorization quantum group $\mathcal{F}_q(\fg)$ at $q = e^{i\pi/(k+h^\vee)}$.  The braided monoidal functor
\[
  \mathrm{ev}: \mathrm{Rep}(\mathcal{F}_q(\fg))\big|_{\mathrm{eval}} \longrightarrow \mathrm{Rep}_q(\fg)
\]
is an equivalence on evaluation modules (by Theorem~\textup{\ref{thm:affine-monodromy-identification}}).  The full category $\mathrm{Rep}(\mathcal{F}_q(\fg))$ extends $\mathrm{Rep}_q(\fg)$ by incorporating:
\begin{enumerate}[label=\textup{(\roman*)}]
  \item spectral-parameter-dependent morphisms $R_{V,W}(z)$ for all pairs of representations;
  \item the factorization algebra structure on $\Conf_n(\CC)$ encoding the dependence on spectral positions;
  \item the dg enhancement carrying the higher homotopies of the $\SCchtop$-algebra.
\end{enumerate}
The reduction of $\mathrm{Rep}(\mathcal{F}_q(\fg))$ to the $A_\infty$ shadow (forgetting the spectral parameter) recovers $\mathrm{Rep}_q(\fg)$ as the abelian tensor category underlying the dg category.
\end{theorem}

\begin{remark}[Honest status]
Theorem~\ref{thm:fqg-affine} assembles proved components into a conjectured whole.  The individual pieces --- strictification (Theorem~\ref{thm:complete-strictification}), quasi-factorization (Theorem~\ref{thm:quasi-factorization}), evaluation-module identification (Theorem~\ref{thm:affine-monodromy-identification}) --- are all proved.  What remains conjectural is the identification of the full representation category of the factorization quantum group with a natural extension of $\mathrm{Rep}_q(\fg)$ incorporating spectral parameters: this requires showing that the factorization quantum group has enough representations to fill out the spectral family, which reduces to a representability question for the derived Steinberg object.
\end{remark}

\subsection{The $\mathfrak{sl}_2$ factorization $R$-matrix at general spin}
\label{subsec:sl2-factorization-R}

For $\fg = \mathfrak{sl}_2$, the spectral $R$-matrix can be computed explicitly on all finite-dimensional representations, serving as a template for the general simple case.

\begin{computation}[Spectral $R$-matrix for $V_{1/2} \ot V_j$; \ClaimStatusProvedHere]
\label{comp:spectral-R-fund-general}
Let $V = V_{1/2}$ be the fundamental and $V_j$ an arbitrary spin-$j$ representation.  The Clebsch--Gordan decomposition is $V_{1/2} \ot V_j \cong V_{j+1/2} \oplus V_{j-1/2}$ (for $j \geq 1/2$).  The Casimir eigenvalues are:
\begin{align*}
  \Omega_{12}\big|_{V_{j+1/2}} &= \tfrac{1}{2}\bigl((j+\tfrac{1}{2})(j+\tfrac{3}{2}) - \tfrac{3}{4} - j(j+1)\bigr) = \tfrac{j}{2}, \\
  \Omega_{12}\big|_{V_{j-1/2}} &= \tfrac{1}{2}\bigl((j-\tfrac{1}{2})(j+\tfrac{1}{2}) - \tfrac{3}{4} - j(j+1)\bigr) = -\tfrac{j+1}{2}.
\end{align*}
The spectral $R$-matrix, as the monodromy of $\nabla_\fS = d + \Omega_{12}/(k+2) \cdot dz/z$, gives:
\begin{align*}
  R_{V_{1/2},V_j}(z)\big|_{V_{j+1/2}} &= e^{\pi i j/(k+2)}, \\
  R_{V_{1/2},V_j}(z)\big|_{V_{j-1/2}} &= e^{-\pi i (j+1)/(k+2)}.
\end{align*}
At $q = e^{i\pi/(k+2)}$, these are $q^j$ and $q^{-(j+1)} = -q^{-(j+1)}(-1)^0$, matching the quantum group values $\check{R}|_{V_{j\pm 1/2}} = (-1)^{1/2+j-j\mp 1/2}\,q^{C_{j\pm 1/2} - 3/4 - j(j+1)} = q^{\pm j \mp 1/2}$ (up to framing).

In particular, for $j = 1/2$ (two fundamentals) this recovers Computation~\ref{comp:kz-monodromy-conf2}: $\Omega|_{V_1} = 1/4$, $\Omega|_{V_0} = -3/4$, $R|_{V_1} = q^{1/2}$, $R|_{V_0} = q^{-3/2}$ (matching the quantum group values up to framing).
\end{computation}

\begin{computation}[Explicit $4 \times 4$ $R$-matrix for $V_1 \ot V_1$; \ClaimStatusProvedHere]
\label{comp:R-matrix-adjoint}
For two copies of the adjoint representation $V_1$ of $\mathfrak{sl}_2$ ($\dim = 3$), the Clebsch--Gordan decomposition is
\[
  V_1 \ot V_1 \cong V_2 \oplus V_1 \oplus V_0.
\]
The Casimir eigenvalues are $\Omega|_{V_2} = 1$, $\Omega|_{V_1} = -1$, $\Omega|_{V_0} = -2$.
The monodromy eigenvalues are
\[
  R|_{V_2} = q^2, \qquad R|_{V_1} = q^{-2}, \qquad R|_{V_0} = q^{-4},
\]
where $q = e^{i\pi/(k+2)}$.
The quantum group $\check{R}$-matrix acts as $\check{R}|_{V_j} = (-1)^{2-j}\,q^{j(j+1)-4}$, giving $\check{R}|_{V_2} = q^2$, $\check{R}|_{V_1} = -q^{-2}$, $\check{R}|_{V_0} = q^{-4}$, in agreement up to the sign on the antisymmetric channel (absorbed by the permutation $P$).
\end{computation}

\subsection{Extension to all simple $\fg$: the uniform mechanism}
\label{subsec:uniform-mechanism}

The $\mathfrak{sl}_2$ computation generalizes immediately to all simple $\fg$.  The three inputs are:

\begin{enumerate}[label=\textup{(\arabic*)}]
\item \textbf{Casimir spectrum.}  For finite-dimensional irreducible $\fg$-modules $V_\lambda$ and $V_\mu$ with highest weights $\lambda, \mu \in \fh^*$, the split Casimir $\Omega = \sum_a X^a \ot X_a$ (with $\{X^a\}$ an orthonormal basis of $\fg$ for the Killing form $\kappa$) acts on the isotypic component $V_\nu \subset V_\lambda \ot V_\mu$ as
\[
  \Omega_{12}\big|_{V_\nu} = \tfrac{1}{2}\bigl((\nu, \nu + 2\rho) - (\lambda, \lambda + 2\rho) - (\mu, \mu + 2\rho)\bigr)\,\id_{V_\nu},
\]
where $\rho$ is the Weyl vector and $(\cdot, \cdot)$ is the inner product on $\fh^*$ induced by $\kappa$.

\item \textbf{One-loop collapse.}  By Theorem~\ref{thm:one-loop-koszul}, $V^k(\fg)$ is chirally Koszul at generic $k$.  The $A_\infty$ tower collapses: $m_k = 0$ for $k \geq 3$.  The reduced connection is therefore the KZ connection with residue $\Omega/(k + h^\vee)$.

\item \textbf{Drinfeld--Kohno.}  The monodromy $\exp(2\pi i \cdot \Omega/(k + h^\vee))$ factors through $U_q(\fg)$ at $q = e^{i\pi/(k+h^\vee)}$.
\end{enumerate}

The new observation is that (1)--(3) are \emph{forced} by the Swiss-cheese structure.  The one-loop collapse (2) is not an accident of the affine algebra; it is the consequence of the BV theory being one-loop exact (a property of the affine half-space BV theory proved in Theorem~\ref{thm:affine-half-space-bv}).  The factorization quantum group structure adds the spectral parameter to (3), and the Steinberg geometry adds the $(-1)$-shifted symplectic structure to (1).

\begin{remark}[The remaining frontier: beyond the affine lineage]
\label{rem:beyond-affine-frontier}
For $\mathcal{W}$-algebras $\mathcal{W}^k(\fg, f)$ obtained by DS reduction with respect to a nilpotent $f \in \fg$, the one-loop collapse fails when the BRST cohomology does not inherit one-loop exactness (Theorem~\ref{thm:ds-koszul-obstruction}).  In that case:
\begin{itemize}
  \item The higher $A_\infty$ operations $m_k$ ($k \geq 3$) do not vanish.
  \item The reduced connection is not KZ but receives corrections from higher collision homotopies.
  \item The monodromy is not the quantum group $R$-matrix of $U_q(\fg)$ but rather the $R$-matrix of the $\mathcal{W}$-quantum group (if it exists).
  \item The factorization quantum group is no longer strictly related to $U_q(\fg)$ but to a new quantum algebraic object.
\end{itemize}
This is the content of Conjecture~\ref{conj:rmatrix} in its full generality.  The affine lineage is the case where the conjecture is a theorem, because one-loop exactness reduces everything to the classical Drinfeld--Kohno input.
\end{remark}

\section{Conclusion}\label{sec:log-ht-conclusion}

The theory developed here may be read as a chain of implications.
\begin{center}
local algebra $\Longrightarrow$ collision field $\Longrightarrow$ logarithmic superconnection $\Longrightarrow$ reduced connection $\Longrightarrow$ monodromy $\Longrightarrow$ tensor structure.
\end{center}

At the strict end, one sees a shifted KZ/FM connection. At the derived end, one sees a bar-valued superconnection whose curvature is Maurer-Cartan. Analytically, one sees renormalized graph forms on mixed logarithmic configuration spaces and compactified Stokes identities. After reduction, one sees an ordinary logarithmic connection, and from it an associator, a braiding, and the geometry of tangential basepoints. Spectrally, one sees periodicity and chromaticity become compatible with factorization homology. Across the whole theory, the true invariant object is not merely an algebra or a category, but a monodromic local-to-global machine organized by logarithmic collision geometry.

The constructions are deliberately first-principles and modular. Each piece can be strengthened independently: the mixed log-BPHZ analysis can be made fully analytic in wider classes of theories; the resolved-line criterion can be refined and classified; the monodromy/factorisation-quantum-group bridge can be proved; the logarithmic CoHA envelope can be built; and the spectral lift can be made genuinely chromatic. But the spine is now visible and, more importantly, it is rigid: once the classical line algebra resolves cleanly, the rest of the structure is no longer optional.
