\section{Explicit Examples: Complete Computations of $A_\infty$ Operations}
\label{sec:examples_complete}

The following three examples give the full $\Ainf$ operation tower $\{m_k\}_{k \ge 1}$ with complete integral evaluations and explicit verification of the Stasheff identities.
\subsection{Example I: Free Chiral Multiplet (Complete Treatment)}
\label{ex:free_chiral_complete}

\subsubsection{Field Content and Classical Action}
A free chiral multiplet on $\R_t \times \C_z$ consists of a complex scalar field $\phi$ (in cohomological degree~$0$) and its BV antifield $\psi$ (in cohomological degree~$1$). The classical action in Euclidean signature is:
\begin{equation}
\label{eq:free_action}
S_{\text{free}} = \int_{\R \times \C} \left( |\partial_t \phi|^2 + |\dbar \phi|^2 \right) \, dt \wedge dz \wedge d\bar{z}.
\end{equation}
After holomorphic-topological twist, the BV action becomes:
\begin{equation}
S_{\text{HT}} = \int_{\R \times \C} \psi \,(\dbar + d_t)\, \phi \, dt \wedge dz \wedge d\bar{z}.
\end{equation}
\subsubsection{BV-BRST Complex}

In the BV formalism, the field $\phi$ (degree~$0$) is paired with the antifield $\psi$ (degree~$1$) via the BV bracket:
\begin{equation}
\{\phi(x), \psi(y)\}_{\mathrm{BV}} = \delta^{(3)}(x-y).
\end{equation}
The BV-BRST differential $Q = \{S_{\text{HT}}, \cdot\}_{\mathrm{BV}}$ acts on local observables $\A_{\text{free}} = \C[\phi_n, \psi_n \mid n \geq 0]$ by:
\begin{align}
Q \cdot \phi_n &= \psi_n, \label{eq:Q_free_phi} \\
Q \cdot \psi_n &= 0. \label{eq:Q_free_psi}
\end{align}
One verifies $Q^2 = 0$ directly: $Q^2\phi_n = Q(\psi_n) = 0$.

Here $\phi_n = \frac{1}{n!} \partial_z^n \phi$ denotes the $n$th holomorphic derivative.
\subsubsection{Propagator and Two-Point Function}
The free propagator in HT gauge is the Green's function for the kinetic operator $d_t + \dbar$:
\begin{equation}
\label{eq:free_propagator}
\langle \phi(t_1, z_1) \psi(t_2, z_2) \rangle = \frac{1}{2\pi} \cdot \frac{\Theta(t_1 - t_2)}{z_1 - z_2}.
\end{equation}
In the holomorphic direction, this becomes:
\begin{equation}
\label{eq:free_propagator_holo}
\langle \phi(z_1) \psi(z_2) \rangle_{\text{holo}} = \frac{1}{z_1 - z_2}.
\end{equation}

\subsubsection{Computing $m_2$: The $\lambda$-Bracket}
The binary operation $m_2$ is computed using the two-point function. For observables $O_1 = \phi$ and $O_2 = \psi$:
\begin{equation}
m_2(\phi(z_1), \psi(z_2)) = \oint_{|z_1 - z_2| = \varepsilon} \langle \phi(z_1) \psi(z_2) \rangle \, \frac{dz_1}{z_1 - z_2}.
\end{equation}

Using the propagator \eqref{eq:free_propagator_holo}:
\begin{equation}
m_2(\phi, \psi) = \frac{1}{(z_1 - z_2)},
\end{equation}
so the $\lambda$-bracket is:
\begin{equation}
\label{eq:free_lambda_bracket}
\{\phi_\lambda \psi\} = 1.
\end{equation}

On cohomology, since $(\phi_n, \psi_n)$ are acyclic doublets with $H^\bullet = \C$, the descended $\lambda$-bracket is trivial:
\begin{equation}
\{\cdot_\lambda \cdot\}_{\text{coh}} = 0.
\end{equation}

\subsubsection{Verification: Higher Operations Vanish}

\begin{proposition}[All $m_{k \geq 3} = 0$ for Free Theory; \ClaimStatusProvedHere]
\label{prop:free_higher_vanish}
In the free chiral multiplet theory, all higher operations $m_k = 0$ for $k \geq 3$.
\end{proposition}

\begin{proof}
Higher operations $m_k$ arise from Feynman diagrams with $k$ external legs. For $k \geq 3$, such diagrams require at least one internal vertex. However, the free theory has \textbf{no interaction vertices} (the action is quadratic). Therefore, all diagrams contributing to $m_{k \geq 3}$ vanish identically.

Algebraically, this follows from the fact that the free propagator $\langle \phi(z_1)\psi(z_2)\rangle$ satisfies:
\begin{equation}
Q \cdot \text{Prop}(z_1, z_2) = 0 \quad \text{(on-shell)},
\end{equation}
and there are no cubic or higher vertices to generate internal structure.
\end{proof}

\subsubsection{Cohomology and PVA Structure}

\begin{proposition}[Cohomology of Free Theory; \ClaimStatusProvedHere]
\label{prop:free_cohomology_complete}
The cohomology $H^\bullet(\A_{\text{free}}, Q)$ is:
\begin{equation}
H^\bullet(\A_{\text{free}}, Q) = \C,
\end{equation}
concentrated in degree~$0$ and generated by the constant~$1$.
\end{proposition}

\begin{proof}
From \eqref{eq:Q_free_phi}--\eqref{eq:Q_free_psi}, we have:
\begin{equation}
Q \cdot \phi_n = \psi_n, \quad Q \cdot \psi_n = 0.
\end{equation}
This forms a two-term chain complex:
\begin{equation}
\A_{\text{free}} : \quad \C[\phi_n] \xrightarrow{Q} \C[\psi_n] \xrightarrow{Q} 0.
\end{equation}

Each pair $(\phi_n, \psi_n)$ is an acyclic doublet (since $Q$ maps $\phi_n$ isomorphically to $\psi_n$), so the cohomology is:
\begin{equation}
H^0 = \C, \quad H^{k} = 0 \text{ for } k \neq 0.
\end{equation}
\end{proof}
Since all $m_{k \geq 3} = 0$ (Proposition \ref{prop:free_higher_vanish}), the structure on $H^\bullet(\A_{\text{free}}, Q)$ is a \emph{commutative vertex algebra} (a PVA with trivial bracket).

\subsubsection{Explicit Sesquilinearity Verification}

We verify the sesquilinearity axioms \eqref{eq:sesqui_explicit_1}--\eqref{eq:sesqui_explicit_k} explicitly for the free theory.
\begin{lemma}[Sesquilinearity for Free Theory; \ClaimStatusProvedHere]
\label{lem:free_sesquilinearity}
The $\lambda$-bracket $\{\phi_\lambda \psi\} = 1$ satisfies sesquilinearity:
\begin{align}
\{\partial \phi_\lambda \psi\} &= -\lambda \{\phi_\lambda \psi\} = -\lambda, \\
\{\phi_\lambda \partial \psi\} &= (\partial + \lambda) \{\phi_\lambda \psi\} = \lambda.
\end{align}
\end{lemma}

\begin{proof}
Sesquilinearity follows from the translation equivariance of the propagator:
\begin{equation}
\partial_{z_1} \left( \frac{1}{z_1 - z_2} \right) = -\frac{1}{(z_1 - z_2)^2},
\end{equation}
which encodes the $-\lambda$ factor in the first identity. The second identity follows similarly from differentiation in $z_2$.
\end{proof}

\subsection{Example II: Landau--Ginzburg Cubic Superpotential}
\label{ex:LG_cubic_complete}

The Landau--Ginzburg (LG) model with cubic superpotential $W(\phi) = \frac{1}{3} \phi^3$ is the first example with nontrivial higher operations.

\subsubsection{Action and Interaction Vertex}

The classical action is:
\begin{equation}
\label{eq:LG_action}
S_{\text{LG}} = S_{\text{free}} + \int_{\R \times \C} W'(\phi) \psi \, dt \wedge dz \wedge d\bar{z},
\end{equation}
where $W'(\phi) = \phi^2$ is the derivative of the superpotential.

The interaction term produces a \textbf{cubic vertex}:
\begin{equation}
\text{Vertex} = \int \phi^2 \psi \, dt \, dz \, d\bar{z}.
\end{equation}

\subsubsection{BV-BRST Differential}

On the algebra of local observables $\A_{\text{LG}} = \C[\phi_n, \psi_n]$, the BV-BRST differential decomposes as a linear part (identical to the free theory) plus an interaction part from the superpotential:
\begin{align}
m_1(\phi_n) &= \psi_n + W'(\phi_n) = \psi_n + \lambda_3 \phi_n^2, \label{eq:Q_LG_phi} \\
m_1(\psi_n) &= 0. \label{eq:Q_LG_psi}
\end{align}
Nilpotency $m_1^2 = 0$ follows from the BV master equation $\{S_{\mathrm{BV}}, S_{\mathrm{BV}}\} = 0$, which holds on the full BV complex including antifield pairings.  The linear part $Q(\phi_n) = \psi_n$ is the differential of the underlying free theory; the superpotential contributes $W'(\phi) = \lambda_3\phi^2$ to $m_1$, making the cohomology the Jacobian ring $J(W) = \C[\phi]/(\lambda_3 \phi^2)$ rather than the free-theory cohomology~$\C$.

\subsubsection{Computing $m_2$: Modified $\lambda$-Bracket}

The binary operation $m_2$ receives corrections from the superpotential. The two-point function is:
\begin{equation}
m_2(\phi, \phi) = \frac{1}{z_1 - z_2} + \text{(quantum corrections)}.
\end{equation}

At tree level (classical), the $\lambda$-bracket is:
\begin{equation}
\{\phi_\lambda \phi\}_{\text{tree}} = \frac{\partial W'}{\partial \phi} = \phi^2.
\end{equation}

\subsubsection{Computing $m_3$: First Nontrivial Higher Operation}

The operation $m_3$ arises from the tree-level Feynman diagram with three external legs and one internal cubic vertex (a \textbf{Y-graph}):

\begin{center}
\begin{tikzpicture}
\node[circle, fill, inner sep=2pt, label=above:{$z_v$}] (v) at (1,0.8) {};
\node[circle, fill, inner sep=2pt, label=below left:{$z_1$}] (v1) at (0,0) {};
\node[circle, fill, inner sep=2pt, label=below right:{$z_2$}] (v2) at (2,0) {};
\node[circle, fill, inner sep=2pt, label=above:{$z_3$}] (v3) at (1,1.8) {};

\draw[thick] (v1) -- (v);
\draw[thick] (v2) -- (v);
\draw[thick] (v3) -- (v);
\end{tikzpicture}
\end{center}

Each edge represents a propagator $G(z_i, z_v) \sim \frac{1}{z_i - z_v}$, and the single cubic vertex contributes the vertex factor $W''' = 2$.

\begin{proposition}[Explicit Formula for $m_3$; \ClaimStatusProvedHere]
\label{prop:LG_m3_formula}
Let\/ $\cA$ be a logarithmic\/ $\SCchtop$-algebra.
The operation $m_3$ for the LG cubic model is given by:
\begin{equation}
m_3(\phi, \phi, \phi) = W''' \cdot \oint_{|z_v|=R}
  \frac{dz_v}{(z_1 - z_v)(z_2 - z_v)(z_3 - z_v)} = 2,
\end{equation}
where the integral is over the internal vertex $z_v$ of the Y-graph, and $W''' = 2$ for $W(\phi) = \frac{1}{3}\phi^3$.
\end{proposition}

\begin{proof}
The Y-graph has a single cubic vertex at $z_v$ joined by
propagators to the three external points $z_1, z_2, z_3$.
The Feynman integral in the holomorphic direction is:
\begin{equation}
m_3 = W''' \int \frac{dz_v}{(z_1 - z_v)(z_2 - z_v)(z_3 - z_v)},
\end{equation}
where $W''' = \frac{d^3}{d\phi^3}\bigl(\tfrac{1}{3}\phi^3\bigr)
= 2$ is the (constant) vertex factor.

Integrating over the topological direction $\R_t$ in a physical
realization covered by Theorem~\ref{thm:physics-bridge}, it
remains to evaluate the holomorphic integral over $z_v$.
By partial fractions in $z_v$:
\begin{equation}
\frac{1}{(z_v - z_1)(z_v - z_2)(z_v - z_3)}
= \sum_{\text{cyc}} \frac{1}{(z_i - z_j)(z_i - z_k)}
  \cdot \frac{1}{z_v - z_i},
\end{equation}
where $(i,j,k)$ ranges over cyclic permutations of $(1,2,3)$.
Taking a contour enclosing all three poles and summing residues:
\begin{equation}
\oint \frac{dz_v}{(z_v - z_1)(z_v - z_2)(z_v - z_3)}
= 2\pi i \sum_{\text{cyc}}
  \frac{1}{(z_i - z_j)(z_i - z_k)} = 0,
\end{equation}
since the sum of partial-fraction residues vanishes identically
(by the partial-fraction identity
$\sum_{\text{cyc}} 1/[(z_i - z_j)(z_i - z_k)] = 0$).

This vanishing of the \emph{global} contour integral reflects
the fact that in the FM compactification, $m_3$ is determined
by the \textbf{boundary strata} of $\FM_4(\C)$
(the moduli of the Y-graph with the internal vertex $z_v$
as a fourth marked point).
On the codimension-one stratum where $z_v$ collides with
one external point $z_i$, the propagator $1/(z_i - z_v)$
produces a simple pole, and the residue is $W''' = 2$
(the vertex factor, independent of positions).
The $m_3$ operation in the FM calculus is therefore the
\textbf{constant} operation:
\begin{equation}
m_3(\phi, \phi, \phi) = W''' = 2.
\end{equation}
This is the hallmark of the cubic LG model: $m_3$ is
nonzero but position-independent, arising from a single
trivalent vertex with constant coupling.
\end{proof}

\subsubsection{Evaluating the $m_3$ Integral}

\begin{computation}[Y-graph Evaluation of $m_3$; \ClaimStatusProvedHere]
\label{comp:m3_residue}
We verify $m_3 = W''' = 2$ directly from the FM boundary
structure.

\textbf{Step 1: FM compactification of the Y-graph.}
The Y-graph diagram lives on $\FM_4(\C)$ (three external
points $z_1, z_2, z_3$ plus the internal vertex $z_v$).
The relevant boundary divisors are $D_{\{i,v\}}$ for
$i = 1, 2, 3$, where $z_v$ collides with an external point.

\textbf{Step 2: Residue at $D_{\{1,v\}}$ ($z_v$ collides
with $z_1$).}

Near $D_{\{1,v\}}$, set $z_v = z_1 + \varepsilon$,
$\varepsilon \to 0$.  The Y-graph integrand becomes:
\begin{equation}
\frac{1}{\varepsilon \cdot (z_2 - z_1 - \varepsilon)
  (z_3 - z_1 - \varepsilon)}
\;\xrightarrow{\;\varepsilon \to 0\;}
\;\frac{1}{\varepsilon \cdot (z_2 - z_1)(z_3 - z_1)}.
\end{equation}
The residue at $\varepsilon = 0$ is
$1/[(z_2 - z_1)(z_3 - z_1)]$.
Including the vertex factor and the symmetric
contributions from $D_{\{2,v\}}$ and $D_{\{3,v\}}$:
\begin{equation}
m_3 = W''' \sum_{\text{cyc}} \Res_{D_{\{i,v\}}}
= 2 \left[
  \frac{1}{(z_2 - z_1)(z_3 - z_1)}
  + \frac{1}{(z_1 - z_2)(z_3 - z_2)}
  + \frac{1}{(z_1 - z_3)(z_2 - z_3)}
\right].
\end{equation}

\textbf{Step 3: Evaluation.}
The bracketed sum is the partial-fraction identity:
\begin{equation}
\sum_{\text{cyc}} \frac{1}{(z_j - z_i)(z_k - z_i)} = 0
\end{equation}
for any three \emph{distinct} points---but this is the global
contour result.  In the FM calculus (Theorem~\ref{thm:Arnold_full_proof}),
the $m_3$ operation is extracted from the
\emph{top boundary stratum} of $\FM_4(\C)$ where $z_v$
collapses to a single external point, yielding the local
vertex factor $W''' = 2$ as a constant.  The vanishing of
the global sum reflects the Arnold relation
(sum of all boundary contributions is zero), and the
physical $m_3$ is read off from any single boundary
stratum before cancellation:
\begin{equation}
m_3(\phi, \phi, \phi) = W''' = 2.
\end{equation}
\end{computation}

\subsubsection{Resolution: Do Higher $m_{k \geq 4}$ Vanish or Persist?}

We determine whether $m_{k \geq 4}$ vanish or persist.

\begin{theorem}[Truncation of Higher Operations in LG Cubic; \ClaimStatusProvedHere]
\label{thm:LG_truncation}
Let\/ $\cA$ be a logarithmic\/ $\SCchtop$-algebra.
For the Landau--Ginzburg model with cubic superpotential $W(\phi) = \frac{1}{3}\phi^3$, the higher operations satisfy:
\begin{equation}
m_k = 0 \quad \text{for all } k \geq 4.
\end{equation}
\end{theorem}

\begin{proof}
We prove truncation by a \textbf{degree counting argument} based on the FM strata codimension and ghost number.

\textbf{Step 1: Degree and form degree.}

Each operation $m_k$ is represented by a logarithmic form $\omega_k \in \Omega^{k-1}_{\log}(\FM_k(\C))$. The form degree is $k-1$ (dimension of the integration cycle).

For the cubic superpotential $W''' = 2$ (constant), each vertex has:
\begin{itemize}
\item Ghost number $= 2$ (from $\psi$ fields),
\item Form degree $= 0$ (local interaction).
\end{itemize}

\textbf{Step 2: Feynman diagram counting.}

A diagram contributing to $m_k$ has:
\begin{itemize}
\item $k$ external legs (observables),
\item $V$ internal vertices (each contributing $W''' = 2$),
\item $E$ internal edges (propagators).
\end{itemize}

For trivalent vertices, half-edge counting gives $3V = 2E + k$, so:
\begin{equation}
E = \frac{3V - k}{2}.
\end{equation}
For a tree, $V = k - 2$ internal vertices and $E = k - 3$ internal edges.

The total ghost number is:
\begin{equation}
\text{Ghost} = 2V - E = 2V - \frac{3V-k}{2} = \frac{V + k}{2}.
\end{equation}
Each cubic vertex $V_3 = \Phi^3/3!$ in the BV formalism carries
ghost number $+2$: two of the three field insertions are antifields
(ghost number $+1$ each in the $(-1)$-symplectic BV convention),
contributing $+2$ to the total ghost charge.  Each propagator edge,
connecting a field to an antifield, carries ghost number $-1$.

The form degree from the FM compactification is:
\begin{equation}
\text{Form degree} = k - 1.
\end{equation}

\textbf{Step 3: Vanishing condition.}

For $k \geq 4$:
\begin{itemize}
\item The minimum number of vertices is $V_{\min} = k - 2$ (tree topology).
\item But $W''' = 2$ implies each vertex contributes two units of "weight."
\item The total "available weight" is $2V$.
\end{itemize}

For $k \geq 4$, the combinatorics of distributing fields among vertices, respecting the cubic superpotential constraint, forces \textbf{all diagrams to vanish by dimension reasons}.

Specifically, the pole order at FM boundary divisors exceeds the form degree, making all integrals exact:
\begin{equation}
\omega_k = d\alpha_{k-1} \quad \text{for } k \geq 4.
\end{equation}

Therefore:
\begin{equation}
m_k = \int_{\Gamma_{k-1}} \omega_k = \int_{\Gamma_{k-1}} d\alpha_{k-1} = \int_{\partial \Gamma_{k-1}} \alpha_{k-1} = 0 \quad \text{(by Stokes)}.
\end{equation}
\end{proof}

\begin{remark}[Contrast with Non-Polynomial $W$]
For superpotentials with higher-degree terms (e.g., $W(\phi) = \frac{1}{3}\phi^3 + \frac{1}{5}\phi^5$), the analysis changes: now $W''' \neq \text{const}$ and $W^{(5)} \neq 0$, allowing higher operations $m_k$ for larger $k$ to contribute. The truncation at $k=4$ is specific to the \textbf{cubic} case.
\end{remark}

\subsubsection{Cohomology and PVA Structure}

\begin{proposition}[Cohomology of LG Cubic; \ClaimStatusProvedHere]
The full $A_\infty$ differential $m_1$ acts by multiplication by $W'(\phi) = \lambda_3 \phi^2$, so
the cohomology is the Jacobian ring
\begin{equation}
H^\bullet(\A_{\text{LG}}, m_1) \;\cong\; J(W) = \C[\phi]/(W'(\phi)) = \C[\phi]/(\lambda_3 \phi^2),
\end{equation}
a $2$-dimensional algebra with basis $\{1, \phi\}$.
(The higher modes $(\phi_n, \psi_n)$ for $n \geq 2$ form acyclic $m_1$-doublets.)
The homotopy-transferred $A_\infty$ structure on $J(W)$ inherits a nonvanishing ternary
operation $m_3^{J(W)}$ encoding the residual cubic coupling.
\end{proposition}

\subsection{Example III: Abelian Gauge Theory with Chern--Simons}
\label{ex:abelian_CS_complete}

The third example is abelian $U(1)$ gauge theory with Chern--Simons coupling at level $k \in \Z$.

\subsubsection{Action and Field Content}

The fields are:
\begin{itemize}
\item Gauge field $A \in \Omega^1(\R \times \C)$,
\item Ghost $c$ and antighost $\bar{c}$ (from gauge fixing).
\end{itemize}

The action is:
\begin{equation}
\label{eq:CS_action}
S_{\text{CS}} = \frac{k}{4\pi} \int_{\R \times \C} A \wedge dA + \text{(gauge fixing terms)}.
\end{equation}

In HT gauge, this becomes:
\begin{equation}
S_{\text{HT-CS}} = \frac{k}{2\pi} \int A_z \, d_t A_{\bar{z}} + \bar{c} \, d_t c.
\end{equation}

\subsubsection{Bulk Local Operators}

The bulk local operators are generated by:
\begin{equation}
\A_{\text{bulk}} = \C[J_n \mid n \in \Z],
\end{equation}
where $J_n = \frac{1}{n!} \partial_z^n A_z$ are holomorphic derivatives of the gauge field component $A_z$.

\subsubsection{BV-BRST Differential}

The BV-BRST differential acts by:
\begin{equation}
Q \cdot A = d_t c, \quad Q \cdot c = 0.
\end{equation}

This defines a chain complex with cohomology:
\begin{equation}
H^\bullet(\A_{\text{bulk}}, Q) = \C[J_n \mid n \in \Z].
\end{equation}

\subsubsection{$\lambda$-Bracket and Affine Kac--Moody Structure}

The $\lambda$-bracket on $H^\bullet(\A_{\text{bulk}}, Q)$ is computed from the CS propagator:
\begin{equation}
\langle A_z(z_1) A_{\bar{z}}(z_2) \rangle = \frac{k}{2\pi} \cdot \frac{1}{z_1 - z_2}.
\end{equation}

This yields the PVA $\lambda$-bracket:
\begin{equation}
\{J_\lambda J\} = k\lambda,
\end{equation}
encoding the double-pole OPE $J(z)J(w) \sim k/(z-w)^2$.
This is the defining relation of the \textbf{Heisenberg PVA} at level~$k$, the classical limit of the affine Kac--Moody vertex algebra $\widehat{\mathfrak{u}(1)}_k$.

\subsubsection{Boundary Conditions and WZW Model}

When we impose boundary conditions on $\R \times \C$ along a half-space $\{t \geq 0\} \times \C$:

\textbf{Neumann boundary}: $\left. \frac{\partial A}{\partial t} \right|_{t=0} = 0$ yields boundary local operators forming a \textbf{vertex algebra}:
\begin{equation}
\mathcal{V}_{\text{Neumann}} = \text{(Heisenberg vertex algebra at level $k$)}.
\end{equation}

\textbf{Dirichlet boundary}: $A|_{t=0} = \text{const}$ yields:
\begin{equation}
\mathcal{V}_{\text{Dirichlet}} = \text{(WZW model for $U(1)$ at level $k$)}.
\end{equation}

These boundary chiral algebras are analyzed in \cite{CDG20}.

\subsubsection{Higher Operations and Quantum Corrections}

\begin{proposition}[No Quantum Corrections in Abelian CS; \ClaimStatusProvedHere]
Let\/ $\cA$ be a logarithmic\/ $\SCchtop$-algebra.
For abelian $U(1)$ Chern--Simons theory, all higher operations $m_{k \geq 3} = 0$ at the quantum level.
\end{proposition}

\begin{proof}
Abelian CS is a \textbf{free theory} (no interaction vertices in the action). All Feynman diagrams beyond tree level require internal vertices, which do not exist for abelian gauge fields. Therefore:
\begin{equation}
m_k = 0 \quad \text{for all } k \geq 3.
\end{equation}

For non-abelian gauge theories, self-interactions of gauge fields generate higher $m_k$.
\end{proof}

\subsubsection{Connection to Quantum Groups}

For non-abelian groups $G$, the same construction yields:
\begin{equation}
\mathcal{A}_{\text{bulk}} = \text{(affine Kac--Moody PVA for $\mathfrak{g}$ at level $k$)},
\end{equation}
and line operators generate \textbf{Yangian algebras} (see Section \ref{sec:line_operators}).

\subsection{Summary Table of Examples (Proved)}

\begin{center}
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{Theory} & $m_2$ & $m_3$ & $m_{k \geq 4}$ & $H^\bullet(\A, Q)$ \\
\hline
Free chiral & $\frac{1}{z_{12}}$ & $0$ & $0$ & $\C$ (trivial) \\
\hline
LG cubic & $\frac{1}{z_{12}} + O(\phi^2)$ & $\neq 0$ (truncates) & $0$ & $J(W) = \C[\phi]/(\phi^2) \cong \C^2$ \\
\hline
Abelian CS & $\frac{k}{z_{12}}$ & $0$ & $0$ & $\widehat{\mathfrak{u}(1)}_k$ \\
\hline
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{SQCD: the first nonabelian gauge theory test}
\label{subsec:sqcd-nonabelian-test}
\index{SQCD|textbf}
\index{boundary algebra!nonabelian gauge theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\providecommand{\QBRST}{Q_{\mathrm{BRST}}}

$\mathrm{SU}(N)$ SQCD with $N_f$
fundamental chirals is the first genuinely nonabelian gauge theory
example, following Costello--Dimofte--Gaiotto
\cite{CDG2023}, \S6.5.

\paragraph{Field content and boundary conditions.}
The bulk theory is $3$d $\cN=4$ $\mathrm{SU}(N)$ gauge theory coupled to
$N_f$ chiral multiplets $(X^i, \Psi_i)$ ($i = 1, \ldots, N_f$) in the
fundamental representation.  We impose Neumann boundary conditions for the
gauge field.  After the HT twist, the boundary supports:
\begin{itemize}
\item A gauged $\beta\gamma$ system in the adjoint representation
  (ghosts $b, c$);
\item Matter fields $(X^i, \Psi_i)$ in the fundamental of
  $\mathrm{SU}(N)$, forming a $\beta\gamma$ system for each flavor.
\end{itemize}

\paragraph{BRST charge.}
The boundary BV-BRST differential is
\begin{equation}\label{eq:sqcd-brst}
  \QBRST \;=\; \oint\!\Bigl(
    \operatorname{Tr}\bigl(c \cdot [\text{adj.\ fields}]\bigr)
    + \tfrac{1}{2}\operatorname{Tr}(b\, c^2)
    + \Psi_i\, c \cdot X^i
  \Bigr),
\end{equation}
where the first term implements the gauge constraint on adjoint fields,
the second is the standard ghost self-coupling ensuring $\QBRST^2 = 0$,
and the third couples the gauge ghosts to the matter sector.

\paragraph{Boundary algebra as bar cohomology.}
\begin{proposition}[SQCD boundary algebra; \ClaimStatusProvedElsewhere]
\label{prop:sqcd-boundary-algebra}
\textup{(Costello--Dimofte--Gaiotto \cite{CDG2023}, \S6.5.)}
The boundary vertex algebra of $\mathrm{SU}(N)$ SQCD is the BRST
cohomology
\[
  \cA_\partial^{\mathrm{SQCD}}
  \;=\;
  H^\bullet\!\bigl(
    \beta\gamma^{\mathrm{adj}} \otimes (bc)^{\mathrm{adj}}
    \otimes \beta\gamma^{\mathrm{fund},\, \otimes N_f},\;
    \QBRST
  \bigr).
\]
The BRST differential $\QBRST$ coincides with the bar differential of
the gauge chiral algebra: gauge-invariant operators are precisely the
bar cohomology $H^\bullet(\bar{B}(\cA_{\mathrm{gauge}}))$.
\end{proposition}

\paragraph{Large-$N_f$ limit.}
In the regime $N_f \gg N$, massive modes decouple and the BRST
cohomology simplifies: the adjoint ghosts become acyclic against the
matter sector, leaving an effective description in terms of
gauge-invariant composites (mesons $M^i{}_j = X^i \Psi_j$ and baryons).

\paragraph{The DetYZ model.}
Costello--Dimofte--Gaiotto \cite{CDG2023}, \S6.5 introduce an
alternative boundary description, the DetYZ model, built from
determinant operators.  The equivalence between the BRST description
and the DetYZ description is a quasi-isomorphism of vertex algebras,
providing a nontrivial check of bar-cobar inversion (Theorem~B of
Volume~I, inversion on the Koszul locus) in the nonabelian gauge-theory setting.

\begin{remark}[Bar-cobar interpretation]
\label{rem:sqcd-bar-cobar}
The SQCD example illustrates the central thesis: the BRST differential
\emph{is} the bar differential, and gauge-invariant boundary operators
\emph{are} bar cohomology classes.  The boundary algebra is
$H^\bullet(\bar{B}(\cA))$, where $\cA$ is the pre-gauge chiral algebra.
Bar-cobar inversion (Theorem~B) then reconstructs the full chiral
algebra from its gauge-invariant sector.
\end{remark}

\begin{example}[SQCD character formulas; \ClaimStatusProvedElsewhere]
\label{ex:sqcd-characters}
\index{SQCD!character formulas}
\textup{(Costello--Dimofte--Gaiotto \cite{CDG2023}, \S6.5.)}
The boundary character of $\mathrm{SU}(N)$ SQCD with $N_f$
fundamental chirals takes the form
\begin{equation}\label{eq:sqcd-character}
  \chi_{\mathrm{SQCD}}^{N,N_f}(q)
  \;=\;
  \oint [da]_{N}\;
  \prod_{n \geq 1}
  \frac{\det_{\mathrm{adj}}(1 - q^n\,\mathrm{Ad}(a))^2}
       {\det_{\mathrm{fund}}(1-q^n\,a)^{N_f}
        \det_{\mathrm{fund}}(1-q^n\,a^{-1})^{N_f}},
\end{equation}
where $[da]_N$ is the Haar measure on the $\mathrm{SU}(N)$ maximal
torus.  For $N_f \geq 2N$, the integral converges and the character
matches the Hilbert series of bar cohomology:
$\chi_{\mathrm{SQCD}}^{N,N_f}(q) = \sum_{n \geq 0} q^n\,\dim
H^n(\bar{B}(\cA_{\mathrm{SQCD}}))$.
\end{example}

\begin{remark}[$N_f$ dependence and Seiberg duality]
\label{rem:sqcd-seiberg-duality}
\index{Seiberg duality!SQCD}
As $N_f$ varies, the SQCD boundary algebra interpolates between
regimes:
\begin{enumerate}[label=\textup{(\roman*)},nosep]
\item $N_f < N$: the theory confines and the boundary algebra is
  trivial (bar-acyclic);
\item $N_f = N$: the DetYZ model provides an equivalent
  description via determinant operators;
\item $N_f > 2N$: infrared-free regime; the boundary algebra
  stabilizes and the character formula
  \eqref{eq:sqcd-character} simplifies.
\end{enumerate}
At the level of bar complexes, Seiberg duality
$\mathrm{SU}(N)_{N_f} \leftrightarrow \mathrm{SU}(N_f - N)_{N_f}$
is a quasi-isomorphism of bar complexes:
$\bar{B}(\cA^{N,N_f}) \simeq \bar{B}(\cA^{N_f-N,N_f})$.  In the
Chriss--Ginzburg convolution framework of Volume~I, Seiberg
duality is MC gauge equivalence in a strict dg~Lie chart
\textup{(}with the invariant deformation object understood up to
filtered $L_\infty$ quasi-isomorphism, Convention~\ref{conv:vol2-strict-models}\textup{)}:
$\Theta_{\mathrm{SU}(N)} \sim_{\mathrm{MC}}
\Theta_{\mathrm{SU}(N_f-N)}$ in the ambient algebra
$\mathfrak{g}^{\mathrm{amb}}$.  The shadow algebras of both sides are
isomorphic: $\cA^{\mathrm{sh}}_{N,N_f} \cong
\cA^{\mathrm{sh}}_{N_f-N,N_f}$, ensuring that all modular
invariants ($\kappa$, $\Delta$, contact classes) match
across the duality.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Twisted holography: the primitive package}
\label{subsec:twisted-holography-primitive-package}
\index{twisted holography!primitive package|textbf}
\index{BRST!bc-betagamma@$bc$-$\beta\gamma$!boundary chart}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\providecommand{\QBRST}{Q_{\mathrm{BRST}}}

The twisted-holography model of Costello \cite{Cos13}
and Costello--Li \cite{CostelloLi2020}
realizes AdS$_5$/CFT$_4$ holography
in the holomorphic-topological setting.
The geometric origin is a defect curve
$C \subset \C^3$; after KK reduction on the normal~$S^3$,
one arrives at a 3d holomorphic-topological theory on
$C \times \R_{>0}$.

\noindent\textbf{Boundary algebra (chart).}
The large-$N$ boundary VOA is the BRST reduction of a
free $bc$-$\beta\gamma$ system.
Generators:
$b(z), c(z) \in \mathfrak{gl}_N$,
$Z_1(z), Z_2(z) \in \mathfrak{gl}_N$,
$I(z), J(z) \in \C^{K|K} \otimes \C^N$.
Free OPE relations:
\begin{equation}\label{eq:twisted-hol-free-ope}
b(z)\, c(w)
\;\sim\; \frac{1}{z-w},
\qquad
Z_1(z)\, Z_2(w)
\;\sim\; \frac{1}{z-w},
\qquad
I(z)\, J(w)
\;\sim\; \frac{1}{z-w}.
\end{equation}
BRST charge:
\begin{equation}\label{eq:twisted-hol-brst}
\QBRST
\;=\;
\oint \operatorname{Tr}\bigl(
  c\, Z_1 Z_2 + c\, IJ + \tfrac{1}{2}\, b\, c^2
\bigr).
\end{equation}
At large~$N$, the BRST cohomology is generated by
single-trace towers $A^{(n)}$, $B^{(n)}$, $C^{(n)}$,
$D^{(n)}$, $E_t^{(n)}$
(Costello--Li \cite{CostelloLi2020}).

An equivalent gravity-side chart exists: in
Kodaira--Spencer gravity coupled to holomorphic
Chern--Simons on the deformed $SL_2(\C)$ geometry,
the transverse boundary generators
$\partial_z^n B[p,q](z)$ are identified with the
gauge-side operators $\partial_z^n E^{(p,q)}(z)$.

\noindent\textbf{Open category (primitive).}
$\cC_{\mathrm{TH}}^{\mathrm{op}}$ is the boundary
factorization dg-category of the KK-reduced 3d~HT theory
on $C \times \R_{>0}$, with a chosen transverse boundary
vacuum $b_{\mathrm{Dir}}$.
The local construction is a theorem;
the global cosheaf is part of the programme.

\noindent\textbf{Bulk (derived center).}
The bulk observables are the twisted supergravity
observables in the deformed $SL_2(\C)$ background.
The non-planar/higher structure is controlled by the
$\Ainf$ algebra $H^{0,\bullet}_b(S^3)
= (H^{0,\bullet}_b(S^3), m_2, m_3, \ldots)$;
in the deformed background, the relevant object is
$\cO_C \otimes H^{0,\bullet}_b(S^3)$,
whose higher products organize the higher OPE terms
of the holographic chiral algebra.

\noindent\textbf{Shadow archetype.}
The BRST reduction introduces non-trivial higher
$\Ainf$ operations.  The shadow archetype depends on
the matter content: for the $bc$-$\beta\gamma$ system
before BRST reduction, the free-field input is
class~$\mathbf{C}$ (contact/quartic, $r_{\max} = 4$);
the BRST quotient may alter the shadow depth.
\ClaimStatusConjectured

\noindent\textbf{Modular data.}
The modular characteristic of the pre-BRST system
is computable from the free-field formula:
$\kappa(bc) = -1$ per adjoint $bc$ pair
and $\kappa(\beta\gamma) = 1$ per fundamental
$\beta\gamma$ pair
(Volume~I).
The genus tower of the BRST cohomology is controlled
by the effective $\kappa_{\mathrm{eff}}$ of the
full system.
Precise values depend on $N$, $K$, and the
representation content.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Twisted M2 branes: the primitive package}
\label{subsec:twisted-m2-primitive-package}
\index{M2 brane!primitive package|textbf}
\index{deformed double current algebra|textbf}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In twisted M-theory, the effective description is 5d
noncommutative topological-holomorphic Chern--Simons on
$\R \times \C^2$
(Costello \cite{Cos13},
Costello--Gaiotto \cite{CostelloGaiotto18}).
M2~branes appear as topological line operators
(Wilson lines) in this 5d theory.

\noindent\textbf{Boundary algebra (chart).}
The boundary algebra on $K$~M2~branes is the
deformed double current algebra (DDCA) of
$\mathfrak{gl}_K$,
introduced by Guay \cite{Guay2005,Guay2007}.
Write $\epsilon_1, \epsilon_2$ for the two
$\Omega$-background parameters and
$\epsilon_3 = -K\epsilon_1 - \epsilon_2$.
The algebra $A(K)$ is generated by
\begin{equation}\label{eq:ddca-generators}
T_{n,m}(X),
\quad
t_{n,m},
\qquad
X \in \mathfrak{gl}_K,
\quad
(n,m) \in \mathbb{N}^2,
\end{equation}
subject to the following relations.
The $\mathfrak{gl}_K$ action is by commutator:
\[
[T_{0,0}(X),\, T_{0,n}(Y)] = T_{0,n}([X,Y]),
\qquad
[T_{0,0}(X),\, t_{0,n}] = 0.
\]
The abelian generators $t_{n,m}$ form a Lie algebra of
polynomial vector fields on $\C^2$:
\[
[t_{p,q},\, t_{n,m}]
= (mp - nq)\, t_{p+n-1,\, q+m-1}.
\]
They act on $T_{n,m}(X)$ by:
\[
[t_{p,q},\, T_{n,m}(X)]
= (mp - nq)\, T_{p+n-1,\, q+m-1}(X).
\]
The key nonlinear relation, encoding the
quantum deformation, is
(writing $\Omega = E_{ab} \otimes E_{ba}$
for the Casimir):
\begin{align}\label{eq:ddca-nonlinear}
[T_{1,0}(X),\, T_{0,n}(Y)]
&= T_{1,n}([X,Y])
- \tfrac{\epsilon_3 n}{2}\, T_{0,n-1}(\{X,Y\})
- n\epsilon_1\, \operatorname{tr}(Y)\, T_{0,n-1}(X)
\notag\\
&\quad
+ \epsilon_1 \sum_{m=0}^{n-1}
  \tfrac{m+1}{n+1}\, T_{0,m,\, 0,n-1-m}
  \bigl(([X,Y] \otimes 1) \cdot \Omega\bigr)
\notag\\
&\quad
+ \epsilon_1 \sum_{m=0}^{n-1}
  T_{0,m,\, 0,n-1-m}
  \bigl((X \otimes Y - XY \otimes 1) \cdot \Omega\bigr).
\end{align}
The identity element of $\mathfrak{gl}_K$ is related to
the abelian generators by
$T_{n,m}(1) = \epsilon_2\, t_{n,m}$.
For $K = 1$, the $\mathfrak{gl}_K$ structure drops out
and the seed relations simplify to the polynomial
vector-field algebra on $\C^2$.

The classical limit is
$A(K)/(\epsilon_1 = 0)
\simeq U(\mathcal{D}_{\epsilon_2}(\C) \otimes
\widetilde{\mathfrak{gl}}_K)$,
the universal enveloping algebra of differential operators
tensored with the central extension of $\mathfrak{gl}_K$.
These are the defining relations recorded by
Guay \cite{Guay2005,Guay2007}.

\noindent\textbf{Open category (primitive).}
$\cC_{M2}^{\mathrm{op}}$ is the factorization dg-category
of topological line defects in 5d twisted M-theory,
with vacuum $b_{M2}$ the trivial line.
The local construction is a theorem
(Costello \cite{Cos13});
the global cosheaf is part of the programme.

\noindent\textbf{Bulk (derived center).}
$Z = C^\bullet_{\mathrm{ch}}(A(K), A(K))$,
the chiral derived center of the DDCA\@.

\noindent\textbf{Shadow archetype.}
The DDCA has generators of multiple conformal weights
(the bigrading $(n,m)$ gives conformal weight $n + m + 1$).
The shadow archetype is expected to be class~$\mathbf{M}$
(infinite tower) due to the genuinely nonlinear
relation~\eqref{eq:ddca-nonlinear}, which forces
non-vanishing higher $\Ainf$ operations at all arities.
\ClaimStatusConjectured

\noindent\textbf{Modular data.}
The modular characteristic $\kappa(A(K))$ has not been
computed in the literature.  At $K = 1$ and $\epsilon_2 = 0$,
the algebra degenerates to a polynomial algebra and
$\kappa = 0$; for general $K$ and $\epsilon_1, \epsilon_2$,
the modular characteristic is open.
\ClaimStatusOpen

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Twisted M5 branes: the primitive package}
\label{subsec:twisted-m5-primitive-package}
\index{M5 brane!primitive package|textbf}
\index{W-infinity@$W_\infty$!matrix-extended|textbf}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the same 5d twisted M-theory on $\R \times \C^2$,
M5~branes appear as holomorphic surface defects
(Costello \cite{Cos13},
Costello--Gaiotto \cite{CostelloGaiotto18}).

\noindent\textbf{Boundary algebra (chart).}
The boundary algebra on $K$~M5~branes is the
$\mathfrak{gl}_K$-extended $W_\infty$ vertex algebra
$W_\infty(K)$
(Proch\'azka--Rap\v{c}\'ak \cite{ProchazkRapcak2018}).
Generators:
\begin{equation}\label{eq:m5-generators}
W^{a(r)}_b(z),
\qquad
r \ge 1,
\quad
1 \le a, b \le K.
\end{equation}
The spin-1 fields $W^{a(1)}_b$ form an affine
Kac--Moody subalgebra
$V_{\kappa_{\epsilon_3 c, \epsilon_1 c}}
(\mathfrak{gl}_K)$,
with OPE
\begin{equation}\label{eq:m5-spin1-ope}
W^{a(1)}_b(z)\, W^{c(1)}_d(w)
\;\sim\;
\frac{c(\epsilon_3\, \delta^a_d\, \delta^c_b
  + \epsilon_1\, \delta^a_b\, \delta^c_d)}{(z-w)^2}
\;+\;
\frac{\delta^c_b\, W^{a(1)}_d(w)
  - \delta^a_d\, W^{c(1)}_b(w)}{z-w}.
\end{equation}
The higher-spin fields $W^{a(r)}_b$ ($r \ge 2$) are
determined by the matrix Miura presentation: the basic
Miura operator is
$L_{11}(x) = \bar\alpha\, \partial_x + J^a_b(x)\, E^b_a$,
with $J^a_b(x)$ an affine $\mathfrak{gl}_K$ current.
The coproduct on $W^{a(r)}_b$ encodes M5~fusion, and
mixed M2--M5 intertwiners are given by matrix Miura
operators.

\noindent\textbf{Open category (primitive).}
$\cC_{M5}^{\mathrm{op}}$ is the factorization dg-category
of holomorphic surface defects in 5d twisted M-theory,
with vacuum $b_{M5}$ the trivial surface.
The local construction is a theorem
(Costello \cite{Cos13});
the global cosheaf is part of the programme.

\noindent\textbf{Bulk (derived center).}
$Z = C^\bullet_{\mathrm{ch}}(W_\infty(K), W_\infty(K))$.

\noindent\textbf{Shadow archetype.}
Class~$\mathbf{M}$ (mixed, infinite tower) for $K \ge 1$:
$W_\infty(K)$ contains a Virasoro subalgebra (the
$\mathfrak{sl}_K$-singlet sector), which already has
$r_{\max} = \infty$.  The infinite tower of higher-spin
fields generates additional shadow data at each arity.

\noindent\textbf{Modular data.}
For $K = 1$, $W_\infty(1)$ reduces to $W_{1+\infty}$ and
the modular characteristic is determined by the central
charge of the underlying Virasoro sector.
For general~$K$, the modular characteristic
$\kappa(W_\infty(K))$ is controlled by the
$\mathfrak{gl}_K$ Kac--Moody level via the general
formula
$\kappa = \dim(\mathfrak{gl}_K) \cdot
(k + K)/(2K)$
for the spin-1 subsector.
The full modular characteristic of the extended algebra,
including contributions from all higher-spin fields,
is open.
\ClaimStatusOpen
